### 2021-11-10 paper-aicts-cherkashin.tex
Initial import

%% The first command in your LaTeX source must be the \documentclass command.
%%
%% Options:
%% twocolumn : Two column layout.
%% hf: enable header and footer.
\documentclass[
% twocolumn,
% hf,
]{ceurart}
\usepackage{amsmath,amssymb,longtable,hhline}
\usepackage{doi}
\def\doitext{DOI:}

%%
%% One can fix some overfulls
\sloppy

%%
%% Minted listings support
%% Need pygment <http://pygments.org/> <http://pypi.python.org/pypi/Pygments>
% \usepackage{listings}
\usepackage{minted}
\setminted{fontsize=\footnotesize,mathescape}
\usepackage{hyperref}

\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=true,           % non-Latin characters in Acrobat’s bookmarks
    pdftoolbar=false,        % show Acrobat’s toolbar?
    pdfmenubar=false,        % show Acrobat’s menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={},    % title
    pdfauthor={Evgeny Cherkashin},     % author
    pdfsubject={model driven architecture},   % subject of the document
    pdfnewwindow=true,      % links in new PDF window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=green,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=blue           % color of external links
  }

\usepackage{pifont}
\graphicspath{{pics/}}

%% auto break lines
% \lstset{breaklines=true}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% Rights management information.
%% CC-BY is default license.
\copyrightyear{2021}
\copyrightclause{Copyright for this paper by its authors.
  Use permitted under Creative Commons License Attribution 4.0
  International (CC BY 4.0).}

%%
%% This command is for the conference information
\conference{2nd International Workshop on Advanced Information and Computation Technologies and Systems, December XX--XX, 2021, Irkutsk, Russia}

%%
%% The "title" command
\title{Technologies of Semantic WEB as an environment of application development and integration}
%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
\author[1,2]{Evgeny A. Cherkashin}[%
orcid=0000-0003-2428-2471,
email=eugeneai@icc.ru,
url=https://github.org/eugeneai,
]
\address[1]{Matrosov Institute for System Dynamics and Control Theory of Siberian Branch of Russian Academy of Sciences, 134 Lermontov St, Irkutsk, 664033, Russian Federation}

\address[2]{Institute for Mathematics and Information Technologies, Irkutsk State University, 20~Gagarina Bulv, Irkutsk, 664003, Russian Federation}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX{} document is presented as an
  article formatted for publication by CEUR-WS in a conference
  proceedings. Based on the ``ceurart'' document class, this article
  presents and explains many of the common variations, as many well as
  of the formatting elements an author may use in the preparation of
  the documentation of their work.
\end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\begin{keywords}
  knowledge graph \sep
  geographical information system \sep
  semantic web \sep
  storing flexible data \sep
  one-page web application
\end{keywords}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

CEUR-WS's article template provides a consistent \LaTeX{} style for
use across CEUR-WS publications, and incorporates accessibility and
metadata-extraction functionality. This document will explain the
major features of the document class\footnote{You can use this
  document as the template for preparing your publication. We
  recommend using the latest version of the ceurart style.}.

If you are new to publishing with CEUR-WS, this document is a valuable
guide to the process of preparing your work for publication.

The ``\verb|ceurart|'' document class can be used to prepare articles
for any CEUR-WS publication, and for any stage of publication, from
review to final ``camera-ready'' copy with {\itshape very} few changes
to the source.

This class depends on the following packages
for its proper functioning:

\begin{itemize}
\item \verb|natbib.sty| for citation processing;
\item \verb|geometry.sty| for margin settings;
\item \verb|graphicx.sty| for graphics inclusion;
\item \verb|hyperref.sty| optional package if hyperlinking is required in
  the document;
\item \verb|fontawesome5.sty| optional package for bells and whistles.
\end{itemize}

All the above packages are part of any
standard \LaTeX{} installation.
Therefore, the users need not be
bothered about downloading any extra packages.

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \verb|\vspace| command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

\section{Template parameters}

There are a number of template
parameters which modify some part of the \verb|ceurart| document class.
This parameters are enclosed in square
brackets and are a part of the \verb|\documentclass| command:
% \begin{lstlisting}
%   \documentclass[parameter]{ceurart}
% \end{lstlisting}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item {\verb|twocolumn|}: Two column layout.
\item {\verb|hf|}: Enable header and footer\footnote{You can enable
    the display of page numbers in the final version of the entire
    collection. In this case, you should adhere to the end-to-end
    pagination of individual papers.}.
\end{itemize}

\section{Front matter}

\subsection{Title Information}

The titles of papers should be either all use the emphasizing
capitalized style or they should all use the regular English (or
native language) style. It does not make a good impression if you or
your authors mix the styles.

Use the \verb|\title| command to define the title of your work. Do not
insert line breaks in your title.

\subsection{Title variants}

% \verb|\title| command have the below options:
% \begin{itemize}
% \item \verb|title|: Document title. This is default option.
% \begin{lstlisting}
% \title[mode=title]{This is a title}
% \end{lstlisting}
% You can just omit it, like as follows:
% \begin{lstlisting}
% \title{This is a title}
% \end{lstlisting}

% \item \verb|alt|: Alternate title.
% \begin{lstlisting}
% \title[mode=alt]{This is a alternate title}
% \end{lstlisting}

% \item \verb|sub|: Sub title.
% \begin{lstlisting}
% \title[mode=sub]{This is a sub title}
% \end{lstlisting}
% You can just use \verb|\subtitle| command, as follows:
% \begin{lstlisting}
% \subtitle{This is a sub title}
% \end{lstlisting}

% \item \verb|trans|: Translated title.
% \begin{lstlisting}
% \title[mode=trans]{This is a translated title}
% \end{lstlisting}

% \item \verb|transsub|: Translated sub title.
% \begin{lstlisting}
% \title[mode=transsub]{This is a translated sub title}
% \end{lstlisting}
% \end{itemize}

\subsection{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification. Multiple authors may share one affiliation. Authors'
names should not be abbreviated; use full first names wherever
possible. Include authors' e-mail addresses whenever possible.

\verb|\author| command have the below options:

\begin{itemize}
\item \verb|style|: Style of author name (chinese)
\item \verb|prefix|: Prefix
\item \verb|suffix|: Suffix
\item \verb|degree|: Degree
\item \verb|role|: Role
\item \verb|orcid|: ORCID
\item \verb|email|: E-mail
\item \verb|url|: URL
\end{itemize}

Author names can have some kinds of marks and notes:
\begin{itemize}
\item affiliation mark: \verb|\author[<num>]|.
% \item email: \verb|\ead{<email>}|,
% \item url: \verb|\ead[url]{<url>}|.
\end{itemize}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}
\item Group the authors per affiliation.
\item Use an explicit mark to indicate the affiliations.
\end{enumerate}

Author block example:
% \begin{lstlisting}
% \author[1,2]{Author Name}[%
%     prefix=Prof.,
%     degree=D.Sc.,
%     role=Researcher,
%     orcid=0000-0000-000-0000,
%     email=name@example.com,
%     url=https://name.example.com
% ]

% \address[1]{Affiliation #1}
% \address[2]{Affiliation #2}
% \end{lstlisting}

% \subsection{Abstract and Keywords}

% Abstract shall be entered in an environment that starts
% with \verb|\begin{abstract}| and ends with
% \verb|\end{abstract}|.

% \begin{lstlisting}
% \begin{abstract}
%   This is an abstract.
% \end{abstract}
% \end{lstlisting}

% The key words are enclosed in a \verb|keywords|
% environment. Use \verb|\sep| to separate keywords.

% \begin{lstlisting}
% \begin{keywords}
%   First keyword \sep
%   Second keyword \sep
%   Third keyword \sep
%   Fourth keyword
% \end{keywords}
% \end{lstlisting}

At the end of front matter add \verb|\maketitle| command.

\section{Sectioning Commands}

Your work should use standard \LaTeX{} sectioning commands:
\verb|\section|, \verb|\subsection|,
\verb|\subsubsection|, and
\verb|\paragraph|. They should be numbered; do not remove
the numbering from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is not allowed.

\section{Tables}

The ``\verb|ceurart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed \textit{above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \verb|table| to enclose the table's contents and the
table caption. The contents of the table itself must go in the
\verb|tabular| environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table*}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table*}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \verb|table*| to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table}

\section{Math Equations}

You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}

A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \verb|math| environment,
which can be invoked with the usual
\verb|\begin| \ldots \verb|\end| construction or with
the short form \verb|$| \ldots \verb|$|. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX};
this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty} \frac{1}{n} = 0,
\end{math}
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}

A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \verb|equation|
environment. An unnumbered display equation is produced by the
\verb|displaymath| environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX{}; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty} \frac{1}{n} = 0.
\end{equation}
Notice how it is formatted somewhat differently in
the \verb|displaymath|
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  S_{n} = \sum_{i=1}^{n} x_{i} ,
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \lim_{x \to 0} (1 + x)^{1/x} = e
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{sample-franklin}
  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
    Ewing, Inc. [Public domain], via Wikimedia
    Commons. (\url{https://goo.gl/VLCRBB}).}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader. Figure captions go below the figure. Your figures should
also include a description suitable for screen readers, to
assist the visually-challenged to better understand your work.

Figure captions are placed below the figure.

\section{Citations and Bibliographies}

The use of Bib\TeX{} for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \verb|\end{document}|
command:
% \begin{lstlisting}
% \bibliography{bibfile}
% \end{lstlisting}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the Bib\TeX{} file.

\begin{acknowledgments}

  The results were obtained within the state assignment of the Ministry of Education and Science of Russia, the project ``Methods and technologies of a cloud-based service-oriented digital platform for collecting, storing and processing large volumes of multi-format interdisciplinary data and knowledge based on the use of artificial intelligence, a model-driven approach and machine learning'', No.~FWEW-2021-0005 (State registration No.~121030500071-2).

  The results obtained with the use of the network infrastructure of Telecommunication center of collective use ``Integrated information-computational network of Irkutsk scientific-educational complex'' (\url{http://net.icc.ru}).

The~R\&D on GIS-viewer of faults involved the Centre of Geodynamics and Geochronology equipment at the Institute of the Earth's Crust, Siberian Branch of the Russian Academy of Sciences (grant No.~075-15-2021-682).   This study direction was partially carried out within the basic budgetary research project ``Modern geodynamics, mechanisms of destruction of the lithosphere and hazardous geological processes in Central Asia'', No.~FWEF-2021-0009.

The~development of the~infrastructure for Mothur command transformation to Rapidminer dataflow diagram is supported by the project of Irkutsk scientific center of Siberian branch of Russian Academy of sciences, grant No 4.2.

\end{acknowledgments}
%%
%% Define the bibliography file to be used
% \bibliography{sample-ceur}

\begin{thebibliography}{99}

\bibitem{lod} C.~Bizer, T.~Heath, T.~Berners-Lee, Linked Data -- The Story So Far. Int. J. Semantic Web Inf. Syst., 5 (2009), pp.~1--22. \doi{10.4018/jswis.2009081901}

\bibitem{lunina} O.~V.~Lunina.  The digital map of the Pliocene–Quaternary crustal faults in the Southern East Siberia and the adjacent Northern Mongolia. Geodynamics \& Tectonophysics.  V.~7(3) (2016). pp.~407-434. (in Russian) \doi{10.5800/GT-2016-7-3-0215}

\bibitem{hogan} A.~Hogan, E.~Blomqvist, M.~Cochez, C.~D’Amato \emph{et al}. Knowledge Graphs, 2020. \url{https://arxiv.org/abs/2003.02320v5}

\bibitem{afs} A.~A.~Gladkov, O.~V.~Lunina. Cartographic service ``Activetectonics''. \url{http://activetectonics.ru/} (access date: 20-Sep-2021)

\bibitem{foss} M.~Leidig, R.~Teeuw. Free software: A review, in the context of disaster management. International Journal of Applied Earth Observation and Geoinformation, 42 (2015), pp.~49-56. \doi{10.1016/j.jag.2015.05.012}.

\bibitem{lgd} C.~Stadler, J.~Lehmann, K.~Höffner, S.~Auer. LinkedGeoData: A core for a web of spatial open data. Semantic Web 3 (2012) 333–354. \doi{10.3233/SW-2011-0052}

\bibitem{geolink} M.~Cheatham, A.~Krisnadhi, R.~Amini, P.~Hitzler, \emph{et al}. The GeoLink knowledge graph, Big Earth Data, 2:2 (2018), pp.~131-143. \doi{10.1080/20964471.2018.1469291}

\bibitem{abid} T.~Abid, H.~Zarzour. Integrating linked open data in geographical information system. Procs. of. International Conference on Information Technology for Organization Development. Oct 19-20, 2014, University of Tebessa, Tebessa, Algeria (2014).

\bibitem{iwaniak1} A.~Iwaniak, I.~Kaczmarek, M.~Strzelecki, J.~Lukowicz, P.~Jankowski. Enriching and improving the quality of linked data with GIS. Open Geosciences, Vol.~8, 1~(2016) pp.~323-336. \doi{10.1515/geo-2016-0020}

\bibitem{iwaniak17} A.~Iwaniak, M.~Leszczuk, M.~Strzelecki, F.~Harvey, I.~Kaczmarek. A Novel Approach for Publishing Linked Open Geodata from National Registries with the Use of Semantically Annotated Context Dependent Web Pages. International Journal of Geo-Information. 6, 252 (2017). \doi{10.3390/ijgi6080252}

\bibitem{zont19} E.~Cherkashin, A.~Shigarov, V.~Paramonov. Representation of MDA transformation with logical objects. International Multi-Conference on Engineering, Computer and Information Sciences (SIBIRCON), Novosibirsk, Russia. (2019) 0913--0918 \doi{10.1109/SIBIRCON48586.2019.8958008}

\bibitem{authoring} E.~Cherkashin, A.~Shigarov, V.~Paramonov, A.~Mikhailov, Digital archives supporting document content inference, Procs. of 42-nd International Convention on Information and Communication Technology Electronics and Microelectronics (MIPRO), May, 20–24, 2019. pp. 1037-1042. \doi{10.23919/MIPRO.2019.8757196}

\bibitem{gisviewer} E.~Cherkashin, O.~Lunina, L.~Demyanov, A.~Tsygankov. Web-GIS viewer for active faults data represented as a knowledge graph. Proceedings for 4th Scientific-practical Workshop Information Technologies: Algorithms, Models, Systems. Irkutsk, Russia, September 14, 2021. CEUR-WS.org, online \url{http://ceur-ws.org/Vol-2984/paper8.pdf}

\end{thebibliography}
\section{Online Resources}

The sources for the viewer are being developed at Github, URL:~\url{https://github.com/De17eon/GRL}.

\end{document}

%%
%% End of file

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper-aicts-cherkashin.tex"
%%% End:

### 2021-11-10 talk-aicts-2021-12-06.tex
Renaming presentation

\documentclass[10pt]{beamer}
\usepackage{amsmath,amssymb,longtable,hhline}
\usepackage{mathrsfs}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{anyfontsize}
\usepackage{minted}
\usepackage{alltt}

\usemintedstyle{tango}
\newcommand{\ltprgsize}{\fontsize{5}{5}\selectfont}
%\newcommand{\ltprgsize}{\footnotesize}
\setminted{fontsize=\footnotesize,mathescape}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=true,           % non-Latin characters in Acrobat’s bookmarks
    pdftoolbar=false,        % show Acrobat’s toolbar?
    pdfmenubar=false,        % show Acrobat’s menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={},    % title
    pdfauthor={Evgeny Cherkashin},     % author
    pdfsubject={model driven architecture},   % subject of the document
    pdfnewwindow=true,      % links in new PDF window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=green,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=blue           % color of external links
}

\usepackage{pifont}

\usetheme{Warsaw}
\usecolortheme{crane}
%\useinnertheme{rectangles}
%\setbeamertemplate{itemize item}{\scriptsize\hbox{\donotcoloroutermaths\ding{113}}}
\definecolor{darkding}{RGB}{200,56,0}
\setbeamertemplate{itemize item}{\scriptsize\hbox{\color{darkding}{\bfseries\ding{113}}}}
\setbeamertemplate{itemize subitem}{\tiny\raise1.5pt\hbox{\donotcoloroutermaths$\blacktriangleright$}}
\setbeamertemplate{itemize subsubitem}{\tiny\raise1.5pt\hbox{\donotcoloroutermaths$\blacktriangleright$}}
\setbeamertemplate{enumerate item}{\insertenumlabel.}
\setbeamertemplate{enumerate subitem}{\insertenumlabel.\insertsubenumlabel}
\setbeamertemplate{enumerate subsubitem}{\insertenumlabel.\insertsubenumlabel.\insertsubsubenumlabel}
\setbeamertemplate{enumerate mini template}{\insertenumlabel}

\beamertemplatenavigationsymbolsempty

\usepackage{iftex,ifxetex}
\ifPDFTeX
  \usepackag
%\useoutertheme{split}
%\useinnertheme{rounded}
\setbeamertemplate{background canvas}[vertical shading][bottom=white!80!cyan!20,top=cyan!10]
%\setbeamertemplate{sidebar canvas left}[horizontal shading][left=white!40!black,right=black]

\graphicspath{{pics/}}

\providecommand{\email}[1]{\texttt{#1}}
\usepackage{changepage}
\newcommand{\GB}[1]{\colorbox{green}{#1}}
\newcommand{\BB}[1]{\colorbox{blue}{#1}}
\newcommand{\RB}[1]{\colorbox{red}{#1}}
\newcommand{\btprgsize}{\fontsize{7}{7}\selectfont}

% --------------------------

\begin{document}

\Setbeamertemplatee[utf8]{inputenc}
  \usepackage[T1]{fontenc}
  \usepackage[russian]{babel}
  \usepackage{lmodern}
  \usefonttheme{serif}
\else
  \ifluatex
    \usepackage{unicode-math}
    \defaultfontfeatures{Ligatures=TeX,Numbers=OldStyle}
    \setmathfont{Latin Modern Math}
    \setsansfont{Linux Biolinum O}
    \setmonofont{Fira Mono}[Scale=MatchLowercase]
    \usefonttheme{professionalfonts}
    % \setmathfont[
    %     Ligatures=TeX,
    %     Scale=MatchLowercase,
    %     math-style=upright,
    %     vargreek-shape=unicode
    %     ]{euler.otf}
  \fi
\fi

%\useoutertheme{split}
%\useinnertheme{rounded}
\setbeamertemplate{background canvas}[vertical shading][bottom=white!80!cyan!20,top=cyan!10]
%\setbeamertemplate{sidebar canvas left}[horizontal shading][left=white!40!black,right=black]

\graphicspath{{pics/}}

\providecommand{\email}[1]{\texttt{#1}}
\usepackage{changepage}
\newcommand{\GB}[1]{\colorbox{green}{#1}}
\newcommand{\BB}[1]{\colorbox{blue}{#1}}
\newcommand{\RB}[1]{\colorbox{red}{#1}}
\newcommand{\btprgsize}{\fontsize{7}{7}\selectfont}

% --------------------------

\begin{document}

\setbeamertemplate{background canvas}[vertical shading][bottom=white,top=white]
\setbeamercolor{background canvas}{bg=white}

\title{Technologies of Semantic WEB as an environment of application development and integration}
\author[E.~Cherkashin]{\bfseries%
  Evgeny Cherkashin}
\institute{\normalsize Matrosov Institute for System Dynamics and Control Theory of Siberian Branch of Russian Academy of Sciences, Irkutsk, Russia\\%
  \email{\href{mailto:eugeneai@icc.ru}{eugeneai@icc.ru}}%
}
\date[2021]{AICTS'2021, December, 06, 2021 \\
Irkutsk, Russia}
%\date{\today}
\maketitle

\begin{frame}
\frametitle{Research and Development objectives}
\textbf{Main objective} of the activity is to construct data integration tools based on the \textbf{standardized} Semantic WEB technologies.

The following aspects are under consideration:
  \begin{enumerate}
  \item Ontological model representation
  \item Application data representation
  \item Model transformation
  \item Document publication
  \item Application integration via knowledge graph data documents
  \end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Representation of ontological models}
The ontologies are represented with \texttt{<subject,~predicate,~object>} \textbf{triples} as \textbf{graphs}, and there is frequently a \textbf{context}, the graph itself.
\begin{columns}
  \begin{column}{0.5\linewidth}
    \includegraphics[width=\linewidth]{SPO.pdf}
    \mbox{}\\[1em]
    \includegraphics[width=\linewidth]{socrates-human.pdf}
  \end{column}
  \begin{column}{0.5\linewidth}
    % Graph comprises
    % \begin{itemize}
    % \item T-Box
    % \item {\color{blue}{A-Box}}
    % \end{itemize}
    The subjects, the predicates and \emph{some} objects are \textbf{URI/IRI}.
    \emph{E.g.,} \texttt{http://purl.org/dc/terms/} defines the \textbf{namespace} ``\texttt{dc}''.\\[1ex]
    \emph{Other} subjects are \textbf{literals}.\\[1ex]
    All \textbf{XML} properties are applicable.
    \begin{itemize}
    \item XML format for data representation (optional!)
    \item global identification
    \item different specification usage in one document
    \end{itemize}
  \end{column}
\end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Blank nodes (BNodes)}
  As one can use only triples, we cannot represent \texttt{P(s,o1,o2,\ldots)}. We have to split it on triples and join them via a \textbf{Blank Node} (BNode), which has no special IRI.
  \begin{columns}
    \begin{column}{0.5\linewidth}
    \includegraphics[width=\linewidth]{bnode-ex.pdf}
    \end{column}
    \begin{column}{0.5\linewidth}
\begin{minted}[fontsize=\tiny]{text}
<http://example.org/web-data>
  dc:title "Web Data" ;
  ex:professor _:entity ;
  ex:students _:students ;
  ex:generatedBy _:activity1 .

_:entity
  ex:fullName "Alice Carol" ;
  ex:homePage <http://example.net/...
  ex:hasAddress _:address .

_:address
  a ex:Address ;
  ex:streetAddress "123 Main St." ;
  ex:postalCode "A1A1A1" ;
  ex:addressLocality "London" .

_:students
  a rdf:Bag ;
  ex:hasMember _:s1 ;
  ex:hasMember _:s2 .

_:activity1
  a ex:Event;
  ex:creator _:entity ;
  ex:atTime "Tuesday 11 February, 06:51:00 CST" .

_:activity2
  a ex:Event, ex:Update ;
  ex:actionOver _:activity1 ;
  ex:creator _:entity2 ;
  ex:atTime "Monday 17 February, 08:12:00 CST" .
\end{minted}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Data formats for graph representation}
  \begin{columns}
    \begin{column}{0.5\linewidth}
\begin{itemize}
\item N-Triples
\begin{minted}[fontsize=\tiny]{text}
<http://mythology.Greek.org/#Cronus>
    <http://www.example.org/schemas/relationship/fatherOf>
    <http://mythology.Greek.org/#Zeus>.
\end{minted}
\item Turtle
\begin{minted}[fontsize=\tiny]{text}
@prefix rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#> .
@prefix dc: <http://purl.org/dc/elements/1.1/> .
@prefix ex: <http://example.org/stuff/1.0/> .
<http://www.w3.org/TR/rdf-syntax-grammar>
  dc:title "RDF/XML Syntax Specification (Revised)" ;
  ex:editor [
    ex:fullname "Dave Beckett";
    ex:homePage <http://purl.org/net/dajobe/>
  ] .
\end{minted}
\item Notation 3 (N3)
\begin{minted}[fontsize=\tiny]{text}
@prefix dc: <http://purl.org/dc/elements/1.1/> .
<http://en.wikipedia.org/wiki/Tony_Benn>
  dc:title "Tony Benn" ;
  dc:publisher "Wikipedia" .
\end{minted}
\end{itemize}
\end{column}
\begin{column}{0.5\linewidth}
\begin{itemize}
\item RDF/XML
\begin{minted}[fontsize=\tiny]{xml}
<rdf:RDF
    xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
    xmlns:dc="http://purl.org/dc/elements/1.1/">
  <rdf:Description rdf:about="http://en.wikipedia.org/wiki/Tony_Benn">
    <dc:title>Tony Benn</dc:title>
    <dc:publisher>Wikipedia</dc:publisher>
  </rdf:Description>
</rdf:RDF>
\end{minted}
\item JSON-LD
\begin{minted}[fontsize=\tiny]{jsonld}
{
  "@context": {
    "name": "http://xmlns.com/foaf/0.1/name",
    "homepage": {
      "@id": "http://xmlns.com/foaf/0.1/workplaceHomepage",
      "@type": "@id"
    },
    "Person": "http://xmlns.com/foaf/0.1/Person"
  },
  "@id": "http://me.markus-lanthaler.com",
  "@type": "Person",
  "name": "Markus Lanthaler",
  "homepage": "http://www.tugraz.at/"
}
\end{minted}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[fragile]
\frametitle{Resource storage and access}
Semantic WEB documents are stored as \textbf{files}, \textbf{documents}, and, in general, [cloud] \textbf{resources} on servers.\\[1ex]
\begin{columns}
  \begin{column}{0.5\linewidth}
    Popular server software are
    \begin{itemize}
    \item Openlink Virtuoso (\texttt{DBPedia.org})
    \item Apache Jena (also a Java library)
    \item GraphDB (has good control interface)
    \item ClioPatria (not so popular, has integrated Prolog engine)
    \end{itemize}
  \end{column}
  \begin{column}{0.5\linewidth}
    \textbf{SPARQL} is a language to formulate questions (queries) for knowledge databases
\begin{minted}[fontsize=\tiny]{sparql}
SELECT ?publisher ?publisherLabel (AVG(?pages) AS ?avgPages)
WHERE
{
  ?book wdt:P123 ?publisher;
        wdt:P1104 ?pages.
  SERVICE wikibase:label { bd:serviceParam
        wikibase:language "[AUTO_LANGUAGE]". }
}
GROUP BY ?publisher ?publisherLabel
HAVING(COUNT(?book) > 1)
ORDER BY DESC(?avgPages)
\end{minted}
  \end{column}
\end{columns}
\vspace{2em}
Further info is at \url{https://www.w3.org/wiki/SparqlImplementations}.
\end{frame}

\begin{frame}[fragile]
  \frametitle{Ontological instruments:  ClioPatria}
  \begin{center}
    \includegraphics[width=0.8\linewidth]{pics/ClioPatria-shot.png}
  \end{center}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Semantic web technologies \& Knowledge graphs}
  Semantic Web (WEB 3.0) is characterized with
  \begin{itemize}
  \item Technological basis, oriented to the web % User applications are always comprises interconnected distributed web-services.
  \item Standardized data formats, storage, and processing
  \item Open principles of data publishing
  % \item Knowledge Graph (KG) construction techniques
  \item Services for data storage and access provision
  \item Generalized and special user interfaces are used for data presentation\vspace{1em}
  \end{itemize}
%\end{frame}

% Knowledge graphs changed the aspects to the knowledge base as being a part of whole totality of knowledge, implying the obeying the global standards and techniques of its acquisition and processing.

%\begin{frame}
%  \frametitle{Knowledge Graphs}
For the Knowledge Graphs (KG), the following is of interest.
 \begin{itemize}
  \item Converged notions \textbf{data} and \textbf{knowledge} as something is \textbf{known}
  \item Contain data, relations, and metadata (vocabularies)
  \item Distinguished \textbf{node filling in} and \textbf{processing} graph triples, \emph{e.g.}, withing SPARQL queries with UPDATEs
  \item Allow \textbf{postpone} the formal definition of a schema
  \item Three types of graph schemata: \textbf{semantic} (aimed at generalization), \textbf{validating} (\textbf{e.g.} semantics, \textbf{completeness} w.r.t. sets of relations), and \textbf{emergent} (infer a set of generalized structures and \textbf{reconstruct} the KG).
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Knowledge graph: Validating semantic example}
  \centering
  \includegraphics[width=0.9\linewidth]{verify-semantics.png}
\end{frame}

\def\textbfr#1{\textbf{{\color{red}#1*}}}

\begin{frame}
  \frametitle{Linked Open Data (LOD) star evaluation}
  Data are available in
  \begin{itemize}
  \item[\textbfr{1}] any format \textbf{openly}
  \item[\textbfr{2}] a \textbf{structured format}, such as Microsoft Excel file format (.xls)
  \item[\textbfr{3}] a \textbf{non-proprietary structured format}, such as .csv
  \item[\textbfr{4}] \textbf{W3C standards}, like using RDF and employing URIs
  \item[\textbfr{5}] a hypercontent form \textbf{having links to other Linked Open Data sources}
  \end{itemize}
  \begin{flushright}
   \includegraphics[width=0.7\linewidth]{5-star-lod.png}
  \end{flushright}
\end{frame}

\begin{frame}
  \frametitle{Information infrastructure for supporting Baikal microbiome research}
  \centering
  \includegraphics[width=\linewidth]{microbiome.png}
\end{frame}

\begin{frame}
  \frametitle{Microbiome study aims}
  \centering
  \includegraphics[width=\linewidth]{microbiome-study.png}
\end{frame}

\begin{frame}
  \frametitle{Microbiome study process}
  \centering
  \includegraphics[width=\linewidth]{microbiome-study2.png}
\end{frame}

\begin{frame}
  \frametitle{The aim of the research and development}

  The object of the research is genetic data processing. We would like to involve biologists in it. The subject is the amplicon data processing with MiSeq SOP\footnote{Standard Operational Procedure} (a technique).

  The primary \textbf{aim} of the research is to construct infrastructure which comprises
  \begin{itemize}
  \item Big Data database for sequence storage;
  \item metadata storage and adapters;
  \item visual construction of a processing model;
  \item cloud genetic data processing unit;
  \item metadata inference unit;
  \item data integration unit based on Semantic Web and Linked Open Data principles.
  \end{itemize}

\end{frame}

\begin{frame}
  \frametitle{The process of data analysis (MiSeq SOP)}
  \begin{enumerate}
  \item Reconstruct \emph{cotigs} (contiguous gene parts) from ``left'' and ``right'' \emph{readings}.
  \item Trim \textbf{bar-code} and other \emph{primers}.
  \item Filter sequences according to formal criteria (ambiguity, average length, maximal length of homopolymer).
  \item Classify \emph{unique} sequences and count their appearance in groups (samples).
  \item Alignment with reference sequences from SILVA database.
  \item Filter non-hanging sequences.
  \item Filter chimeras, fund unique sequences again.
  \item Classify sequences with respect to existing taxa hierarchy. Get \textbf{OTU}s.
  \end{enumerate}

  After these stages a large number of OTU\footnote{Operation Taxonomic Unit} classified has been obtained.
\end{frame}

\begin{frame}
  \frametitle{Dataflow representation of NGS analysis of amplicons}
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \begin{raggedright}
        \includegraphics[width=1\linewidth]{Dataflow-color-en.png}
      \end{raggedright}
    \end{column}
    \begin{column}{0.4\textwidth}\footnotesize
      \begin{tabular}{ll}
        Term & Description \\
        \hline
        NGS & New Generation\\ & Sequencing\\
        Amplicon & A DNA or RNA part \\
                 & copied many times \\
        Mothur & A software toolset for\\ & NGS research \\
        Rapidminer & A visual tool for \\
             & data mining modeling\\
             &  and execution
      \end{tabular}
      ${}$\\[1em]
      Green blocks are Mothur modules. Others are Rapidminer modules.
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Rapidminer module}
\begin{minted}[fontsize=\tiny]{cpp}
. . . vector<string> AlignCommand::setParameters(){ // PART OF MODULE SOURCE
try {
  CommandParameter ptemplate("reference", "InputTypes", "", "", "none", "none", "none","",false,true,true); parameters.push_back(ptemplate);
  CommandParameter pcandidate("fasta", "InputTypes", "", "", "none", "none", "none","fasta-alignreport-accnos",false,true,true); parameters.push_back(pcandidate);
  CommandParameter psearch("search", "Multiple", "kmer-blast-suffix", "kmer", "", "", "","",false,false,true); parameters.push_back(psearch);
  CommandParameter pksize("ksize", "Number", "", "8", "", "", "","",false,false); parameters.push_back(pksize);
  CommandParameter pmatch("match", "Number", "", "1.0", "", "", "","",false,false); parameters.push_back(pmatch);
// . . . . . . .
\end{minted}
  \begin{columns}
    \begin{column}{0.6\linewidth}
\begin{minted}[fontsize=\tiny]{java}
package com.rapidminer.ngs.operator; // GENERATED JAVA MODULE
// imports

class MothurChimeraCcodeOperator extends MothurGeneratedOperator {
  private InputPort fastaInPort = getInputPorts().createPort("fasta");
  private InputPort referenceInPort = getInputPorts().createPort("reference");
  private OutputPort chimeraOutPort = getOutputPorts().createPort("chimera");
  private OutputPort mapinfoOutPort = getOutputPorts().createPort("mapinfo");
  private OutputPort accnosOutPort = getOutputPorts().createPort("accnos");

  public MothurChimeraCcodeOperator (OperatorDescription description) {
    super(description);
  }
  @Override
  public void doWork() throws OperatorException {
    super();
    // . . . . . .
  }
  @Override
  public String getOutputPattern(String type) {
    if (type=="chimera") return
 "[filename],[tag],ccode.chimeras-[filename],ccode.chimeras";
    if (type=="mapinfo") return "[filename],mapinfo";
    if (type=="accnos") return
 "[filename],[tag],ccode.accnos-[filename],ccode.accnos";
    return super.getOutputPattern(type);
  }
}
\end{minted}
    \end{column}
    \begin{column}{0.4\linewidth}
\begin{minted}[fontsize=\tiny]{turtle}
@prefix xml: <http://www.w3.org/XML/1998/namespace> .
@prefix xsd: <http://www.w3.org/2001/XMLSchema#> .
ngsp:spec a ngsp:Specification ;
    ngsp:module mothur:NoCommand,
        mothur:align-check,
        mothur:align-seqs,
# . . . . .
mothur:align-check a ngsp:Module ;
    ngsp:outputPattern [ a cnt:Chars ;
            ngsp:parameterName "type" ;
            ngsp:pattern [ ngsp:patternString
                    "[filename],align.check" ;
                    dc:identifier "aligncheck" ] ;
            cnt:chars # . . . .
# . . . . .
mothur:align-check-idir-parameter a ngsp:Parameter ;
    ngsp:important false ;
    ngsp:multipleSelectionAllowed false ;
    ngsp:optionsDefault "" ;
    ngsp:required false ;
    ngsp:type mothur:String ;
    dc:title "inputdir" .

mothur:align-check-map-parameter a ngsp:Parameter ;
    ngsp:important true ;
    ngsp:multipleSelectionAllowed false ;
    ngsp:optionsDefault "" ;
    ngsp:required true ;
    ngsp:type mothur:InputTypes ;
    dc:title "map" .
# . . . . .
\end{minted}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Procedural data (Mothur tooling of Galaxy)}
  \begin{columns}[T]
    \begin{column}{0.5\linewidth}
\begin{minted}[fontsize=\tiny]{xml+cheetah}
<tool profile="16.07" id="mothur_make_contigs"
  name="Make.contigs" version="@WRAPPER_VERSION@.0">
  <description>Aligns paired ...</description>
  <!-- . . . . -->
  <command><![CDATA[ @SHELL_OPTIONS@
## Symlinks creation or On the fly ...
#if input_type.type == 'list_collection'
  #for pair in input_type.list_paired_collection:
    ln -s {pair.forward} `basename {pair.forward}` &&
    ln -s {pair.reverse} `basename {pair.reverse}` &&
    echo -e "{pair.name}\t`basename {pair.forward}`\t
    `basename {pair.reverse}`" >> combo_fastq.dat &&
  #end for ## . . . . . .
echo 'make.contigs(
  #if input_type.type == 'list_collection':
    file=combo_fastq.dat,
  #else:
    ffastq=ffastq.dat,
    rfastq=rfastq.dat,
  #end if  ## . . . . . . .
  gapextend=gapextend,
  rename=rename
  processors='{GALAXY_SLOTS:-8}'
)' | sed 's/ //g' | mothur | tee mothur.out.log
  ]]></command>
  <inputs>
    <conditional name="input_type">
      <param name="type" type="select" label="Select ...">
        <option value="regular" selected="true">Two ...</option>
        <option value="simple_collection">One pair ...</option>
        <option value="list_collection">Multiple ....</option>
      </param>
      <when value="regular">
        <param name="forward_fastq" type="data" />
        <param name="reverse_fastq" type="data" />
      </when>
    </conditional>
    <param name="align" type="select" label="...." help="">
      <option value="needleman" selected="true">needleman ...</option>
      <option value="gotoh">gotoh</option>
      <option value="kmer">kmer</option>
    </param>
  </inputs>
</tool>
\end{minted}
    \end{column}
    \begin{column}{0.5\linewidth}
\begin{minted}[fontsize=\tiny]{turtle}
@prefix dc: <http://purl.org/dc/elements/1.1/> .

[] a gal:Suite ;
 ngsp:module [ a gal:Module,
     ngsp:Module ;
   gal:command " ## . . . .  ";
   gal:exit_code [ gal:level "fatal" ;
     gal:range "1:" ] ;
       gal:inputs [ gal:checked "false" ;
         gal:conditional [ gal:param [ gal:help "" ;
           gal:option [ gal:value "yes" ;
              dc:description "yes" ],
                [ gal:value "no" ;
                  dc:description "no" ] ;
                dc:description "Trim with an oligos file?" ;
                dc:title "add" ;
              rdfs:range "select" ] ;
           gal:when [ gal:value "no" ],
           [ gal:param [ gal:min "0" ;
             gal:value "0" ;
             dc:description "pdiffs - number of differences to allow in the primer (default 0)" ;
             dc:title "pdiffs" ;
             rdfs:range "integer" ],
             [ gal:min "0" ;
               gal:value "0" ;
               dc:description "bdiffs - number of differences to allow in the barcode (default 0)" ;
               dc:title "bdiffs" ;
               rdfs:range "integer" ],
             [ gal:min "0" ;
               gal:value "0" ;
               dc:description "tdiffs - total number of differences to allow in primer and barcode (default 0)" ;
             dc:title "tdiffs" ;
             rdfs:range "integer" ] ] ] ]
   dc:identifier "mothur_make_contigs" ;
   dc:title "Make.contigs",
            "make.contigs" ;
   schema:sku 1 ''

\end{minted}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Model--Driven Architecture}

  \begin{columns}
    \begin{column}{0.4\textwidth}
      \includegraphics[width=1\linewidth]{mda-most-general.pdf}
    \end{column}
    \begin{column}{0.7\linewidth}
      \begin{description}
      \item[CIM] Computationally Independent Model;
      \item[CM] Model of Computations;
      \item[PIM] Platform Independent Model;
      \item[PM] Platform Model;
      \item[PSM] Platform--Specific Model;
      \item[CODE] Source code of software;
      \item[DATA] Initial database state.
      \end{description}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Architecture of transformation modules}
  \centering
  \includegraphics[width=0.9\linewidth]{architect_tree_pres-en-wo-OCL.pdf}
\end{frame}

\begin{frame}
  \frametitle{Model Driven Architecture and Linked Open Data}
  \begin{center}
    \includegraphics[width=0.9\linewidth]{mda-overview.pdf}
  \end{center}
\end{frame}

\begin{frame}[fragile]\frametitle{Architecture of services}
  \begin{columns}
    \begin{column}{0.7\linewidth}
      \includegraphics[width=1\linewidth]{architecture-mda-lod-ext-galaxy.pdf}
    \end{column}
    \begin{column}{0.3\linewidth}
      \textbf{Abbreviations}\\[1ex]\scriptsize
      T-Module is Transformation module\\
      MDA is Model-Driven Architecture\\
      CIM is Computationally Independent Model\\
      PIM is Platform Independent Model\\
      PSM is Platform Specific Model\\
      T-Box is Terminological Box\\
      A-Box is Instance Box\\
      A-Box is Instance Box\\
      NGS is Next-Generation Sequencing\\
      DB is Database
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Used ontologies}

  Standardized ontologies

  \begin{itemize}
  \item Friend-of-a-friend (\textbf{foaf}) for agent information: individuals, legal entities, program agents.
  \item Provenance (\textbf{prov}) for making references between documents.
  \item Dublin Core (\textbf{dc}) for published resource metadata mark up.
  \item DBPedia resource (\textbf{dbr}) to refer external classes and instance objects.
  \item Schema.org (\textbf{schema}) for Google, Yandex, Yahoo, \emph{etc}. searchable objects, structural elements.
  \item The Bibliographic Ontology (\textbf{bibo}) used for literature reference mark up.
  \item Open annotation (\textbf{oa}) as an ``bookmark'' ontology.
  \end{itemize}

  Non-standard ontologies

  \begin{itemize}
  \item Ontology \textbf{nssp} for Mothur source code processing results.
  \item Ontology \textbf{uml} for XMI representation.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Instrumentation: Ontology metadata server LOV}
  \begin{center}
    \includegraphics[width=0.8\linewidth]{lov-main.png}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Logtalk as transformation definition language}
  We have chosen Logtalk as it
  \begin{itemize}
  \item inherits widely known Prolog language syntax and runtime;
  \item implemented as macro package, performance penalties are about 1.5\%;
  \item has flexible semantics: we can define transformations and constraints within the same syntax;
  \item implement object-oriented knowledge (rules) structuring, encapsulation and replacement;
  \item compositional way of transformation implementation;
  \item powerful engine to post constraints on object-to-object messages (events);
  \item has implementation for many Prolog engines.
  \end{itemize}
  The <<regular>> language allow us to use its libraries not directly related to MDA transformations.
\end{frame}

\begin{frame}[fragile]
  \frametitle{RDF (TTL) representation and and its query object}
  \begin{columns}
    \begin{column}{0.4\linewidth}
\begin{minted}[fontsize=\tiny]{turtle}
@prefix xml: <http://www.w3.org/XML/1998/namespace> .
@prefix xsd: <http://www.w3.org/2001/XMLSchema#> .
ngsp:spec a ngsp:Specification ;
    ngsp:module mothur:NoCommand,
        mothur:align-check,
        mothur:align-seqs,
# . . . . .
mothur:align-check a ngsp:Module ;
    ngsp:outputPattern [ a cnt:Chars ;
            ngsp:parameterName "type" ;
            ngsp:pattern [ ngsp:patternString
                    "[filename],align.check" ;
                    dc:identifier "aligncheck" ] ;
            cnt:chars # . . . .
# . . . . .
mothur:align-check-idir-parameter a ngsp:Parameter ;
    ngsp:important false ;
    ngsp:multipleSelectionAllowed false ;
    ngsp:optionsDefault "" ;
    ngsp:required false ;
    ngsp:type mothur:String ;
    dc:title "inputdir" .

mothur:align-check-map-parameter a ngsp:Parameter ;
    ngsp:important true ;
    ngsp:multipleSelectionAllowed false ;
    ngsp:optionsDefault "" ;
    ngsp:required true ;
    ngsp:type mothur:InputTypes ;
    dc:title "map" .

mothur:align-check-name-parameter a ngsp:Parameter ;
    ngsp:chooseOnlyOneGroup "namecount" ;
    ngsp:important false ;
    ngsp:multipleSelectionAllowed false ;
# . . . . .
\end{minted}
    \end{column}
    \begin{column}{0.6\linewidth}
\begin{minted}[fontsize=\scriptsize]{logtalk}
:- object(query(_XMI)).
:- protected(xmi/1).
:- public([class/2, attribute/3, method/3]).
xmi(XMI) :- parameter(1, XMI).
    % Recognition of Class in RDF
class(Name, ID):-
    ::xmi(XMI),
    XMI::rdf(ID,rdf:type,uml:'Class'),
    XMI::rdf(ID,rdfs:label, literal(Name)).
    % Recognition of an attribute
attribute(Name, ClassID, ID):-
    ::xmi(XMI),
    XMI::rdf(ClassID, xmi:ownedAttribute, ID),
    XMI::rdf(ID, rdfs:label, literal(Name)).
    % Recognition of a method specification.
method(Name, ClassID, ID):-
    ::XMI(XMI),
    XMI::rdf(ClassID, xmi:ownedOperation, ID),
    XMI::rdf(ID, rdfs:label, literal(Name)).
% . . . . . . . . . . .
:- end_object.
\end{minted}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Code Block (idea is taken from \texttt{llvmlite}${}^*$)}
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \flushleft
\begin{minted}[fontsize=\footnotesize{}]{logtalk}
:- object(code_block, specializes(root)).
% Public interface of the object
:- public([append/1, prepend/1, clear/0,
   render/1, render_to/1, remove/1,
   item/1, items/1]).
% Code block items
:- dynamic([item_/1]).
:- private([item_/1]).
% Methods specialized during inheritance
:- protected([renderitem/2, render_to/2]).
% . . . . . . . . . . . .
% Delegate rendering to object itself
renderitem(Object, String):-
    current_object(Object), !,
    Object::render(String).
% Convert a literal to its string
% representation
renderitem(literal(Item), String):-!,
    atom_string(Item, String).
% Just print the item (debugging).
renderitem(Item, String):-
    root::iswritef(String, '%q', [Item]).
:- end_object.
\end{minted}
    \end{column}
    \begin{column}{0.4\textwidth}
      \includegraphics[width=1\linewidth]{code_block.pdf}
  ${}^*$) \url{https://github.com/numba/llvmlite}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{PSM of a Python Class as a specialization of Code Block}
%\begin{multicols}{2}
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \flushleft
\begin{minted}[fontsize=\scriptsize]{logtalk}
:- object(class, specializes(code_block),
   imports([named])). % Category of named entities
:- public([classlist/1, methods/1, attributes/1]).
% . . . . . . . . . . . . . .
renderitem(Item, Result):-      % proceed with default
    ^^renderitem(Item, Result). % rendering
render(Result):-         % Source generator
    ^^render(Name),      % implemented in a category
    ( ::item(classlist(List)) ->
     % . . . . . . . . . . .
        [Name]) ),
    ( ::item(attributes(Attributes))->
     % . . . . . . . . . . .
        [DefAttrList]),
      Attributes::items(InstanceAttrs),
      findall(S, ( % initialize attributes
         % . . . . . . . . .
         ), AttrAssigns),
        root::unindent,
        AttrList=[ConstructorDef|AttrAssigns];
         % . . . . . . . . .
        AttrList=[ConstructorDef, Pass] ),
    ( ::item(methods(Methods))-> % If any ...
      Methods::render(MethodList);
      MethodList=[] ),
    lists::append(AttrList,MethodList,StringList),
    root::unindent, Result=[Signature|StringList].
:- end_object.
\end{minted}
    \end{column}
    \begin{column}{0.4\linewidth}
      \includegraphics[width=1\linewidth]{code_block_class.pdf}
    \end{column}
  \end{columns}
  % \end{multicols}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Logtalk Categories}
  A category of named entities
\begin{minted}[fontsize=\scriptsize]{logtalk}
:- category(named).
:- public([name/1, render/1]).
:- protected([renderitem/2]).
name(Name):- ::prepend(name(Name)).
renderitem(name(Name), String):-!, atom_string(Name, String).
render(String):-  % What is code generation from items
    ::item(name(Name)), ::renderitem(name(Name), String).
:-end_category.
\end{minted}
Category of named and typed entities
\begin{minted}[fontsize=\scriptsize]{logtalk}
:- category(namedtyped, extends(named)).
:- public([type/1,render/2, separator_option/2,list_separator/1]).
:- protected([renderitem/2]).
type(Type):- ::append(type(Type)).
renderitem(Item, String):- ^^renderitem(Item, String),!.
renderitem(type(Type),String):-!, ::list_separator(Separator),
    writef::swritef(String, '%w%w', [Separator, Type]).
render(Middle, String):- ^^render(SName),
    (   ::item(type(Type)) ->
        ::renderitem(type(Type), SType),
        string_concat(SName, Middle, _1),
        string_concat(_1, SType, String) ;
        SName = String  ).
render(String):-  ::render("", String).
list_separator(Separator):-
    ::separator_option(Name, Default),!, % Global options
    root::option(Name, Separator, Default).
:- end_category.

\end{minted}
\end{frame}

\begin{frame}
  \frametitle{Discussion (MDA application)}
  Interesting positive impressions obtained:
  \begin{itemize}
  \item Logtalk and RDF are flexible, sufficiently universal and convenient implementation infrastructures for MDA;
  \item The best implemenation means is Prolog predicate wrapping and Logtalk object encapsulation of rules;
  \item Not all Logtalk properties are investigated: there might be more sophisticated programming techniques developed, \emph{e.g.}, on the base of message watchers.
  \end{itemize}
  Technical problems making the approach somewhat problematic:
  \begin{itemize}
  \item Very simple tasks take too much efforts, \emph{e.g.}, text processing: convert an identifier into the CamelCase;
  \item It takes too long to surf Internet in order to find a vocabulary for a domain, but it is more productive than development new one and classes;
  \item Prolog is not a popular language in MDA, neither Logtalk.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Future activities}
  The future activities supposed to be as follows:
  \begin{enumerate}
  \item Having dataflow models of MiSeq SOP and other techniques, device an intelligent subsystem, which will construct computational procedures for a predefined set of data processing tasks (\emph{AI's problem solving}).
  \item Integrate dataflow and data storage with Galaxy.
    \begin{itemize}
       \item Realize adapters of data/metadata storage and retrieving, as well as the storage.
       \item Create a more sophisticated source code parser or PIM model of computation so we will able to infer metadata for the output on the metadata of input (partially done in November).
         % \item Adapt our ``pattern-directed'' approach of semantically marked up document authoring to present input and output to the scientific communities.
       \item Adapt our document authoring tools to Galaxy allowing LOD representation of results.
  \end{itemize}
  \item Implement integration to biological/gene databases.
  \item Write a handbook on Logtalk programming strategies with its author Paulo Mora.
  \end{enumerate}
\end{frame}

% \begin{frame}
%   \frametitle{Conclusion}
%   The following results have been obtained as for today:
%   \begin{itemize}
%   \item Biologists' activities are mastered, investigated and regular patterns are described.
%   \item A technique for interpretation of Mothur interfaces has been developed, implemented, refined and extended.
%   \item Transformation tools are tested in application areas and no significant technical problems were detected.
%   \item A technique of document authoring is being developed and adapted to the domain.
%   \item Integration with Galaxy project is the primary aim of the future development.
%   \end{itemize}
%   The source codes are available at \url{https://github.com/isu-enterprise/icc.xmitransform}, \url{https://github.com/eugeneai/icc.mothurpim}.

%   This research is supported by Irkutsk scientific center of SB RAS, project No 4.2;
% \end{frame}

\begin{frame}
  \frametitle{Discussion (MDA application)}
  Interesting positive impressions obtained:
  \begin{itemize}
  \item Logtalk and RDF are flexible, sufficiently universal and convenient implementation infrastructures for MDA;
  \item The best implemenation means is Prolog predicate wrapping and Logtalk object encapsulation of rules;
  \item Not all Logtalk properties are investigated: there might be more sophisticated programming techniques developed, \emph{e.g.}, on the base of message watchers.
  \end{itemize}
  Technical problems making the approach somewhat problematic:
  \begin{itemize}
  \item Very simple tasks take too much efforts, \emph{e.g.}, text processing: convert an identifier into the CamelCase;
  \item It takes too long to surf Internet in order to find a vocabulary for a domain, but it is more productive than development new one and classes;
  \item Prolog is not a popular language in MDA, neither Logtalk.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Document authoring and storage}
  In most cases documents are created as a result of
  \begin{itemize}
  \item creative activity of a person with a text processors (authoring);
  \item printing a digital copy or a data record in a database;
  \item aggregation operation over database records (report).
  \end{itemize}
  Then it is stored either as a physical paper and/or a digital document (PDF, DOCX, HTML).

  Since 2000-th, Semantic Web and Linked Open Data (LOD) is being developed, allowing
  \begin{itemize}
  \item structural storage of data within published documents;
  \item processing stored data computationally;
  \item integration of data structures and data objects globally.
  \end{itemize}

  The \textbf{aim of this research} is to develop technologies, software and services allowing construction of digital archives supporting document data inclusion and inference from existing documents.
\end{frame}

\begin{frame}
  \frametitle{Structure of a document}
  \centering
  \includegraphics[width=1\linewidth]{document-structural-view.pdf}
\end{frame}

\begin{frame}
  \frametitle{Open Annotation (oa)}
\begin{adjustwidth}{-3em}{-3em}
  \centering
 \includegraphics[width=1\linewidth]{Open-Annotation_CB_Bookmarking_and_Semantically_Tagging_A_webpage_spec20130128.png}
\end{adjustwidth}
\end{frame}

\begin{frame}
  \frametitle{Architecture}
  \begin{adjustwidth}{-2.5em}{-2.5em}
    \begin{center}
      \includegraphics[width=0.5\linewidth]{architecture-mda-lod-ext-tot.pdf}
    \end{center}
  \end{adjustwidth}

% \begin{itemize}
% \item content and metadata repository with SPARQL (6) and full-text search (3); the engines for LOD processing (5) based on the logical inference;
% \item service for analysis of the stored content (2) enriching the archive with semantic information describing the content;
% \item tools for development of LOD applications and their user interfaces (1);
% \item browser based authoring tools (4).
% \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Generated list of title page preambles}
    \begin{adjustwidth}{-2.5em}{-2.5em}
    \begin{center}
      \includegraphics[width=0.7\linewidth]{template-title-pages.jpg}
    \end{center}
  \end{adjustwidth}
\end{frame}

\begin{frame}
  \frametitle{Generated part of a study program}
   \begin{adjustwidth}{-2.5em}{-2.5em}
    \begin{center}
      \includegraphics[width=0.7\linewidth]{template-courses.jpg}
    \end{center}
  \end{adjustwidth}
\end{frame}
\begin{frame}
  \frametitle{Imported time distribution for lecture, seminary, \ldots}
   \begin{adjustwidth}{-2.5em}{-2.5em}
    \begin{center}
      \includegraphics[width=0.9\linewidth]{work-program-volume.jpg}
    \end{center}
  \end{adjustwidth}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Representation of  document parts with RDFa}
  % \begin{block}{}
  %   \textbf{Целью} исследования является создание методики разработки процедур трансформации (PD\footnote{Platform [description] model.}) в виде ОО-модулей.
  % \end{block}
  % \begin{columns}
  %   \begin{column}{0.5\linewidth}
  %     % \includegraphics[width=1\linewidth]{pics/scenario-ru-wo-mothur.pdf}
  %   \end{column}
  %   \begin{column}{0.6\linewidth}
  %     Задачи исследования:
  %     \begin{itemize}
  %     \item Изучить синтаксические структуры Logtalk в аспекте структурирования знаний;
  %     \item Предложить методику представления трансформации в виде ОО-модулей;
  %     \item Реализовать библиотеку объектов и классов для МДА;
  %     \item Тестирование библиотеки на примере.
  %     \end{itemize}
  %   \end{column}
  % \end{columns}

\begin{adjustwidth}{-1.5em}{-1.5em}
\begin{minted}[escapeinside=||,fontsize=\btprgsize]{text}
<html lang="ru" xmlns=http://www.w3.org/1999/xhtml
|\GB{xmlns:taa}|=http://irnok.net/engine/rdfa-manipulation
xml:lang="ru" metal:define-macro="page">
<head> . . . . </head>
<body prefix="rdf: http://www.w3.org/1999/...-ns# foaf: http://xmlns.com/foaf/...
imei: imei.html# course: https://irnok.net/college/plan/01..16-...\
%D0\%BA_PB-SM.plm.xml.xlsx-....2.3.1.html#"  resource="#post"
typeof="schema:CreativeWork sioc:Post prov:Entity">
<!-- The application control panel -->

<main lang="ru" resource="#annotation" typeof="oa:Annotation" id="main-doc-cnt">
<div property="oa:hasTarget" resource="#course-work-prog"></div>
<article property="oa:hasBody" typeof="foaf:Document curr:WorkingProgram"
         resource="#course-work-program" id="main-document">
  <div |\GB{taa:content}|="imei:title-page"></div>
  <div |\GB{taa:content}|="imei:neg-UMK"></div>
  <section id="TOC" class="break-after"> <h2>Table of Contents</h2>
    <div id="tableOfContents"></div>
  </section>
  <section id="course-description" resource="#description"
           property="schema:hasPart" typeof="schema:CreativeWork">
    <div property="schema:hasPart" resource="#purpose"
         typeof="dc:Text cnt:ContentAsText" >
      <div property="cnt:chars" datatype="xsd:string">
        <h2 property="dc:title" datatype="xsd:string">
           Aims and objectives of the discipline (module)</h2>
        <p>The aim of teaching the discipline ...</p>
      </div>
   </div>
  . . . . . . . .
\end{minted}
\end{adjustwidth}
\end{frame}

\begin{frame}
  \frametitle{Complete document}
  \begin{adjustwidth}{-2.5em}{-2.5em}
    \begin{center}
      \begin{columns}
        \begin{column}{0.5\linewidth}
          \includegraphics[width=1\linewidth]{work-program-title.jpg}
        \end{column}
        \begin{column}{0.5\linewidth}
          \includegraphics[width=1\linewidth]{work-program-agreement.jpg}
        \end{column}
    \end{columns}
    \end{center}
  \end{adjustwidth}
\end{frame}

\begin{frame}
  \frametitle{Discussion}
A tools (components) for digital archive implementation, which allows
to device information systems and document processing services
with the following features:
\begin{itemize}
\item load LOD marked up document, extract, store in a graph and index RDF data;
\item retrieve RDF data as triples or as a result of full-text search query;
\item combine existing LOD data and its content in new documents dynamically with browser based context inference machine;
\item use server-site inference machine (Prolog) to process RDF data upon
  request from browser's part of the system;
  % \item organizes a platform for document semantic markup,
\item convert created RDFa marked up HTML5 documents into Excel and Word formats.
\end{itemize}

  \textbf{Applications}
  \begin{itemize}
  \item Document authoring automation;
  \item Context-depended editing;
  % \item Export into office formats;
  \item Self-organizing global document flows;
  \item Documents as data sources for information systems.
  \end{itemize}

\end{frame}

\begin{frame}
\frametitle{Cartographical WEB-service with knowledge graph of South-Siberian faults}
\textbf{Aim} is to construct a WEB-GIS browser for faults stored in the KG.
\begin{itemize}
\item Scalability to external data with converters (TODO)
\item Interdisciplinary data representation
\item Application development with nowadays WEB techniques
\item Digital platform for data publication in ``Digital Baikal'' project
\end{itemize}
% \end{frame}

% \begin{frame}
% \frametitle{Key points}
Development \textbf{plan}
\begin{itemize}
\item Investigate the current data formats
\item Develop T-Box
\item Fill in A-Box
\item Expose the KG with a server
\item Implement browsing SPARQL query results with GIS
\item Develop object browser
\end{itemize}
MVP is a WEB-GIS with the most of the listed features.
\end{frame}

\begin{frame}
  \frametitle{GIS source data table properties}
  \begin{itemize}
  \item Only one table, one row for each GIS object
  \item There are many \texttt{NULL} values
  \item More than 1000 objects
  \item More than 70 attributes (according to O.V.~Lunina, PhD)
  \end{itemize}
  \begin{center}
    \includegraphics[width=1\linewidth]{gis-fault-table.png}
  \end{center}
\end{frame}

\begin{frame}
\frametitle{Developed ontology}
The ontology contains nonintersection properties for its classes
\begin{center}
    \includegraphics[width=1\linewidth]{fault-onto.png}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Serving ontology and its A-box}
As server \texttt{GraphDB} is used.
\begin{center}
    \includegraphics[width=1\linewidth]{graph-db.png}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Web GIS}
\begin{center}
    \includegraphics[width=1\linewidth]{gis-ex.png}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Used technologies for constructing WEB-GIS browser}
\begin{center}
    \includegraphics[width=1\linewidth]{gis-tech.png}
\end{center}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Ontological instruments: editor Proteg\'e}
  \begin{columns}
    \begin{column}{0.4\textwidth}
      \includegraphics[width=1\linewidth]{pics/NLP-fault-tbox-shot.png}
    \end{column}
    \begin{column}{0.6\textwidth}
      \includegraphics[width=1\linewidth]{pics/NLP-fault-abox-shot.png}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Modification of GeoBase supporting Semantic WEB}
\begin{minted}{prolog}
schema('fault','in','continent').  % Connect our relations with GeoBase
schema('fault','with','feature').  % vocabulary.
schema('name','of','fault').       %
% schema('feature','of','fault').  % This relation is already in the T-Box''

% <<Загрузка>> онтологий
schema(Prop, 'of', SubjName):-     % used on translation stage
        var(SubjName),
        geob_prop(Prop,_).

schema(Prop, 'of', SubjName):-     % used on stage of interpretation
        nonvar(SubjName),          % a Class is supplied
        geob_prop(Prop, GProp),
        geob_ent_class(SubjName, Subj),
        rdf_reachable(Subj, rdfs:subClassOf, Parent),
        rdf(GProp, rdfs:domain, Parent),!.

geob_prop(Prop, GProp):-           % Property check
        rdf_global_id(geob:Prop, GProp),
        rdf(GProp,rdf:type,owl:'ObjectProperty'),!.

geob_class(Class, GClass):-        % Class check
        rdf_global_id(geob:Class, GClass),
        rdf(GClass,rdf:type,owl:'Class'),!.

geob_ent_class(Ent, Class):-
        sub_atom(Ent,0,1,R,H),
        sub_atom(Ent,1,R,0,T),
        string_upper(H,U),
        atom_concat(U,T,GEnt),
        geob_class(GEnt, Class).

geob_ent(E, A):-
        nonvar(E),
        geob_ent_class(E,Class),
        rdf(A, rdf:type, Class, geodata). % A Subject of geodata
\end{minted}
\end{frame}

\begin{frame}[fragile]
  \frametitle{GeoBase to ActiveFaults Natural language interface}
  \begin{center}
    \includegraphics[width=1\linewidth]{Geobase-ui.png}
  \end{center}
  \textbf{New problem} for student graduation project: Implement Natural Language to SPARQL translator.
\end{frame}

\begin{frame}
  \frametitle{Conclusion}
  The following results have been obtained as for today:
  \begin{itemize}
  \item A technique for \textbf{domain} model representation has been developed and tested.
  \item A programming technique using object-oriented logical language Logtalk is devised.
  \item Prototypes of various transformation procedures are implemented.
  \item Transformation tools are tested in application areas and no significant technical problems were mentioned.
  \end{itemize}
  Further development directions are as follows:
  \begin{itemize}
  \item A technique for document automatic markup with vocabulary entities.
  \item A transformation implemenation techniques, minimizing usage of dynamic objects, targeting on macro properties of Logtalk.
  \item Form a toolset out of existing prototypes obeying nowadays software development requirements.
  \end{itemize}
  The source codes are available at \url{https://github.com/isu-enterprise/icc.xmitransform}, \url{https://github.com/eugeneai/icc.mothurpim}.
\end{frame}

\begin{frame}
  \frametitle{Technologies used (open source)}
  Python-3.x.x (\url{http://python.org}),\\
  ZCA (\url{https://muthukadan.net/docs/zca.html}),\\
  SWIG (\url{http://swig.org/}),\\
  SWI-Prolog (\url{https://www.swi-prolog.org/}),\\
  Logtalk (\url{https://logtalk.org/}),\\
  ClioPatria (\url{https://cliopatria.swi-prolog.org/home}),\\
  Pengines (\url{https://pengines.swi-prolog.org/docs/index.html}),\\
  LOV (\url{https://lov.linkeddata.es/dataset/lov/}),\\
  Elastic Search (\url{https://www.elastic.co/}),\\
  Kyotocabinet (\url{https://fallabs.com/kyotocabinet/}),\\
  DBPedia (\url{https://wiki.dbpedia.org/}),\\
  Mothur (\url{https://mothur.org/}),\\
  Galaxy (\url{https://usegalaxy.org/}),\\
  R (\url{https://www.r-project.org/}), \\
  Dust.js (\url{https://akdubya.github.io/dustjs/})
\end{frame}

\begin{frame}
  \begin{center}
  \Large Thanks for Your Attention!
\end{center}
\end{frame}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2021-11-10 paper-aicts-cherkashin.tex
Added an abstract.

@@
 \begin{abstract}
-  A clear and well-documented \LaTeX{} document is presented as an
-  article formatted for publication by CEUR-WS in a conference
-  proceedings. Based on the ``ceurart'' document class, this article
-  presents and explains many of the common variations, as many well as
-  of the formatting elements an author may use in the preparation of
-  the documentation of their work.
+  Semantic Web technologies and standards are becoming one of productive assets for software development and integration on various levels of design and implementation.
+This include OSI level 6 data representation, database and file storage formats, the source data for indexes, hypertext markup, source data for document authoring, representing abstract software models, \emph{etc}.
+They also smoothly combined with development and data analysis tools, giving raise a new common-ground environment for realizing integration on the semantic level.
+
+  This paper presents the view on the Semantic Web technologies as data representation for software development tools with model analysis, transformation and retrospection capabilities.
+Three examples are presented, exposing our experience.
 \end{abstract}
@@
   \centering
-  \includegraphics[width=\linewidth]{sample-franklin}
+  % \includegraphics[width=\linewidth]{sample-franklin}
   \caption{1907 Franklin Model D roadster. Photograph by Harris \&

### 2021-11-10 paper-aicts-cherkashin.tex
Revised keywords.

@@
   knowledge graph \sep
-  geographical information system \sep
   semantic web \sep
-  storing flexible data \sep
-  one-page web application
+  model transformation \sep
+  model-driven architecture \sep
+  object-oriented logical programming \sep
+  web application \sep
+  geographical information system
 \end{keywords}

### 2021-11-11 paper-aicts-cherkashin.tex
Added some into paragraphs.

@@
 \section{Introduction}
+
+Since the beginning of 2000s the development of the Semantic Web technologies became [awesome] mostly in describing software domains, representing semantics withing data objects, and define first-order knowledge \cite{}.
+A number of tools for automation of conceptual model design and model conversion were proposed as well \cite{}.
+The R\& D at present is carried on in developing applications, improving and standardizing vocabularies, devising tools for solving particular problems.
+
+Unification of routine tasks resulted in development common techniques for storing and processing data, which are aggregated around term ``knowledge graph'' (KG).
+These include data representation formats, vocabulary standards, query languages, tools for accessing and management KG.
+Software development techniques are mediated with properties of the new-created resources and their APIs\footnote{Abbreviation of Application Programming Interface}.
+Many internet resources provide common ways of formalized domain data access using query endpoints, and these endpoints allow one to organize federated data access.
+
+[....
+About itch of personal vocabularies development instead to utilize similar ones ..
+Problem of searching resources ....
+metadata storages ...
+a general vocabularies ...
+Term ontology = vocabulary ]
+
+[we have by our opiinion a good examples of SW and KG applications in the following fields ....
+]  [ ...
+The KG and SW tech are well integrated with existing knowledge based AI systems [ if not use the equivalence]].
+
+The aim of this paper is to present a contemporary approach for SW and KG utilization in development applications.
 

### 2021-11-23 paper-aicts-cherkashin.tex
Added new page to paper.

@@
 % \lstset{breaklines=true}
+
+\newcommand{\ns}[1]{\textbf{\texttt{#1}}}
 
@@
 A number of tools for automation of conceptual model design and model conversion were proposed as well \cite{}.
-The R\& D at present is carried on in developing applications, improving and standardizing vocabularies, devising tools for solving particular problems.
+The R\&D at present is carried on in developing applications, improving and standardizing vocabularies, devising tools for solving particular problems.
 
@@
 
-[....
-About itch of personal vocabularies development instead to utilize similar ones ..
-Problem of searching resources ....
-metadata storages ...
-a general vocabularies ...
-Term ontology = vocabulary ]
+Domain specialists sometimes prefer develop local vocabularies instead of spending time for the routine work of Internet surfing for existing conceptual models and their enrichment.
+This is due to underdevelopment of the metadata resources (meta vocabularies) due to lack a domain classification system as compared, \emph{e.g.} to Universal Decimal Classification (UDC) used in bibliography.
+Vocabularies also can have common parts, describing the same or similar entities with different term or ways, more adopted to a specific problem solving.
 
-[we have by our opiinion a good examples of SW and KG applications in the following fields ....
-]  [ ...
-The KG and SW tech are well integrated with existing knowledge based AI systems [ if not use the equivalence]].
+Unitizing standardized vocabularies as domain conceptual models (ontologies) in our projects allow us to compare other research groups point of view to a problem with a formalized context, figure out underestimated and not fully formularized aspects of a domain under investigation/automation.
+This is the similar information source as scientific papers, but having concrete formal presentation of the information objects, allowing one to get rid of particular uncertainties.
+Another advantage of standardization is construction of a common ground of research, development, data formalization and interchange formats and semantics.
+This include aggregation of results of pattern recognition, applications of machine learning with their following use as facts in knowledge-based systems.
 
-The aim of this paper is to present a contemporary approach for SW and KG utilization in development applications.
+The aim of this paper is to present a contemporary approach for SW and KG utilization in development applications in context of improving software developers and researchers performance.
+A number of examples were reviewer showing advantageous points of the technologies application.
+
+\section{Contemporary SW tools review}
+\label{sec:sw-tools}
+
+Semantic Web technologies from the developer point of view include the following assets:
+\begin{enumerate}
+\item Global identification of the entities via URI\footnote{Abbreviation of Universal Resource Identifiers}, IRI\footnote{Abbreviation of International Resource Identifier}, URL\footnote{Abbreviation of Universal Resource Locator}, which are essentially the same notions but reflecting different representation properties;
+\item Conceptual set-theoretic and category-theoretic data representation as relation between resources and literals represented with triples of a form \texttt{<subject, predicate, object>};
+\item Triples used in a description of a domain form a corresponding sematic graph, conditionally subdivided into terminological subgraph (T-Box) and instance subgraph (A-Box);
+\item Graph presentation formats and techniques as files, databases, markups of documents, a data flow between applicants; \label{kg1}
+\item Public vocabularies, mostly A-Boxes (databases), for example \texttt{DBPedia.org}, a formalized version of \texttt{Wikipedia.org}. % wikidata
+\item For graphs being databases a query services are implemented to allow users obtain its data by portions, as well as implement modifications;
+\item API and libraries for processing files and documents with semantic data acquisition (\emph{e.g.} by extraction of the markup).
+\item Software for graph data processing;
+\item Validation and verification services allowing one to explore correctness and consistency of the technologies use and domain conceptual models. \label{kg2}
+\end{enumerate}
+The assets \ref{kg1}--\ref{kg2} form knowledge graph technological aspect of the SW that gained in the last decade special R\&D directions.
+
+Each entity being described in SW has its unique IRI, which is its virtual or real internet address locating it, \emph{e.g.} in a document most usually in a context of other resource.
+Using IRIs in applications solves fundamental problem of cross-domain and cross-application object identification and forming a basis of common data processing environment.
+Entities of various abstract levels can be designated an IRI, including vocabulary definitions (\emph{namespaces}), data types, data containers (files, databases).
+
+Triples are combination of two or three IRIs.
+Subject and predicate are always IRIs.
+Objects are either IRI or a literal.
+Literals are values of basic types (numbers, strings, dates, \emph{etc.}), usually accompanied with metadata (type designator and/or a language tag).
+Types are defined in a vocabulary \texttt{<\url{http://www.w3.org/2001/XMLSchema}\#>}, having well-known namespace abbreviation \ns{xs}\footnote{Hereafter, all well-known namespace names will be given after its IRI in brackets with bold monospace font.}.
+Language tag is used to define string values taking into account its affiliation to a cultural context.
+It is recognized with KG tools and utilized at presentation layers of the software.
+
+Namespaces are convenient instruments of resource definition.
+For example \verb|integer| literal \verb|42^^<http://www.w3.org/2001/XMLSchema#integer>| can be reduced to \verb|42^^xs:integer|, where \verb|xs| is namespace prefix of term \verb|integer|.
+Most of KG representation formats allow to define a default namespace, which terms are used without prefix.
+Almost any conceptual model of a domain represented as KG comprises entities taken from multiple vocabularies, and namespaces allow to specify the origin of a term.
+Three-component tuple (triple) being outside its KG context sometimes added the fourth component, its KG IRI (quadruple).
+
+Much R\&D efforts since 2001 \cite{tbl} were spend to develop SW data representation formats and their coupling with document representation.
+The simplest way to provide a printed document with semantic representation is pointing user of software agent to an URL serving the data.
+But this approach requires to implement two different procedures for document generation and its metadata.
+That's why a RDFa\footnote{Abbreviation of Resource Description Framework attribute language} has been developed.
+It allows developer of XHTML\footnote{HTML structures with strict XML restrictions obeyed} documents to add semantic layer inside the document.
+The semantic markup is represented with special attributes and their values.
+The hierarchical structure of XHTML is a matrix of hierarchical structure of document entities semantic relations.
+The layer is extracted by XHTML/RDFa processing libraries.
+There are other formats having the same expression capabilities as RDF but used for other purposes: \texttt{N-triples}, \texttt{Turtle}, \texttt{Notation~3}, \texttt{JSON-LD}.
+The last one is an efficient way of coupling SW and various generic data storage, indexing and processing services.
+
+Among published global vocabularies, \texttt{DBPedia} was of most valuable for our projects.
+It is used for concrete object or notion of physical world reference, \emph{e.g.} ``passport'' in law documents authoring \cite{}, naming (labeling) referenced objects in user interface with corresponding words in natural language \cite{github-db}.
+Another useful resource is \texttt{Wordnet} converted in RDF triples.
+It describes semantics of english words and various mutual relations.
+
+Access by program agents to the global resources are standardized as well.  Each serious vocabulary have API access point (endpoints) of form
+\verb|http[s]://<address>/sparql|\footnote{For example, see DBpedia access point at the URL \url{https://dbpedia.org/sparql}} supporting direct \texttt{GET}-queries returning a user form, and \texttt{SPARQL} \texttt{POST} query with answering \texttt{JSON} encoded tabular result.
+The form is used to interactively test \texttt{SPARQL} queries before their incorporation in software agents.
+All \texttt{SPARQL} endpoints support federated queries, \emph{i.e.} run subqueries on another servers.
+For a personal user KG storage, a number of servers were developed, \emph{e.g.}, \texttt{GraphDB}, \texttt{Jena Fuseki}, \texttt{ClioPatria}.
+Each has its extension engines.
+\texttt{ClioPatria} extends with Prolog modules and allows one to implement KG processing on server side.
+
+KGs and their subgraphs are programmatically processed using libraries developed for all papular programming environments.
+KG is represented with triple or quadruple set object.
+The set is traversed, triples are filtered according to various criteria.
+Filters are defined usually as patterns, where each of the components either a free variable/placeholder or a concrete value (IRI, literal), which must mach exactly.
+Some libraries allow one execute \texttt{SPARQL} queries locally.
+KG are evolved with adding new triples, as well as removing irrelevant ones.
+The resulting modified KG is stored as files and published on a server.
+An example of a federated SPARQL query is presented in Figure~\ref{fig:sparql-ex1}, which results in counting parts classes as the first level of a three widget.
+Each class is named in Russian if \texttt{DBPedia} contains a corresponding label.
+Otherwise label is figured out of IRI.
+This query is running at a local server and initiates a subquery to \texttt{DBPedia.org} endpoint.
+
+\begin{figure}[bth]
+  \centering
+\begin{minted}{sparql}
+PREFIX dbr: <http://dbpedia.org/resource/>
+PREFIX owl: <http://www.w3.org/2002/07/owl#>
+PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
+select * where {
+  ?class rdfs:subClassOf <http://dbpedia.org/resource/Electronic_component> .
+  bind(replace(strafter(str(?class),str(dbr:)),'_',' ') as ?class_name) .
+  OPTIONAL {
+    SERVICE <http://dbpedia.org/sparql> {
+      ?class rdfs:label ?title .
+      FILTER (
+        langMatches(lang(?title), 'ru')
+      )
+    }
+  }
+} limit 100
+\end{minted}
+  \caption{Example of a federated SPARQL query usage in a electronic component database}
+  \label{fig:sparql-ex1}
+\end{figure}
+
+Formalized representation of a domain, which is provided by SW techniques, is the basis of automatic  inference construction, thus modeling, \emph{e.g.}, problem-solving, resulting decision synthesis, theorem proving, domain verification.
+The inference is implemented in various ways, all of them belong either to semantic inference, usually implemented as solving a corresponding CSP\footnote{Abbreviation of Constraint Satisfaction Problem}, or syntax-driven procedures that mediate the previous structures and infer one of special properties (Automatic Theorem Proving).
+Theoretic research is being carried on classification the properties of vocabularies by complexity of verification and development software implementing it for these classes.
+These software are parts of RDF-processing libraries and stand-alone system extensions.
+
+Our interest is located in inferring new objects using logic programming from domain models represented as KGs.
+It allows developer to combine KG processing with other inferences in the same programming paradigm.
+SWI-Prolog programming language environment and Logtalk extension is used for the purpose.
+SWI system has library for representing and processing KG, and Logtalk is used to realize transformations providing the inference.
+Object-oriented properties of Logtalk enable one to cope with generic disadvantaged of RDF\footnote{Requirements of triple tuple data representation.} with encapsulation, manipulate knowledge sets with extension, inheritance and composition, and represent inference in a generalized form of objects sending messages to each other.
+
+% TODO: Integration out of our architecture.
+
+\section{General architecture of applications based on KG}
+\label{sec:architecture}
+
+Some general words about architectures .... related to WEB.
+
+In order to present the general architecture of application development, we have to briefly speak about KG definitions and properties.
+
+\subsection{Knowledge graph properties as application basis}
+\label{sec:kg-as-basis}
+
+KGs changed the view point to the knowledge base as being a part of whole totality of knowledge, implying the obeying the global standards and techniques of its acquisition and processing \cite{hogan}.
+The following outlines their definitions \cite{hogan}.
+ \begin{itemize}
+  \item Notions \emph{data} and \emph{knowledge} are converged into ``something is \emph{known}'';
+  \item KG contains, in general, data, relations, and metadata;
+  \item The same KG is being filled in and processed by different independent modules, \emph{e.g.}, withing SPARQL queries with UPDATE statements;
+  \item KG fundamental software development allow to \emph{postpone} the formal definition of a data schema;
+  \item There is three points of view to graph schemata: \emph{semantic}, which is aimed at generalization, \emph{validating}, reflecting semantic verification, checking completeness w.r.t.
+sets of predefined general relations, and \emph{emergent}, observed or deduced from exploitation experience, as a set of new generalized structures and \emph{refactoring} of the KG.
+  \end{itemize}
+
+\section{Discussion}
+\label{sec:disc}
+
+[we tried to collaborate with software developers using traditional approaches.  They are not interested as ]
 
@@
 metadata-extraction functionality. This document will explain the
-major features of the document class\footnote{You can use this
-  document as the template for preparing your publication. We
-  recommend using the latest version of the ceurart style.}.
+major features of the document class.
 
@@
 
+\bibitem{tbl} T.B.-L. Scientific American ...
+
 \bibitem{lod} C.~Bizer, T.~Heath, T.~Berners-Lee, Linked Data -- The Story So Far. Int. J. Semantic Web Inf. Syst., 5 (2009), pp.~1--22. \doi{10.4018/jswis.2009081901}

### 2021-11-23 paper-aicts-cherkashin.tex
Added architecture figure.

@@
 
-Some general words about architectures .... related to WEB.
+Some general words about architectures .... related to WEB.  Modeled with component architecture.
 
@@
 
-\subsection{Knowledge graph properties as application basis}
-\label{sec:kg-as-basis}
+% \subsection{Knowledge graph properties as application basis}
+% \label{sec:kg-as-basis}
 
@@
   \end{itemize}
+
+\begin{figure}[bth]
+  \begin{center}
+    \includegraphics[width=0.9\linewidth]{architecture-mda-lod-ext-general.pdf}
+  \end{center}
+  \begin{center}
+      \textbf{Abbreviations}\\[1ex]\scriptsize
+      T-Module is Transformation module,
+      MDA is Model-Driven Architecture,
+      CIM is Computationally Independent Model,
+      PIM is Platform Independent Model,
+      PSM is Platform Specific Model,
+      T-Box is Terminological Box,
+      A-Box is Instance Box,
+      A-Box is Instance Box,
+      NGS is Next-Generation Sequencing,
+      DB is Database.
+  \end{center}
+
+  \caption{Architecture of services}
+  \label{fig:arch-services}
+\end{figure}
 

### 2021-12-05 paper-aicts-cherkashin.tex
Presentation should be ok now

@@
   Semantic Web technologies and standards are becoming one of productive assets for software development and integration on various levels of design and implementation.
-This include OSI level 6 data representation, database and file storage formats, the source data for indexes, hypertext markup, source data for document authoring, representing abstract software models, \emph{etc}.
+These include OSI level 6 data representation, database and file storage formats, the source data for indexes, hypertext markup, source data for document authoring, representing abstract software models, \emph{etc}.
 They also smoothly combined with development and data analysis tools, giving raise a new common-ground environment for realizing integration on the semantic level.
@@
   \end{itemize}
-
+With these general constraints, design of application modules are to be built on standardized data interchange formats, which are to be easily converted from and to KG representation, as well as applications are to be constructed out of service modules supporting SW endpoints (SPARQL) whenever possible.
+Our applications are composition of modules showed in Figure~\ref{fig:arch-services}.
 \begin{figure}[bth]
@@
 \end{figure}
+
+The kernel module of applications is \texttt{KG server} (A2) built on the base of ClioPatria \cite{cliopatria} with web-server and Pengines plug-ins.
+ClioPatria is a web-application implemented in SWI-Prolog \cite{swi} being ISO-compliant Prolog implementation.
+KG functions are realized with SW package, supplied in standard SWI-Prolog distribution.
+Pengines service enables SW web-applications requesting logical inference on the server side \cite{pengines} with integration of browser JavaScript environment for supporting input-output.
+Pengines allows integration of Prolog programs as its plugins.
+Kernel module stores A-~and T-boxes of applications, \emph{i.e.} it is application database.
+Other modules support application (C3) with frequently used services.
+
+Authoring tool (B1) is used to provide document generation and integration \cite{sibiricon}.
+This tool enables document content and structure inference in browser utilizing templates and data from other web pages and websites.
+Programmer implements a view of an object as a web page with an extended version of RDFa.
+The page are generated with any programming system, \emph{e.g.} PHP.
+Authoring tool engine collects web page chunks and composing the document out of these chunks as a XHTML page.
+This composed page can be stored in KG as set of triples and indexed in Text indexing module (C1), printed with browser's printing engine, or converted to PDF.
+The tool allow user to change document content in WISYWIG mode in some extent.
+Its general structure is defined by changing template at server side.
+
+Text indexing engine (C1) provides service for storing documents represented as KG triples, JSONs, and BLOBS of any format with corresponding text layers.
+This module has two implementations.
+One is build on Sphinx Search engine, and the second one is Elasticsearch.
+Elasticsearch supports JSON as the only document format and easily integrated with KG using JSON-LD for triple representation.
+Sphinx Search is much faster in text and database records indexing and consumes less RAM as implemented in C++.
+The choice of an engine depends on the primary document representation to be used in application.
+
+BLOB stored documents having no markup usually contain useful data, which are to be revealed for application use.
+They could be report data or scientific paper content represented as PDF, DJVU files or raster images.
+Data recognition is realized in PDF/Text document processing module (C2).
+Data processing workers query Text indexing module for BLOB documents and add recognized text and high-order structure layers.
+All obtained layers are stored in indexing database of C1, as well as data of interest can be converted in triples and sent to KG.
+
+Processing KG for the various purposed are presented as KB processing module (B3).
 
@@
 \end{itemize}
-
-\section{Front matter}
 
@@
 
+\bibitem{cliopatria}
+\bibitem{sibircon} Logical objects
+
 \end{thebibliography}

### 2021-12-05 talk-aicts-2021-12-06.tex
Added new page to paper.
Presentation should be ok now

@@
 \author[E.~Cherkashin]{\bfseries%
-  Evgeny Cherkashin}
+  Evgeny Cherkashin, Nikita Dorodnykh, Alexey Shigarov, \emph{et al.}}
 \institute{\normalsize Matrosov Institute for System Dynamics and Control Theory of Siberian Branch of Russian Academy of Sciences, Irkutsk, Russia\\%
@@
   \begin{enumerate}
+  \item Application data representation
   \item Ontological model representation
-  \item Application data representation
+  \item Document publication
+  \item Application integration % via knowledge graph data documents
   \item Model transformation
-  \item Document publication
-  \item Application integration via knowledge graph data documents
   \end{enumerate}
@@
 <rdf:RDF
-    xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
+    xmlns:rdf="http://www.w3.org/1999/02/
+               22-rdf-syntax-ns#"
     xmlns:dc="http://purl.org/dc/elements/1.1/">
-  <rdf:Description rdf:about="http://en.wikipedia.org/wiki/Tony_Benn">
+  <rdf:Description rdf:about="http://en.wikipedia.org/
+                              wiki/Tony_Benn">
     <dc:title>Tony Benn</dc:title>
@@
     "homepage": {
-      "@id": "http://xmlns.com/foaf/0.1/workplaceHomepage",
+      "@id": "http://xmlns.com/foaf/0.1/
+              workplaceHomepage",
       "@type": "@id"
@@
   \end{center}
+  Using standard vocabularies form cross-application platform, \emph{e.g.}, interpreting relations.
 \end{frame}
@@
   \item Contain data, relations, and metadata (vocabularies)
-  \item Distinguished \textbf{node filling in} and \textbf{processing} graph triples, \emph{e.g.}, withing SPARQL queries with UPDATEs
+  \item Distinguished \textbf{node filling in} and \textbf{processing} graph triples, \emph{e.g.}, with SPARQL queries with UPDATEs
   \item Allow \textbf{postpone} the formal definition of a schema
@@
 \begin{frame}
-  \frametitle{Information infrastructure for supporting Baikal microbiome research}
+  \frametitle{Useful standard vocabularies}
+
+  Standardized vocabularies
+
+  \begin{itemize}
+  \item Friend-of-a-friend (\textbf{foaf}) for agent information: individuals, legal entities, program agents.
+  \item Provenance (\textbf{prov}) for making references between documents.
+  \item Dublin Core (\textbf{dc}) for published resource metadata mark up.
+  \item DBPedia resource (\textbf{dbr}) to refer external classes and instance objects.
+  \item Open annotation (\textbf{oa}) as an ``bookmark'' ontology.
+  \item The Bibliographic Ontology (\textbf{bibo}) used for literature reference mark up.
+  \item Schema.org (\textbf{schema}) for Google, Yandex, Yahoo, \emph{etc}. searchable objects, structural elements.
+  \end{itemize}
+
+  Non-standard vocabularies
+
+  \begin{itemize}
+  \item Ontology \textbf{nssp} for Mothur source code processing results.
+  \item Ontology \textbf{uml} for XMI representation.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}[fragile]
+  \frametitle{Instrumentation: Ontology metadata server LOV}
+  \begin{center}
+    \includegraphics[width=0.8\linewidth]{lov-main.png}
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Applications}
+  \vfil
+  \begin{center}
+    \Large Applications
+  \end{center}
+  \vfil
+\end{frame}
+
+\begin{frame}
+  \frametitle{Application: Information infrastructure for supporting Baikal microbiome research}
   \centering
@@
 
-\begin{frame}
-  \frametitle{Microbiome study aims}
-  \centering
-  \includegraphics[width=\linewidth]{microbiome-study.png}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Microbiome study process}
-  \centering
-  \includegraphics[width=\linewidth]{microbiome-study2.png}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Microbiome study aims}
+%   \centering
+%   \includegraphics[width=\linewidth]{microbiome-study.png}
+% \end{frame}
+
+% \begin{frame}
+%   \frametitle{Microbiome study process}
+%   \centering
+%   \includegraphics[width=\linewidth]{microbiome-study2.png}
+% \end{frame}
 
@@
 
-\begin{frame}
-  \frametitle{The process of data analysis (MiSeq SOP)}
-  \begin{enumerate}
-  \item Reconstruct \emph{cotigs} (contiguous gene parts) from ``left'' and ``right'' \emph{readings}.
-  \item Trim \textbf{bar-code} and other \emph{primers}.
-  \item Filter sequences according to formal criteria (ambiguity, average length, maximal length of homopolymer).
-  \item Classify \emph{unique} sequences and count their appearance in groups (samples).
-  \item Alignment with reference sequences from SILVA database.
-  \item Filter non-hanging sequences.
-  \item Filter chimeras, fund unique sequences again.
-  \item Classify sequences with respect to existing taxa hierarchy. Get \textbf{OTU}s.
-  \end{enumerate}
-
-  After these stages a large number of OTU\footnote{Operation Taxonomic Unit} classified has been obtained.
-\end{frame}
+% \begin{frame}
+%   \frametitle{The process of data analysis (MiSeq SOP)}
+%   \begin{enumerate}
+%   \item Reconstruct \emph{cotigs} (contiguous gene parts) from ``left'' and ``right'' \emph{readings}.
+%   \item Trim \textbf{bar-code} and other \emph{primers}.
+%   \item Filter sequences according to formal criteria (ambiguity, average length, maximal length of homopolymer).
+%   \item Classify \emph{unique} sequences and count their appearance in groups (samples).
+%   \item Alignment with reference sequences from SILVA database.
+%   \item Filter non-hanging sequences.
+%   \item Filter chimeras, fund unique sequences again.
+%   \item Classify sequences with respect to existing taxa hierarchy. Get \textbf{OTU}s.
+%   \end{enumerate}
+
+%   After these stages a large number of OTU\footnote{Operation Taxonomic Unit} classified has been obtained.
+% \end{frame}
 
@@
   private InputPort fastaInPort = getInputPorts().createPort("fasta");
-  private InputPort referenceInPort = getInputPorts().createPort("reference");
+  private InputPort referenceInPort = getInputPorts().
+                                         createPort("reference");
   private OutputPort chimeraOutPort = getOutputPorts().createPort("chimera");
@@
              gal:value "0" ;
-             dc:description "pdiffs - number of differences to allow in the primer (default 0)" ;
+             dc:description "pdiffs - number of differences . . . . . ." ;
              dc:title "pdiffs" ;
@@
                gal:value "0" ;
-               dc:description "bdiffs - number of differences to allow in the barcode (default 0)" ;
+               dc:description "bdiffs - number of differences . . . . . . " ;
                dc:title "bdiffs" ;
@@
                gal:value "0" ;
-               dc:description "tdiffs - total number of differences to allow in primer and barcode (default 0)" ;
+               dc:description "tdiffs - total number of diffe... . . . . ." ;
              dc:title "tdiffs" ;
@@
 \begin{frame}
-  \frametitle{Architecture of transformation modules}
-  \centering
-  \includegraphics[width=0.9\linewidth]{architect_tree_pres-en-wo-OCL.pdf}
-\end{frame}
-
-\begin{frame}
   \frametitle{Model Driven Architecture and Linked Open Data}
@@
 
-\begin{frame}[fragile]\frametitle{Architecture of services}
+\begin{frame}[fragile]\frametitle{Architecture of services for NGS}
   \begin{columns}
     \begin{column}{0.7\linewidth}
-      \includegraphics[width=1\linewidth]{architecture-mda-lod-ext-galaxy.pdf}
+      \includegraphics[width=1\linewidth]{architecture-mda-lod-ext-general.pdf}
     \end{column}
@@
 \begin{frame}
-  \frametitle{Used ontologies}
-
-  Standardized ontologies
-
-  \begin{itemize}
-  \item Friend-of-a-friend (\textbf{foaf}) for agent information: individuals, legal entities, program agents.
-  \item Provenance (\textbf{prov}) for making references between documents.
-  \item Dublin Core (\textbf{dc}) for published resource metadata mark up.
-  \item DBPedia resource (\textbf{dbr}) to refer external classes and instance objects.
-  \item Schema.org (\textbf{schema}) for Google, Yandex, Yahoo, \emph{etc}. searchable objects, structural elements.
-  \item The Bibliographic Ontology (\textbf{bibo}) used for literature reference mark up.
-  \item Open annotation (\textbf{oa}) as an ``bookmark'' ontology.
-  \end{itemize}
-
-  Non-standard ontologies
-
-  \begin{itemize}
-  \item Ontology \textbf{nssp} for Mothur source code processing results.
-  \item Ontology \textbf{uml} for XMI representation.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}[fragile]
-  \frametitle{Instrumentation: Ontology metadata server LOV}
-  \begin{center}
-    \includegraphics[width=0.8\linewidth]{lov-main.png}
-  \end{center}
+  \frametitle{Architecture of transformation modules}
+  \centering
+  \includegraphics[width=0.9\linewidth]{architect_tree_pres-en-wo-OCL.pdf}
 \end{frame}
@@
 
-\begin{frame}
-  \frametitle{Discussion (MDA application)}
-  Interesting positive impressions obtained:
-  \begin{itemize}
-  \item Logtalk and RDF are flexible, sufficiently universal and convenient implementation infrastructures for MDA;
-  \item The best implemenation means is Prolog predicate wrapping and Logtalk object encapsulation of rules;
-  \item Not all Logtalk properties are investigated: there might be more sophisticated programming techniques developed, \emph{e.g.}, on the base of message watchers.
-  \end{itemize}
-  Technical problems making the approach somewhat problematic:
-  \begin{itemize}
-  \item Very simple tasks take too much efforts, \emph{e.g.}, text processing: convert an identifier into the CamelCase;
-  \item It takes too long to surf Internet in order to find a vocabulary for a domain, but it is more productive than development new one and classes;
-  \item Prolog is not a popular language in MDA, neither Logtalk.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Future activities}
-  The future activities supposed to be as follows:
-  \begin{enumerate}
-  \item Having dataflow models of MiSeq SOP and other techniques, device an intelligent subsystem, which will construct computational procedures for a predefined set of data processing tasks (\emph{AI's problem solving}).
-  \item Integrate dataflow and data storage with Galaxy.
-    \begin{itemize}
-       \item Realize adapters of data/metadata storage and retrieving, as well as the storage.
-       \item Create a more sophisticated source code parser or PIM model of computation so we will able to infer metadata for the output on the metadata of input (partially done in November).
-         % \item Adapt our ``pattern-directed'' approach of semantically marked up document authoring to present input and output to the scientific communities.
-       \item Adapt our document authoring tools to Galaxy allowing LOD representation of results.
-  \end{itemize}
-  \item Implement integration to biological/gene databases.
-  \item Write a handbook on Logtalk programming strategies with its author Paulo Mora.
-  \end{enumerate}
-\end{frame}
+% \begin{frame}
+%   \frametitle{Future activities}
+%   The future activities supposed to be as follows:
+%   \begin{enumerate}
+%   \item Having dataflow models of MiSeq SOP and other techniques, device an intelligent subsystem, which will construct computational procedures for a predefined set of data processing tasks (\emph{AI's problem solving}).
+%   \item Integrate dataflow and data storage with Galaxy.
+%     \begin{itemize}
+%        \item Realize adapters of data/metadata storage and retrieving, as well as the storage.
+%        \item Create a more sophisticated source code parser or PIM model of computation so we will able to infer metadata for the output on the metadata of input (partially done in November).
+%          % \item Adapt our ``pattern-directed'' approach of semantically marked up document authoring to present input and output to the scientific communities.
+%        \item Adapt our document authoring tools to Galaxy allowing LOD representation of results.
+%   \end{itemize}
+%   \item Implement integration to biological/gene databases.
+%   \item Write a handbook on Logtalk programming strategies with its author Paulo Mora.
+%   \end{enumerate}
+% \end{frame}
 
@@
 \begin{frame}
-  \frametitle{Document authoring and storage}
+  \frametitle{Application: Document authoring and storage}
   In most cases documents are created as a result of
@@
 \begin{frame}
-\frametitle{Cartographical WEB-service with knowledge graph of South-Siberian faults}
+\frametitle{Application: Cartographical WEB-service with knowledge graph of South-Siberian faults}
 \textbf{Aim} is to construct a WEB-GIS browser for faults stored in the KG.
@@
 
-% <<Загрузка>> онтологий
+% Interpret any well described relation
+% between a subject (fault) via ''of''.
 schema(Prop, 'of', SubjName):-     % used on translation stage
@@
 
-\begin{frame}
-  \frametitle{Conclusion}
-  The following results have been obtained as for today:
-  \begin{itemize}
-  \item A technique for \textbf{domain} model representation has been developed and tested.
-  \item A programming technique using object-oriented logical language Logtalk is devised.
-  \item Prototypes of various transformation procedures are implemented.
-  \item Transformation tools are tested in application areas and no significant technical problems were mentioned.
-  \end{itemize}
-  Further development directions are as follows:
-  \begin{itemize}
-  \item A technique for document automatic markup with vocabulary entities.
-  \item A transformation implemenation techniques, minimizing usage of dynamic objects, targeting on macro properties of Logtalk.
-  \item Form a toolset out of existing prototypes obeying nowadays software development requirements.
-  \end{itemize}
-  The source codes are available at \url{https://github.com/isu-enterprise/icc.xmitransform}, \url{https://github.com/eugeneai/icc.mothurpim}.
-\end{frame}
-
-\begin{frame}
-  \frametitle{Technologies used (open source)}
-  Python-3.x.x (\url{http://python.org}),\\
-  ZCA (\url{https://muthukadan.net/docs/zca.html}),\\
-  SWIG (\url{http://swig.org/}),\\
-  SWI-Prolog (\url{https://www.swi-prolog.org/}),\\
-  Logtalk (\url{https://logtalk.org/}),\\
-  ClioPatria (\url{https://cliopatria.swi-prolog.org/home}),\\
-  Pengines (\url{https://pengines.swi-prolog.org/docs/index.html}),\\
-  LOV (\url{https://lov.linkeddata.es/dataset/lov/}),\\
-  Elastic Search (\url{https://www.elastic.co/}),\\
-  Kyotocabinet (\url{https://fallabs.com/kyotocabinet/}),\\
-  DBPedia (\url{https://wiki.dbpedia.org/}),\\
-  Mothur (\url{https://mothur.org/}),\\
-  Galaxy (\url{https://usegalaxy.org/}),\\
-  R (\url{https://www.r-project.org/}), \\
-  Dust.js (\url{https://akdubya.github.io/dustjs/})
-\end{frame}
+% \begin{frame}
+%   \frametitle{Conclusion}  % On Web GIS
+%   The following results have been obtained as for today:
+%   \begin{itemize}
+%   \item A technique for \textbf{domain} model representation has been developed and tested.
+%   \item A programming technique using object-oriented logical language Logtalk is devised.
+%   \item Prototypes of various transformation procedures are implemented.
+%   \item Transformation tools are tested in application areas and no significant technical problems were mentioned.
+%   \end{itemize}
+%   Further development directions are as follows:
+%   \begin{itemize}
+%   \item A technique for document automatic markup with vocabulary entities.
+%   \item A transformation implemenation techniques, minimizing usage of dynamic objects, targeting on macro properties of Logtalk.
+%   \item Form a toolset out of existing prototypes obeying nowadays software development requirements.
+%   \end{itemize}
+%   Projects sources are at \url{https://github.com/isu-enterprise} and \url{https://github.com/eugeneai}.
+% \end{frame}
+
+\begin{frame}
+  \frametitle{Conclusion}  % On Web GIS
+  Web-3.0 (SW and KG) are convenient and productive basis of application development, allowing one to
+  \begin{itemize}
+  \item Integrate on various levels of software representation, \emph{i.e.}, from data to abstract models.
+  \item There are assets to utilize global resources in applications (\texttt{DBPedia.org}).
+  \item The inherited software are involved via adapters.
+  \end{itemize}
+  In our projects we has been developed techniques to
+  \begin{itemize}
+  \item Data processing and model transformation using Prolog-based logical inference.
+  \item Logtalk is a perfect instrument for knowledge encapsulation and processing, complex data synthesis.
+  \end{itemize}
+\end{frame}
+
+% \begin{frame}
+%   \frametitle{Technologies used (open source)}
+%   Python-3.x.x (\url{http://python.org}),\\
+%   ZCA (\url{https://muthukadan.net/docs/zca.html}),\\
+%   SWIG (\url{http://swig.org/}),\\
+%   SWI-Prolog (\url{https://www.swi-prolog.org/}),\\
+%   Logtalk (\url{https://logtalk.org/}),\\
+%   ClioPatria (\url{https://cliopatria.swi-prolog.org/home}),\\
+%   Pengines (\url{https://pengines.swi-prolog.org/docs/index.html}),\\
+%   LOV (\url{https://lov.linkeddata.es/dataset/lov/}),\\
+%   Elastic Search (\url{https://www.elastic.co/}),\\
+%   Kyotocabinet (\url{https://fallabs.com/kyotocabinet/}),\\
+%   DBPedia (\url{https://wiki.dbpedia.org/}),\\
+%   Mothur (\url{https://mothur.org/}),\\
+%   Galaxy (\url{https://usegalaxy.org/}),\\
+%   R (\url{https://www.r-project.org/}), \\
+%   Dust.js (\url{https://akdubya.github.io/dustjs/})
+% \end{frame}
 

### 2022-02-01 paper-aicts-cherkashin.tex
Some grammarly made gram checking

--- 2021-11-23 paper-aicts-cherkashin.tex
@@
     pdfauthor={Evgeny Cherkashin},     % author
-    pdfsubject={model driven architecture},   % subject of the document
+    pdfsubject={Technologies of Semantic WEB as an environment of application development and integration},   % subject of the document
     pdfnewwindow=true,      % links in new PDF window
@@
 %%
-%% The abstract is a short summary of the work to be presented in the
+%% The abstract is a summary of the work to be presented in the
 %% article.
 \begin{abstract}
-  Semantic Web technologies and standards are becoming one of productive assets for software development and integration on various levels of design and implementation.
-This include OSI level 6 data representation, database and file storage formats, the source data for indexes, hypertext markup, source data for document authoring, representing abstract software models, \emph{etc}.
-They also smoothly combined with development and data analysis tools, giving raise a new common-ground environment for realizing integration on the semantic level.
+  Semantic Web technologies and their standards are becoming productive assets for software development and integration on various design and implementation levels.
+These include OSI level 6 data representation, database and file storage formats, the source data for text indexes, document authoring, hypertext markup engines, a representation for abstract software models, \emph{etc}.
+They are also smoothly combined with development and data analysis tools, giving raise new common-ground environment for realizing an integration of distributed applications on a semantic level.
 
-  This paper presents the view on the Semantic Web technologies as data representation for software development tools with model analysis, transformation and retrospection capabilities.
-Three examples are presented, exposing our experience.
+  This paper presents a view on the Semantic Web technologies as data representation for software development tools with model analysis, transformation, and retrospection capabilities.
+Three examples show the technologies usage, exposing our experience.
 \end{abstract}
@@
 
-Since the beginning of 2000s the development of the Semantic Web technologies became [awesome] mostly in describing software domains, representing semantics withing data objects, and define first-order knowledge \cite{}.
-A number of tools for automation of conceptual model design and model conversion were proposed as well \cite{}.
-The R\&D at present is carried on in developing applications, improving and standardizing vocabularies, devising tools for solving particular problems.
+Since 2000 the development of the Semantic Web (SW) technologies became actively used usually in describing software domains, representing semantic relations between objects.
+For some domains on the base of the SW description, first-order knowledge theories are constructed \cite{}, but in practical aspect of the SW utilization, the solving data representation and software integration aspects prevail.
+Some tools for automation of conceptual model design and model conversion were proposed \cite{}.
+The R\&D at present are carried on in developing applications, improving and standardizing vocabularies (conceptual domain models), devising tools for solving particular problems.
 
 Unification of routine tasks resulted in development common techniques for storing and processing data, which are aggregated around term ``knowledge graph'' (KG).
-These include data representation formats, vocabulary standards, query languages, tools for accessing and management KG.
+These include data representation formats, vocabulary standards, query languages, tools for accessing and management data in a KG.
 Software development techniques are mediated with properties of the new-created resources and their APIs\footnote{Abbreviation of Application Programming Interface}.
-Many internet resources provide common ways of formalized domain data access using query endpoints, and these endpoints allow one to organize federated data access.
+Many internet resources provide common ways of formalized domain data access using query endpoints, and these endpoints allow one to organize federated KG data access.
 
-Domain specialists sometimes prefer develop local vocabularies instead of spending time for the routine work of Internet surfing for existing conceptual models and their enrichment.
-This is due to underdevelopment of the metadata resources (meta vocabularies) due to lack a domain classification system as compared, \emph{e.g.} to Universal Decimal Classification (UDC) used in bibliography.
-Vocabularies also can have common parts, describing the same or similar entities with different term or ways, more adopted to a specific problem solving.
+In Russia, domain model designers generally prefer to develop local vocabularies instead of spending time for the routine work of Internet surfing for existing conceptual models and their enrichment.
+It is due to underdevelopment of the metadata resources (metavocabularies) due to lack a domain classification system as compared, \emph{e.g.} to Universal Decimal Classification (UDC) used in bibliography.
+Vocabularies also can have common parts, describing the same or similar entities with different term and relations, \emph{e.g.} semantically more adopted to a specific problem solving.
 
-Unitizing standardized vocabularies as domain conceptual models (ontologies) in our projects allow us to compare other research groups point of view to a problem with a formalized context, figure out underestimated and not fully formularized aspects of a domain under investigation/automation.
-This is the similar information source as scientific papers, but having concrete formal presentation of the information objects, allowing one to get rid of particular uncertainties.
-Another advantage of standardization is construction of a common ground of research, development, data formalization and interchange formats and semantics.
-This include aggregation of results of pattern recognition, applications of machine learning with their following use as facts in knowledge-based systems.
+Unitizing standardized vocabularies as domain conceptual models (ontologies) in our projects allow us to compare other research groups point of view to a problem with a formalized context, figure out underestimated and not fully formularized aspects of a domain under investigation or automation.
+The vocabularies are similar to information sources as scientific papers, but have a concrete formal presentation of the information objects, allowing one to get rid of particular uncertainties.
+Another advantage of standardization is a construction of a common ground of research, development, data formalization, and interchange formats and semantics.
+It includes an aggregation of results of pattern recognition, applications of machine learning with their following use as facts in knowledge-based systems.
 
-The aim of this paper is to present a contemporary approach for SW and KG utilization in development applications in context of improving software developers and researchers performance.
-A number of examples were reviewer showing advantageous points of the technologies application.
+The aim of this paper is to present a contemporary experience of SW and KG utilization in the development of applications in a context of performance improvement of R\&D.
+Three examples were reviewed showing advantageous points of the technologies' application.
 
-\section{Contemporary SW tools review}
+\section{Semantic Web tools}
 \label{sec:sw-tools}
 
-Semantic Web technologies from the developer point of view include the following assets:
+Briefly, SW technologies from a developer point of view include the following assets:
 \begin{enumerate}
-\item Global identification of the entities via URI\footnote{Abbreviation of Universal Resource Identifiers}, IRI\footnote{Abbreviation of International Resource Identifier}, URL\footnote{Abbreviation of Universal Resource Locator}, which are essentially the same notions but reflecting different representation properties;
-\item Conceptual set-theoretic and category-theoretic data representation as relation between resources and literals represented with triples of a form \texttt{<subject, predicate, object>};
-\item Triples used in a description of a domain form a corresponding sematic graph, conditionally subdivided into terminological subgraph (T-Box) and instance subgraph (A-Box);
-\item Graph presentation formats and techniques as files, databases, markups of documents, a data flow between applicants; \label{kg1}
-\item Public vocabularies, mostly A-Boxes (databases), for example \texttt{DBPedia.org}, a formalized version of \texttt{Wikipedia.org}. % wikidata
+\item Global identification of the entities via URI\footnote{Abbreviation of Universal Resource Identifiers}, IRI\footnote{Abbreviation of International Resource Identifier}, URL\footnote{Abbreviation of Universal Resource Locator}, which are essentially the same notions but reflecting different representation properties of an identified resource;
+\item Conceptual set-theoretic and category-theoretic data representation notation as relations between resources and literals represented with triples of a form \texttt{<subject, predicate, object>};
+\item Triples used in a description of a domain form a corresponding semantic graph, conditionally subdivided into terminological subgraph (T-Box) and instance subgraph (A-Box);
+\item Graph presentation formats and techniques are implemented in files, databases, markups of documents, a data flow between applicants; \label{kg1}
+\item There are public semantic databases comprising the generic domain objects descriptions, which are mostly A-Boxes (databases), for example \texttt{DBPedia.org}, a formalized version of \texttt{Wikipedia.org}.
+% wikidata
 \item For graphs being databases a query services are implemented to allow users obtain its data by portions, as well as implement modifications;
 \item API and libraries for processing files and documents with semantic data acquisition (\emph{e.g.} by extraction of the markup).
-\item Software for graph data processing;
-\item Validation and verification services allowing one to explore correctness and consistency of the technologies use and domain conceptual models. \label{kg2}
+\item Software for graph data processing to improve its structure or induct general relation patterns;
+\item Validation and verification services allowing one to explore consistency and completeness of the conceptual domain models and their implementations in software. \label{kg2}
 \end{enumerate}
-The assets \ref{kg1}--\ref{kg2} form knowledge graph technological aspect of the SW that gained in the last decade special R\&D directions.
+The assets \ref{kg1}--\ref{kg2} form knowledge graph technological aspect of the SW that gained a special interest within R\&D in the last decade.
 
-Each entity being described in SW has its unique IRI, which is its virtual or real internet address locating it, \emph{e.g.} in a document most usually in a context of other resource.
+Each entity being described in SW has its unique global IRI, which is its virtual or real internet address.
+Document entities' resource IRIs are formed accounting the document structure, \emph{e.g.} a hierarchy of contexts.
 Using IRIs in applications solves fundamental problem of cross-domain and cross-application object identification and forming a basis of common data processing environment.
@@
 Subject and predicate are always IRIs.
-Objects are either IRI or a literal.
-Literals are values of basic types (numbers, strings, dates, \emph{etc.}), usually accompanied with metadata (type designator and/or a language tag).
+Objects are either IRI (a reference to a resource) or a literal.
+Literals are values of basic types (number, string, date, time, \emph{etc.}) usually accompanied with metadata (type designator and/or a language tag).
 Types are defined in a vocabulary \texttt{<\url{http://www.w3.org/2001/XMLSchema}\#>}, having well-known namespace abbreviation \ns{xs}\footnote{Hereafter, all well-known namespace names will be given after its IRI in brackets with bold monospace font.}.
-Language tag is used to define string values taking into account its affiliation to a cultural context.
+Language tag is used to define string values, taking into account its affiliation to a cultural context.
 It is recognized with KG tools and utilized at presentation layers of the software.
 
-Namespaces are convenient instruments of resource definition.
+Namespaces are convenient instruments in triple definition.
 For example \verb|integer| literal \verb|42^^<http://www.w3.org/2001/XMLSchema#integer>| can be reduced to \verb|42^^xs:integer|, where \verb|xs| is namespace prefix of term \verb|integer|.
-Most of KG representation formats allow to define a default namespace, which terms are used without prefix.
+Most of KG representation formats allow one to define a default namespace, which terms are used without prefix.
 Almost any conceptual model of a domain represented as KG comprises entities taken from multiple vocabularies, and namespaces allow to specify the origin of a term.
@@
 
-Much R\&D efforts since 2001 \cite{tbl} were spend to develop SW data representation formats and their coupling with document representation.
-The simplest way to provide a printed document with semantic representation is pointing user of software agent to an URL serving the data.
-But this approach requires to implement two different procedures for document generation and its metadata.
+Much R\&D efforts since 2001 \cite{tbl} were spend to develop SW data formats and their coupling with document representation.
+The simplest way to provide a printed document with semantic representation is pointing user's software agent to an URL serving the semantic data.
+But this approach requires to implement two different procedures for document generation and its metadata serving.
 That's why a RDFa\footnote{Abbreviation of Resource Description Framework attribute language} has been developed.
-It allows developer of XHTML\footnote{HTML structures with strict XML restrictions obeyed} documents to add semantic layer inside the document.
+It allows developer of XHTML\footnote{HTML structures with strict XML restrictions obeyed} documents to enrich the document with its semantic layer.
 The semantic markup is represented with special attributes and their values.
-The hierarchical structure of XHTML is a matrix of hierarchical structure of document entities semantic relations.
+The hierarchical structure of XHTML is a matrix layer for hierarchical structure of document entities semantic relations.
 The layer is extracted by XHTML/RDFa processing libraries.
@@
 
-Among published global vocabularies, \texttt{DBPedia} was of most valuable for our projects.
-It is used for concrete object or notion of physical world reference, \emph{e.g.} ``passport'' in law documents authoring \cite{}, naming (labeling) referenced objects in user interface with corresponding words in natural language \cite{github-db}.
-Another useful resource is \texttt{Wordnet} converted in RDF triples.
-It describes semantics of english words and various mutual relations.
+Among published global vocabularies, \texttt{DBPedia} is of most valuable for our projects.
+It is used for concrete object or notion of physical world reference, \emph{e.g.} the term ``passport'' in law documents authoring \cite{}, naming (labeling) referenced objects in user interface with corresponding words in natural language \cite{github-db}.
+Another useful resource is \texttt{WordNet} converted in RDF triples.
+It describes semantics of english words and various their mutual relations.
 
@@
 The form is used to interactively test \texttt{SPARQL} queries before their incorporation in software agents.
-All \texttt{SPARQL} endpoints support federated queries, \emph{i.e.} run subqueries on another servers.
+All \texttt{SPARQL} endpoints support federated queries, \emph{i.e.} run subqueries on other servers.
 For a personal user KG storage, a number of servers were developed, \emph{e.g.}, \texttt{GraphDB}, \texttt{Jena Fuseki}, \texttt{ClioPatria}.
 Each has its extension engines.
-\texttt{ClioPatria} extends with Prolog modules and allows one to implement KG processing on server side.
+\texttt{ClioPatria} is extended with Prolog modules and allows one to implement KG processing on server.
 
@@
 KG is represented with triple or quadruple set object.
-The set is traversed, triples are filtered according to various criteria.
+The set is traversed, triples are filtered according to various conditions.
 Filters are defined usually as patterns, where each of the components either a free variable/placeholder or a concrete value (IRI, literal), which must mach exactly.
-Some libraries allow one execute \texttt{SPARQL} queries locally.
+Some libraries allow one to execute \texttt{SPARQL} queries locally.
 KG are evolved with adding new triples, as well as removing irrelevant ones.
@@
 An example of a federated SPARQL query is presented in Figure~\ref{fig:sparql-ex1}, which results in counting parts classes as the first level of a three widget.
-Each class is named in Russian if \texttt{DBPedia} contains a corresponding label.
+Each class is named in Russian if \texttt{DBPedia} contains a corresponding label with \texttt{ru} language tag.
 Otherwise label is figured out of IRI.
-This query is running at a local server and initiates a subquery to \texttt{DBPedia.org} endpoint.
+This query is initially run at a local server and server initiates a subquery to \texttt{DBPedia.org} endpoint.
 
@@
 \end{minted}
-  \caption{Example of a federated SPARQL query usage in a electronic component database}
+  \caption{Example of a federated SPARQL query usage in an electronic component database}
   \label{fig:sparql-ex1}
@@
 
-Formalized representation of a domain, which is provided by SW techniques, is the basis of automatic  inference construction, thus modeling, \emph{e.g.}, problem-solving, resulting decision synthesis, theorem proving, domain verification.
-The inference is implemented in various ways, all of them belong either to semantic inference, usually implemented as solving a corresponding CSP\footnote{Abbreviation of Constraint Satisfaction Problem}, or syntax-driven procedures that mediate the previous structures and infer one of special properties (Automatic Theorem Proving).
+Formalized representation of a domain provided by SW techniques is the basis of automatic inference construction, thus modeling, \emph{e.g.}, problem-solving, resulting in decision synthesis, theorem proving, domain verification.
+The inference is implemented in various ways, all of them belong either to semantic ones, usually implemented as solving a corresponding CSP\footnote{Abbreviation of Constraint Satisfaction Problem}, or syntax-driven procedures that mediate the previous structures and infer one of special properties (Automatic Theorem Proving).
 Theoretic research is being carried on classification the properties of vocabularies by complexity of verification and development software implementing it for these classes.
-These software are parts of RDF-processing libraries and stand-alone system extensions.
+This software is a  part of RDF-processing libraries and stand-alone system extensions.
 
-Our interest is located in inferring new objects using logic programming from domain models represented as KGs.
-It allows developer to combine KG processing with other inferences in the same programming paradigm.
+Our interest is located in synthesizing (logically inferring) new objects using logic programming (LP) from domain models represented as KGs.
+LP allows developer to combine KG processing with other inferences in the same programming paradigm.
 SWI-Prolog programming language environment and Logtalk extension is used for the purpose.
 SWI system has library for representing and processing KG, and Logtalk is used to realize transformations providing the inference.
-Object-oriented properties of Logtalk enable one to cope with generic disadvantaged of RDF\footnote{Requirements of triple tuple data representation.} with encapsulation, manipulate knowledge sets with extension, inheritance and composition, and represent inference in a generalized form of objects sending messages to each other.
+Object-oriented properties of Logtalk enable one to cope with generic disadvantaged of RDF with encapsulation, manipulate knowledge sets with object composition instruments such as extension, inheritance and composition, and represent inference in a generalized form of objects sending messages to each other.
+
+In order to present the general architecture of application development, we have to briefly speak about KG definitions and its useful properties.
+% \subsection{Knowledge graph properties as application basis}
+% \label{sec:kg-as-basis}
+KGs changed the view point to the knowledge base as being a part of whole totality of knowledge, implying the obeying the global standards and techniques of its acquisition and processing \cite{hogan}.
+The following outlines their definitions \cite{hogan}.
+ \begin{itemize}
+  \item Notions \emph{data} and \emph{knowledge} are converged into ``\emph{something} \emph{known}'';
+  \item KG contains data and knowledge in form of resources and relations, including metadata;
+  \item The same KG is being filled in and processed by different independent modules, \emph{e.g.}, withing SPARQL queries with UPDATE statements;
+  \item KG fundamentals allow software developer to \emph{postpone} the formal definition of a data schema;
+  \item There are three points of view to graph schemata interpretation:
+    \begin{itemize}
+    \item \emph{semantic}, which is aimed at domain modeling paying attention to generalization relations (classes);
+    \item \emph{validating}, reflecting semantic verification, checking completeness with respect to sets of predefined general relations, and
+    \item \emph{emergent}, observed or deduced from exploitation experience analysis, as a set of new generalized structures and \emph{refactoring} of the KG.
+    \end{itemize}
+  \end{itemize}
+
+In our projects, KG stores global data, knowledge, and models.
+This is due to a general considerations not to overfull it with secondary plane data, which of cause can form other KG resources and used in the federated queries.
+Such way partially solves the main technical problems of KG storage and query execution.
+Being a triple set, whole KGs are usually stored in computer RAM, their structures expose no heuristic information to as query engines, which could be used for optimizations as compared to relation databases technologies.
 
@@
 
-\section{General architecture of applications based on KG}
+\section{General architecture of application environment}
 \label{sec:architecture}
 
-Some general words about architectures .... related to WEB.
+With general constraints implied by KG based domain modeling, design of application modules are to be built on standardized data interchange formats, which are to be easily converted from and to KG representation, as well as applications are to be constructed out of service modules supporting SW endpoints (SPARQL) whenever possible.
 
-In order to present the general architecture of application development, we have to briefly speak about KG definitions and properties.
+A convenient approach of application construction is usage component architectures, where subsystems in runtime are components \emph{providing} interfaces using other components' service available via corresponding interfaces.
+Zope component architecture \cite{zca}, Logtalk, React.js are used to \emph{implement} interfaces.
+Python allows wrap whole subsystems, \emph{e.g.}, Elastic search, in imperative paradigm, and Logtalk is chosen for the declarative counterpart.
+Inter-process component communications is realized usually with \texttt{HTTP} protocol, representing data in XML and JSON formats.
 
-\subsection{Knowledge graph properties as application basis}
-\label{sec:kg-as-basis}
+Our R\&D resulted in a set of components and a general application architecture showed in Figure~\ref{fig:arch-services}.
+\begin{figure}[bth]
+  \begin{center}
+    \includegraphics[width=0.9\linewidth]{architecture-mda-lod-ext-general.pdf}
+  \end{center}
+  \begin{center}
+      \textbf{Abbreviations}%\\[1ex]
+    \end{center}
+    \begin{quote}\scriptsize
+      % T-Module is Transformation module,
+      MDA is Model-Driven Architecture,
+      CIM is Computationally Independent Model,
+      PIM is Platform Independent Model,
+      PSM is Platform Specific Model,
+      T-Box is Terminological Box,
+      KB is Knowledge Base,
+      % A-Box is Instance Box,
+      % NGS is Next-Generation Sequencing,
+      DB is Database.
+    \end{quote}
+  \caption{Architecture of application instrumental environment as a set of components}
+  \label{fig:arch-services}
+\end{figure}
+The kernel module of applications is \texttt{KG server} (B2) built on the base of ClioPatria \cite{cliopatria} with web-server and Pengines plug-ins.
+ClioPatria is a web-application implemented in SWI-Prolog \cite{swi} being ISO-compliant Prolog implementation.
+KG functions are realized with SW package, supplied in standard SWI-Prolog distribution.
+Pengines service enables SW web-applications utilize logical inference on the server side \cite{pengines} with browser JavaScript environment integration.
+Pengine's knowledge base are extended with programmer's Prolog modules.
+Kernel module stores a common global part of A-~and T-boxes of applications, \emph{i.e.} it is application database.
+Other modules provide application (C3) with services.
 
-KGs changed the view point to the knowledge base as being a part of whole totality of knowledge, implying the obeying the global standards and techniques of its acquisition and processing \cite{hogan}.
-The following outlines their definitions \cite{hogan}.
- \begin{itemize}
-  \item Notions \emph{data} and \emph{knowledge} are converged into ``something is \emph{known}'';
-  \item KG contains, in general, data, relations, and metadata;
-  \item The same KG is being filled in and processed by different independent modules, \emph{e.g.}, withing SPARQL queries with UPDATE statements;
-  \item KG fundamental software development allow to \emph{postpone} the formal definition of a data schema;
-  \item There is three points of view to graph schemata: \emph{semantic}, which is aimed at generalization, \emph{validating}, reflecting semantic verification, checking completeness w.r.t.
-sets of predefined general relations, and \emph{emergent}, observed or deduced from exploitation experience, as a set of new generalized structures and \emph{refactoring} of the KG.
-  \end{itemize}
+Authoring tool (B1) is used to provide document generation and integration \cite{sibiricon}.
+This tool enables document content and structure inference in browser utilizing templates and data from other web pages and websites providing LOD capabilities.
+Programmer implements a view of an object as a web page with an extended version of RDFa.
+The page are generated within any programming runtime, \emph{e.g.} PHP.
+Authoring tool engine collects web page chunks and composing the document out of these chunks as a XHTML page.
+This composed page can be stored in a KG as set of triples and indexed in Text indexing module (C1), printed with browser's printing engine, or converted to PDF.
+The tool allow user to change document content in WISYWIG mode in some extent.
+Its general structure is defined by changing template at server side.
+
+Text indexing engine (C1) provides service for storing documents represented as KG triples, JSONs, and BLOBS of any format with corresponding text layers.
+This module has two implementations.
+One is build on Sphinx Search engine, and the second one is Elasticsearch.
+Elasticsearch supports JSON as the only document format and easily integrated with KG using JSON-LD for triple representation.
+Sphinx Search is much faster in text and database records indexing and consumes less RAM as it is implemented in C++.
+The choice of an engine depends on the primary document representation to be used in application.
+
+BLOB stored documents having no markup (PDFs, scanned documents) usually contain useful data, which are to be revealed for application use.
+The documents could be report data or scientific paper content, DJVU files or raster images.
+Data recognition is realized in PDF/Text document processing module (C2).
+Data processing workers query Text indexing module for BLOB documents and add recognized text and high-order structure layers.
+All obtained layers are stored in the database of C1, as well as data of general interest are to be converted in triples of KG.
+
+A generalized KG processing is being executed in KB processing module (B3).
+This component denote a set of rules used to implement the validating and emergent semantics, as well as analysis and synthesis of the new data, including resultant output.
+Module A2 is the source of model data (design of domains, UML, problem statement, \emph{etc.}), which is converted and form T-box of KG.
+The conversion is a kind of Model-driven architecture (MDA) implementation.
+Transformation is carried out under control of Platform model describing a context of the conversion, for example, in case of software source code synthesis, software platform used in implementation of model objects.
+
+Metadata processing module (A1) denote the environment capabilities for expressing semantics of data processing module output out of its input data semantics.
+This allows us to store resultant data in KG for the further usage in module (B3) for problem solving and decision making.
+Module A3 denote external services and KG with useful resources, \emph{e.g.}, DBPedia global objects.
+
+\section{Implementation examples}
+
+In this section, three examples of application implementation are being presented.
 

### 2022-02-01 paper-aicts-cherkashin.tex
Grammarly web grammar checking. Removed CEUR WS instructions.

@@
 %% Options:
-%% twocolumn : Two column layout.
+%% two-column: Two-column layout.
 %% hf: enable header and footer.
@@
   Semantic Web technologies and their standards are becoming productive assets for software development and integration on various design and implementation levels.
-These include OSI level 6 data representation, database and file storage formats, the source data for text indexes, document authoring, hypertext markup engines, a representation for abstract software models, \emph{etc}.
-They are also smoothly combined with development and data analysis tools, giving raise new common-ground environment for realizing an integration of distributed applications on a semantic level.
+These include OSI level 6 data representation, database and file storage formats, the source data for text indexes, document authoring, hypertext markup engines, a representation for abstract software models, \emph{etc.}  They are also smoothly combined with the development and data analysis tools, giving rise to a new common-ground environment for integrating distributed applications on a semantic level.
 
@@
 
-Since 2000 the development of the Semantic Web (SW) technologies became actively used usually in describing software domains, representing semantic relations between objects.
-For some domains on the base of the SW description, first-order knowledge theories are constructed \cite{}, but in practical aspect of the SW utilization, the solving data representation and software integration aspects prevail.
+Since 2000, the Semantic Web (SW) technologies have become actively used in describing software domains, representing semantic relations between objects.
+For some domains on the base of the SW description, first-order knowledge theories are constructed \cite{}, but in the practical aspect of the SW utilization, the solving data representation and software integration aspects prevail.
 Some tools for automation of conceptual model design and model conversion were proposed \cite{}.
@@
 
-Unification of routine tasks resulted in development common techniques for storing and processing data, which are aggregated around term ``knowledge graph'' (KG).
-These include data representation formats, vocabulary standards, query languages, tools for accessing and management data in a KG.
+Unification of routine tasks resulted in standard development techniques for storing and processing data aggregated around the term ``knowledge graph'' (KG).
+These include data representation formats, vocabulary standards, query languages, tools for accessing and managing data in a KG.
 Software development techniques are mediated with properties of the new-created resources and their APIs\footnote{Abbreviation of Application Programming Interface}.
-Many internet resources provide common ways of formalized domain data access using query endpoints, and these endpoints allow one to organize federated KG data access.
+Many internet resources provide standard ways of formalized domain data access using query endpoints, and these endpoints allow one to organize federated KG data access.
 
 In Russia, domain model designers generally prefer to develop local vocabularies instead of spending time for the routine work of Internet surfing for existing conceptual models and their enrichment.
-It is due to underdevelopment of the metadata resources (metavocabularies) due to lack a domain classification system as compared, \emph{e.g.} to Universal Decimal Classification (UDC) used in bibliography.
-Vocabularies also can have common parts, describing the same or similar entities with different term and relations, \emph{e.g.} semantically more adopted to a specific problem solving.
+It is due to underdevelopment of the metadata resources (metavocabularies) due to lack of a domain classification system as compared, \emph{e.g.} to Universal Decimal Classification (UDC) used in the bibliography.
+Vocabularies also can have common parts, describing the same or similar entities with different terms and relations, \emph{e.g.} semantically more adapted to a specific problem-solving.
 
-Unitizing standardized vocabularies as domain conceptual models (ontologies) in our projects allow us to compare other research groups point of view to a problem with a formalized context, figure out underestimated and not fully formularized aspects of a domain under investigation or automation.
-The vocabularies are similar to information sources as scientific papers, but have a concrete formal presentation of the information objects, allowing one to get rid of particular uncertainties.
-Another advantage of standardization is a construction of a common ground of research, development, data formalization, and interchange formats and semantics.
-It includes an aggregation of results of pattern recognition, applications of machine learning with their following use as facts in knowledge-based systems.
+Unitizing standardized vocabularies as conceptual domain models (ontologies) in our projects allows us to compare other research groups' points of view to a problem with a formalized context and figure out underestimated and not fully formularized aspects of a domain under investigation or automation.
+The vocabularies are similar to information sources as scientific papers but have a concrete formal presentation of the information objects, allowing one to get rid of particular uncertainties.
+Another advantage of standardization is constructing a common ground of research, development, data formalization, and interchange formats and semantics.
+It includes an aggregation of pattern recognition results applications of machine learning with their following use as facts in knowledge-based systems.
 
-The aim of this paper is to present a contemporary experience of SW and KG utilization in the development of applications in a context of performance improvement of R\&D.
-Three examples were reviewed showing advantageous points of the technologies' application.
+This paper aims to present a contemporary experience of SW and KG utilization in the development of applications in a context of performance improvement of R\&D.
+Three examples were reviewed, showing good points of the technologies' application.
 
@@
 \item Graph presentation formats and techniques are implemented in files, databases, markups of documents, a data flow between applicants; \label{kg1}
-\item There are public semantic databases comprising the generic domain objects descriptions, which are mostly A-Boxes (databases), for example \texttt{DBPedia.org}, a formalized version of \texttt{Wikipedia.org}.
+\item There are public semantic databases comprising the generic domain objects descriptions, which are mostly A-Boxes (databases), for example, \texttt{DBPedia.org}, a formalized version of \texttt{Wikipedia.org}.
 % wikidata
-\item For graphs being databases a query services are implemented to allow users obtain its data by portions, as well as implement modifications;
+\item For graphs being databases, query services are implemented to allow users to obtain its data by portions, as well as implement modifications;
 \item API and libraries for processing files and documents with semantic data acquisition (\emph{e.g.} by extraction of the markup).
 \item Software for graph data processing to improve its structure or induct general relation patterns;
-\item Validation and verification services allowing one to explore consistency and completeness of the conceptual domain models and their implementations in software. \label{kg2}
+\item Validation and verification services allow one to explore consistency and completeness of the conceptual domain models and their implementations in software. \label{kg2}
 \end{enumerate}
-The assets \ref{kg1}--\ref{kg2} form knowledge graph technological aspect of the SW that gained a special interest within R\&D in the last decade.
+The assets  (\ref{kg1}--\ref{kg2}) form knowledge graph technological aspect of the SW that gained a particular interest within R\&D in the last decade.
 
-Each entity being described in SW has its unique global IRI, which is its virtual or real internet address.
+Each entity described in SW has its unique global IRI, its virtual or real internet address.
 Document entities' resource IRIs are formed accounting the document structure, \emph{e.g.} a hierarchy of contexts.
-Using IRIs in applications solves fundamental problem of cross-domain and cross-application object identification and forming a basis of common data processing environment.
+Using IRIs in applications solves the fundamental problem of cross-domain and cross-application object identification and forms a basis for a standardized data processing environment.
 Entities of various abstract levels can be designated an IRI, including vocabulary definitions (\emph{namespaces}), data types, data containers (files, databases).
 
-Triples are combination of two or three IRIs.
-Subject and predicate are always IRIs.
-Objects are either IRI (a reference to a resource) or a literal.
-Literals are values of basic types (number, string, date, time, \emph{etc.}) usually accompanied with metadata (type designator and/or a language tag).
-Types are defined in a vocabulary \texttt{<\url{http://www.w3.org/2001/XMLSchema}\#>}, having well-known namespace abbreviation \ns{xs}\footnote{Hereafter, all well-known namespace names will be given after its IRI in brackets with bold monospace font.}.
-Language tag is used to define string values, taking into account its affiliation to a cultural context.
+Triples are a combination of two or three IRIs (\texttt{<subjectIRI, predicateIRI, object>}).
+Subject and predicate are always IRIs, whereas objects are either IRI (a reference to a resource) or a literal.
+Literals are values of basic types (\emph{e.g.}, number, string, date, time) usually accompanied with metadata (type designator or a language tag).
+Types are defined in a vocabulary \texttt{<\url{http://www.w3.org/2001/XMLSchema}\#>}, having well-known namespace abbreviation \ns{xs}\footnote{Hereafter, all well-known namespace names will be given after its IRI in brackets with a bold monospace font.}.
+A language tag defines string values attribute, affiliating them to a cultural context.
 It is recognized with KG tools and utilized at presentation layers of the software.
 
-Namespaces are convenient instruments in triple definition.
+Namespaces are convenient instruments in the triple definition.
 For example \verb|integer| literal \verb|42^^<http://www.w3.org/2001/XMLSchema#integer>| can be reduced to \verb|42^^xs:integer|, where \verb|xs| is namespace prefix of term \verb|integer|.
-Most of KG representation formats allow one to define a default namespace, which terms are used without prefix.
-Almost any conceptual model of a domain represented as KG comprises entities taken from multiple vocabularies, and namespaces allow to specify the origin of a term.
-Three-component tuple (triple) being outside its KG context sometimes added the fourth component, its KG IRI (quadruple).
+Most KG representation formats allow one to define a default namespace, which is used without prefix.
+Almost any conceptual model of a domain represented as KG comprises entities taken from multiple vocabularies, and namespaces specify the origin of a term.
+Three-component tuple (triple) outside its KG context sometimes added the fourth KG IRI component (quadruple).
 
-Much R\&D efforts since 2001 \cite{tbl} were spend to develop SW data formats and their coupling with document representation.
-The simplest way to provide a printed document with semantic representation is pointing user's software agent to an URL serving the semantic data.
-But this approach requires to implement two different procedures for document generation and its metadata serving.
-That's why a RDFa\footnote{Abbreviation of Resource Description Framework attribute language} has been developed.
+Much R\&D efforts since 2001 \cite{tbl} were spent to develop SW data formats and their coupling with document representation.
+The simplest way to provide a printed document with semantic representation is by pointing the user's software agent to an URL serving the semantic data.
+However, this approach requires implementing two different procedures for document generation and its metadata serving.
+That is why an RDFa\footnote{Abbreviation of Resource Description Framework attribute language} has been developed.
 It allows developer of XHTML\footnote{HTML structures with strict XML restrictions obeyed} documents to enrich the document with its semantic layer.
 The semantic markup is represented with special attributes and their values.
-The hierarchical structure of XHTML is a matrix layer for hierarchical structure of document entities semantic relations.
-The layer is extracted by XHTML/RDFa processing libraries.
+The hierarchical structure of XHTML is a matrix layer for the hierarchical structure of document entities' semantic relations.
+XHTML/RDFa processing libraries extract the layer.
 There are other formats having the same expression capabilities as RDF but used for other purposes: \texttt{N-triples}, \texttt{Turtle}, \texttt{Notation~3}, \texttt{JSON-LD}.
-The last one is an efficient way of coupling SW and various generic data storage, indexing and processing services.
+The last is an efficient way of coupling SW and various generic data storage, indexing, and processing services.
 
-Among published global vocabularies, \texttt{DBPedia} is of most valuable for our projects.
-It is used for concrete object or notion of physical world reference, \emph{e.g.} the term ``passport'' in law documents authoring \cite{}, naming (labeling) referenced objects in user interface with corresponding words in natural language \cite{github-db}.
+Among published global vocabularies, \texttt{DBPedia} has the most value for our projects.
+It is used for concrete object or notion of physical world reference, \emph{e.g.} the term ``passport'' in documents authoring \cite{}, naming (labeling) referenced objects in a user interface with corresponding words in natural language \cite{github-db}.
 Another useful resource is \texttt{WordNet} converted in RDF triples.
-It describes semantics of english words and various their mutual relations.
+It describes the semantics of English words and various their mutual relations.
 
@@
 KGs and their subgraphs are programmatically processed using libraries developed for all papular programming environments.
-KG is represented with triple or quadruple set object.
+KG is represented with a triple or quadruple set object.
 The set is traversed, triples are filtered according to various conditions.
-Filters are defined usually as patterns, where each of the components either a free variable/placeholder or a concrete value (IRI, literal), which must mach exactly.
+Filters are usually defined as patterns, where each component is either a free variable/placeholder or a concrete value (IRI, literal), which must match exactly.
 Some libraries allow one to execute \texttt{SPARQL} queries locally.
-KG are evolved with adding new triples, as well as removing irrelevant ones.
+KG is evolved by adding new triples and removing irrelevant ones.
 The resulting modified KG is stored as files and published on a server.
@@
 Each class is named in Russian if \texttt{DBPedia} contains a corresponding label with \texttt{ru} language tag.
-Otherwise label is figured out of IRI.
-This query is initially run at a local server and server initiates a subquery to \texttt{DBPedia.org} endpoint.
+Otherwise, the label is figured out of IRI.
+This query is initially run at a local server, and the server initiates a subquery to \texttt{DBPedia.org} endpoint.
 
@@
 Formalized representation of a domain provided by SW techniques is the basis of automatic inference construction, thus modeling, \emph{e.g.}, problem-solving, resulting in decision synthesis, theorem proving, domain verification.
-The inference is implemented in various ways, all of them belong either to semantic ones, usually implemented as solving a corresponding CSP\footnote{Abbreviation of Constraint Satisfaction Problem}, or syntax-driven procedures that mediate the previous structures and infer one of special properties (Automatic Theorem Proving).
-Theoretic research is being carried on classification the properties of vocabularies by complexity of verification and development software implementing it for these classes.
+The inference is implemented in various ways.
+All of them belong either to semantic ones, usually implemented as solving a corresponding CSP\footnote{Abbreviation of Constraint Satisfaction Problem}, or syntax-driven procedures that mediate the previous structures and infer one of the specific properties (Automatic Theorem Proving).
+Theoretic research is being carried on classification the properties of vocabularies by the complexity of verification and development software implementing it for these classes.
 This software is a  part of RDF-processing libraries and stand-alone system extensions.
 
-Our interest is located in synthesizing (logically inferring) new objects using logic programming (LP) from domain models represented as KGs.
-LP allows developer to combine KG processing with other inferences in the same programming paradigm.
-SWI-Prolog programming language environment and Logtalk extension is used for the purpose.
-SWI system has library for representing and processing KG, and Logtalk is used to realize transformations providing the inference.
-Object-oriented properties of Logtalk enable one to cope with generic disadvantaged of RDF with encapsulation, manipulate knowledge sets with object composition instruments such as extension, inheritance and composition, and represent inference in a generalized form of objects sending messages to each other.
+Our interest is synthesizing (logically inferring) new objects using logic programming (LP) from domain models represented as KGs.
+LP allows a developer to combine KG processing with other inferences in the same programming paradigm.
+SWI-Prolog programming language environment and Logtalk extension is used for this purpose.
+SWI system has a library for representing and processing KG, and Logtalk is used to implement transformations providing the inference.
+Object-oriented properties of Logtalk enable one to cope with generic disadvantages of RDF with encapsulation, manipulate knowledge sets with object composition instruments such as extension, inheritance, and composition, and represent inference in a generalized form of objects sending messages to each other.
 
-In order to present the general architecture of application development, we have to briefly speak about KG definitions and its useful properties.
+In order to present the general architecture of application development, we have to speak about KG definitions and useful properties briefly of KGs.
 % \subsection{Knowledge graph properties as application basis}
 % \label{sec:kg-as-basis}
-KGs changed the view point to the knowledge base as being a part of whole totality of knowledge, implying the obeying the global standards and techniques of its acquisition and processing \cite{hogan}.
+KGs changed the viewpoint to the knowledge base as being a part of the whole totality of knowledge, implying the obeying the global standards and techniques of its acquisition and processing \cite{hogan}.
 The following outlines their definitions \cite{hogan}.
@@
   \item Notions \emph{data} and \emph{knowledge} are converged into ``\emph{something} \emph{known}'';
-  \item KG contains data and knowledge in form of resources and relations, including metadata;
+  \item KG contains data and knowledge in the form of resources and relations, including metadata;
   \item The same KG is being filled in and processed by different independent modules, \emph{e.g.}, withing SPARQL queries with UPDATE statements;
-  \item KG fundamentals allow software developer to \emph{postpone} the formal definition of a data schema;
+  \item KG fundamentals allow a software developer to \emph{postpone} the formal definition of a data schema;
   \item There are three points of view to graph schemata interpretation:
@@
 In our projects, KG stores global data, knowledge, and models.
-This is due to a general considerations not to overfull it with secondary plane data, which of cause can form other KG resources and used in the federated queries.
-Such way partially solves the main technical problems of KG storage and query execution.
-Being a triple set, whole KGs are usually stored in computer RAM, their structures expose no heuristic information to as query engines, which could be used for optimizations as compared to relation databases technologies.
+It is due to general considerations not to overfill it with secondary plane data, which of cause can form other KG resources and be used in the federated queries.
+Such a way partially solves the main technical problems of KG storage and query execution.
+Being a triple set, whole KGs are usually stored in computer RAM, and their structures expose no heuristic information to query engines, which could be used for optimizations compared to relation databases technologies.
 
@@
 
-With general constraints implied by KG based domain modeling, design of application modules are to be built on standardized data interchange formats, which are to be easily converted from and to KG representation, as well as applications are to be constructed out of service modules supporting SW endpoints (SPARQL) whenever possible.
+With general constraints implied by KG based domain modeling, the design of application modules are to be built on standardized data interchange formats, which are to be easily converted from and to KG representation, as well as applications are to be constructed out of service modules supporting SW endpoints (SPARQL) whenever possible.
 
@@
 Zope component architecture \cite{zca}, Logtalk, React.js are used to \emph{implement} interfaces.
-Python allows wrap whole subsystems, \emph{e.g.}, Elastic search, in imperative paradigm, and Logtalk is chosen for the declarative counterpart.
+Python allows to wrap whole subsystems, \emph{e.g.}, Elastic search, in imperative paradigm, and Logtalk is chosen for the declarative counterpart.
 Inter-process component communications is realized usually with \texttt{HTTP} protocol, representing data in XML and JSON formats.
@@
 ClioPatria is a web-application implemented in SWI-Prolog \cite{swi} being ISO-compliant Prolog implementation.
-KG functions are realized with SW package, supplied in standard SWI-Prolog distribution.
-Pengines service enables SW web-applications utilize logical inference on the server side \cite{pengines} with browser JavaScript environment integration.
-Pengine's knowledge base are extended with programmer's Prolog modules.
+KG functions are implemented with SW package, supplied in standard SWI-Prolog distribution.
+Pengines service enables SW web-applications to utilize logical inference on the server-side \cite{pengines} with browser JavaScript environment integration.
+Pengine's knowledge base is extended with the programmer's Prolog modules.
 Kernel module stores a common global part of A-~and T-boxes of applications, \emph{i.e.} it is application database.
-Other modules provide application (C3) with services.
+Other modules provide an application (C3) with services.
 
 Authoring tool (B1) is used to provide document generation and integration \cite{sibiricon}.
-This tool enables document content and structure inference in browser utilizing templates and data from other web pages and websites providing LOD capabilities.
+This tool enables document content and structure inference in the browser utilizing templates and data from other web pages and websites providing LOD capabilities.
 Programmer implements a view of an object as a web page with an extended version of RDFa.
-The page are generated within any programming runtime, \emph{e.g.} PHP.
-Authoring tool engine collects web page chunks and composing the document out of these chunks as a XHTML page.
-This composed page can be stored in a KG as set of triples and indexed in Text indexing module (C1), printed with browser's printing engine, or converted to PDF.
-The tool allow user to change document content in WISYWIG mode in some extent.
-Its general structure is defined by changing template at server side.
+The page is generated within any programming runtime, \emph{e.g.} PHP.
+The authoring tool engine collects web page chunks and composes the document as an XHTML page.
+This composed page can be stored in a KG as a set of triples and indexed in the Text indexing module (C1), printed with the browser's printing engine, or converted to PDF.
+The tool allows the user to change document content in WYSIWYG mode.
+Its general structure is defined by changing the template on the server-side.
 
-Text indexing engine (C1) provides service for storing documents represented as KG triples, JSONs, and BLOBS of any format with corresponding text layers.
+Text indexing engine (C1) provides a service for storing documents represented as KG triples, JSONs, and BLOBS of any format with corresponding text layers.
 This module has two implementations.
-One is build on Sphinx Search engine, and the second one is Elasticsearch.
-Elasticsearch supports JSON as the only document format and easily integrated with KG using JSON-LD for triple representation.
+One is built on the Sphinx Search engine, and the second is Elasticsearch.
+Elasticsearch supports JSON as the only document format and is easily integrated with KG using JSON-LD for triple representation.
 Sphinx Search is much faster in text and database records indexing and consumes less RAM as it is implemented in C++.
-The choice of an engine depends on the primary document representation to be used in application.
+The choice of an engine depends on the primary document representation used in an application.
 
-BLOB stored documents having no markup (PDFs, scanned documents) usually contain useful data, which are to be revealed for application use.
-The documents could be report data or scientific paper content, DJVU files or raster images.
-Data recognition is realized in PDF/Text document processing module (C2).
-Data processing workers query Text indexing module for BLOB documents and add recognized text and high-order structure layers.
-All obtained layers are stored in the database of C1, as well as data of general interest are to be converted in triples of KG.
+BLOB stored documents having no markup (PDFs, scanned documents) usually contain valuable data to be revealed for application use.
+The documents could be report data or scientific paper content, DJVU files, or raster images.
+Data recognition is implemented in the PDF/Text document processing module (C2), where data processing workers query the Text indexing module for BLOB documents and add recognized text and high-order structure layers.
+All obtained layers are stored in the database of C1, and data of general interest are to be converted in triples of KG.
 
-A generalized KG processing is being executed in KB processing module (B3).
-This component denote a set of rules used to implement the validating and emergent semantics, as well as analysis and synthesis of the new data, including resultant output.
-Module A2 is the source of model data (design of domains, UML, problem statement, \emph{etc.}), which is converted and form T-box of KG.
+A generalized KG processing is being executed in the KB processing module (B3).
+This component denotes a set of rules used to implement the validating and emergent semantics and analysis and synthesis of the new data, including resultant output.
+Module A2 is the source of model data (design of domains, UML, problem statement, \emph{etc.}), which is converted and form the T-box of KG.
 The conversion is a kind of Model-driven architecture (MDA) implementation.
-Transformation is carried out under control of Platform model describing a context of the conversion, for example, in case of software source code synthesis, software platform used in implementation of model objects.
+Transformation is carried out under the control of the Platform model, describing a context of the conversion.
+For example, in the case of software source code synthesis, the software platform model is used to implement model objects.
 
-Metadata processing module (A1) denote the environment capabilities for expressing semantics of data processing module output out of its input data semantics.
-This allows us to store resultant data in KG for the further usage in module (B3) for problem solving and decision making.
-Module A3 denote external services and KG with useful resources, \emph{e.g.}, DBPedia global objects.
+The metadata processing module (A1) denotes the environment capabilities for expressing the semantics out of the data processing modules' output and the input data semantics.
+It allows us to store resultant data in KG for further usage in module (B3) for problem-solving and decision-making.
+Module A3 denote external services and KG with valuable resources, \emph{e.g.}, DBPedia global objects.
 
@@
 
-CEUR-WS's article template provides a consistent \LaTeX{} style for
-use across CEUR-WS publications, and incorporates accessibility and
-metadata-extraction functionality. This document will explain the
-major features of the document class.
-
-If you are new to publishing with CEUR-WS, this document is a valuable
-guide to the process of preparing your work for publication.
-
-The ``\verb|ceurart|'' document class can be used to prepare articles
-for any CEUR-WS publication, and for any stage of publication, from
-review to final ``camera-ready'' copy with {\itshape very} few changes
-to the source.
-
-This class depends on the following packages
-for its proper functioning:
-
-\begin{itemize}
-\item \verb|natbib.sty| for citation processing;
-\item \verb|geometry.sty| for margin settings;
-\item \verb|graphicx.sty| for graphics inclusion;
-\item \verb|hyperref.sty| optional package if hyperlinking is required in
-  the document;
-\item \verb|fontawesome5.sty| optional package for bells and whistles.
-\end{itemize}
-
-All the above packages are part of any
-standard \LaTeX{} installation.
-Therefore, the users need not be
-bothered about downloading any extra packages.
-
-\section{Modifications}
-
-Modifying the template --- including but not limited to: adjusting
-margins, typeface sizes, line spacing, paragraph and list definitions,
-and the use of the \verb|\vspace| command to manually adjust the
-vertical spacing between elements of your work --- is not allowed.
-
-\section{Template parameters}
-
-There are a number of template
-parameters which modify some part of the \verb|ceurart| document class.
-This parameters are enclosed in square
-brackets and are a part of the \verb|\documentclass| command:
-% \begin{lstlisting}
-%   \documentclass[parameter]{ceurart}
-% \end{lstlisting}
-
-Frequently-used parameters, or combinations of parameters, include:
-\begin{itemize}
-\item {\verb|twocolumn|}: Two column layout.
-\item {\verb|hf|}: Enable header and footer\footnote{You can enable
-    the display of page numbers in the final version of the entire
-    collection. In this case, you should adhere to the end-to-end
-    pagination of individual papers.}.
-\end{itemize}
-
-\subsection{Title Information}
-
-The titles of papers should be either all use the emphasizing
-capitalized style or they should all use the regular English (or
-native language) style. It does not make a good impression if you or
-your authors mix the styles.
-
-Use the \verb|\title| command to define the title of your work. Do not
-insert line breaks in your title.
-
-\subsection{Title variants}
-
-% \verb|\title| command have the below options:
-% \begin{itemize}
-% \item \verb|title|: Document title. This is default option.
-% \begin{lstlisting}
-% \title[mode=title]{This is a title}
-% \end{lstlisting}
-% You can just omit it, like as follows:
-% \begin{lstlisting}
-% \title{This is a title}
-% \end{lstlisting}
-
-% \item \verb|alt|: Alternate title.
-% \begin{lstlisting}
-% \title[mode=alt]{This is a alternate title}
-% \end{lstlisting}
-
-% \item \verb|sub|: Sub title.
-% \begin{lstlisting}
-% \title[mode=sub]{This is a sub title}
-% \end{lstlisting}
-% You can just use \verb|\subtitle| command, as follows:
-% \begin{lstlisting}
-% \subtitle{This is a sub title}
-% \end{lstlisting}
-
-% \item \verb|trans|: Translated title.
-% \begin{lstlisting}
-% \title[mode=trans]{This is a translated title}
-% \end{lstlisting}
-
-% \item \verb|transsub|: Translated sub title.
-% \begin{lstlisting}
-% \title[mode=transsub]{This is a translated sub title}
-% \end{lstlisting}
-% \end{itemize}
-
-\subsection{Authors and Affiliations}
-
-Each author must be defined separately for accurate metadata
-identification. Multiple authors may share one affiliation. Authors'
-names should not be abbreviated; use full first names wherever
-possible. Include authors' e-mail addresses whenever possible.
-
-\verb|\author| command have the below options:
-
-\begin{itemize}
-\item \verb|style|: Style of author name (chinese)
-\item \verb|prefix|: Prefix
-\item \verb|suffix|: Suffix
-\item \verb|degree|: Degree
-\item \verb|role|: Role
-\item \verb|orcid|: ORCID
-\item \verb|email|: E-mail
-\item \verb|url|: URL
-\end{itemize}
-
-Author names can have some kinds of marks and notes:
-\begin{itemize}
-\item affiliation mark: \verb|\author[<num>]|.
-% \item email: \verb|\ead{<email>}|,
-% \item url: \verb|\ead[url]{<url>}|.
-\end{itemize}
-
-The author names and affiliations could be formatted in two ways:
-\begin{enumerate}
-\item Group the authors per affiliation.
-\item Use an explicit mark to indicate the affiliations.
-\end{enumerate}
-
-Author block example:
-% \begin{lstlisting}
-% \author[1,2]{Author Name}[%
-%     prefix=Prof.,
-%     degree=D.Sc.,
-%     role=Researcher,
-%     orcid=0000-0000-000-0000,
-%     email=name@example.com,
-%     url=https://name.example.com
-% ]
-
-% \address[1]{Affiliation #1}
-% \address[2]{Affiliation #2}
-% \end{lstlisting}
-
-% \subsection{Abstract and Keywords}
-
-% Abstract shall be entered in an environment that starts
-% with \verb|\begin{abstract}| and ends with
-% \verb|\end{abstract}|.
-
-% \begin{lstlisting}
-% \begin{abstract}
-%   This is an abstract.
-% \end{abstract}
-% \end{lstlisting}
-
-% The key words are enclosed in a \verb|keywords|
-% environment. Use \verb|\sep| to separate keywords.
-
-% \begin{lstlisting}
-% \begin{keywords}
-%   First keyword \sep
-%   Second keyword \sep
-%   Third keyword \sep
-%   Fourth keyword
-% \end{keywords}
-% \end{lstlisting}
-
-At the end of front matter add \verb|\maketitle| command.
-
-\section{Sectioning Commands}
-
-Your work should use standard \LaTeX{} sectioning commands:
-\verb|\section|, \verb|\subsection|,
-\verb|\subsubsection|, and
-\verb|\paragraph|. They should be numbered; do not remove
-the numbering from the commands.
-
-Simulating a sectioning command by setting the first word or words of
-a paragraph in boldface or italicized text is not allowed.
-
-\section{Tables}
-
-The ``\verb|ceurart|'' document class includes the ``\verb|booktabs|''
-package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
-high-quality tables.
-
-Table captions are placed \textit{above} the table.
-
-Because tables cannot be split across pages, the best placement for
-them is typically the top of the page nearest their initial cite.  To
-ensure this proper ``floating'' placement of tables, use the
-environment \verb|table| to enclose the table's contents and the
-table caption. The contents of the table itself must go in the
-\verb|tabular| environment, to be aligned properly in rows and
-columns, with the desired horizontal and vertical rules.
-
-Immediately following this sentence is the point at which
-Table~\ref{tab:freq} is included in the input file; compare the
-placement of the table here with the table in the printed output of
-this document.
-
-\begin{table*}
-  \caption{Frequency of Special Characters}
-  \label{tab:freq}
-  \begin{tabular}{ccl}
-    \toprule
-    Non-English or Math&Frequency&Comments\\
-    \midrule
-    \O & 1 in 1,000& For Swedish names\\
-    $\pi$ & 1 in 5& Common in math\\
-    \$ & 4 in 5 & Used in business\\
-    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-  \bottomrule
-\end{tabular}
-\end{table*}
-
-To set a wider table, which takes up the whole width of the page's
-live area, use the environment \verb|table*| to enclose the table's
-contents and the table caption.  As with a single-column table, this
-wide table will ``float'' to a location deemed more
-desirable. Immediately following this sentence is the point at which
-Table~\ref{tab:commands} is included in the input file; again, it is
-instructive to compare the placement of the table here with the table
-in the printed output of this document.
-
-\begin{table}
-  \caption{Some Typical Commands}
-  \label{tab:commands}
-  \begin{tabular}{ccl}
-    \toprule
-    Command &A Number & Comments\\
-    \midrule
-    \texttt{{\char'134}author} & 100& Author \\
-    \texttt{{\char'134}table}& 300 & For tables\\
-    \texttt{{\char'134}table*}& 400& For wider tables\\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\section{Math Equations}
-
-You may want to display math equations in three distinct styles:
-inline, numbered or non-numbered display.  Each of the three are
-discussed in the next sections.
-
-\subsection{Inline (In-text) Equations}
-
-A formula that appears in the running text is called an inline or
-in-text formula.  It is produced by the \verb|math| environment,
-which can be invoked with the usual
-\verb|\begin| \ldots \verb|\end| construction or with
-the short form \verb|$| \ldots \verb|$|. You can use any of the symbols
-and structures, from $\alpha$ to $\omega$, available in
-\LaTeX~\cite{Lamport:LaTeX};
-this section will simply show a few
-examples of in-text equations in context. Notice how this equation:
-\begin{math}
-  \lim_{n\rightarrow \infty} \frac{1}{n} = 0,
-\end{math}
-set here in in-line math style, looks slightly different when
-set in display style.  (See next section).
-
-\subsection{Display Equations}
-
-A numbered display equation---one set off by vertical space from the
-text and centered horizontally---is produced by the \verb|equation|
-environment. An unnumbered display equation is produced by the
-\verb|displaymath| environment.
-
-Again, in either environment, you can use any of the symbols and
-structures available in \LaTeX{}; this section will just give a couple
-of examples of display equations in context.  First, consider the
-equation, shown as an inline equation above:
-\begin{equation}
-  \lim_{n\rightarrow \infty} \frac{1}{n} = 0.
-\end{equation}
-Notice how it is formatted somewhat differently in
-the \verb|displaymath|
-environment.  Now, we'll enter an unnumbered equation:
-\begin{displaymath}
-  S_{n} = \sum_{i=1}^{n} x_{i} ,
-\end{displaymath}
-and follow it with another numbered equation:
-\begin{equation}
-  \lim_{x \to 0} (1 + x)^{1/x} = e
-\end{equation}
-just to demonstrate \LaTeX's able handling of numbering.
-
-\section{Figures}
-
-The ``\verb|figure|'' environment should be used for figures. One or
-more images can be placed within a figure. If your figure contains
-third-party material, you must clearly identify it as such, as shown
-in the example below.
-\begin{figure}
-  \centering
-  % \includegraphics[width=\linewidth]{sample-franklin}
-  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
-    Ewing, Inc. [Public domain], via Wikimedia
-    Commons. (\url{https://goo.gl/VLCRBB}).}
-\end{figure}
-
-Your figures should contain a caption which describes the figure to
-the reader. Figure captions go below the figure. Your figures should
-also include a description suitable for screen readers, to
-assist the visually-challenged to better understand your work.
-
-Figure captions are placed below the figure.
-
-\section{Citations and Bibliographies}
-
-The use of Bib\TeX{} for the preparation and formatting of one's
-references is strongly recommended. Authors' names should be complete
---- use full first names (``Donald E. Knuth'') not initials
-(``D. E. Knuth'') --- and the salient identifying features of a
-reference should be included: title, year, volume, number, pages,
-article DOI, etc.
-
-The bibliography is included in your source document with these two
-commands, placed just before the \verb|\end{document}|
-command:
-% \begin{lstlisting}
-% \bibliography{bibfile}
-% \end{lstlisting}
-where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
-suffix, of the Bib\TeX{} file.
-
 \begin{acknowledgments}
 
-  The results were obtained within the state assignment of the Ministry of Education and Science of Russia, the project ``Methods and technologies of a cloud-based service-oriented digital platform for collecting, storing and processing large volumes of multi-format interdisciplinary data and knowledge based on the use of artificial intelligence, a model-driven approach and machine learning'', No.~FWEW-2021-0005 (State registration No.~121030500071-2).
+The results were obtained within the state assignment of the Ministry of Education and Science of Russia, the project ``Methods and technologies of a cloud-based service-oriented digital platform for collecting, storing and processing large volumes of multi-format interdisciplinary data and knowledge based on the use of artificial intelligence, a model-driven approach, and machine learning'', No.~FWEW-2021-0005 (State registration No.~121030500071-2).
 
-  The results obtained with the use of the network infrastructure of Telecommunication center of collective use ``Integrated information-computational network of Irkutsk scientific-educational complex'' (\url{http://net.icc.ru}).
+The results were obtained with the use of the network infrastructure of the Telecommunication center of collective use ``Integrated information-computational network of Irkutsk scientific-educational complex'' (\url{http://net.icc.ru}).
 
@@
 
-The~development of the~infrastructure for Mothur command transformation to Rapidminer dataflow diagram is supported by the project of Irkutsk scientific center of Siberian branch of Russian Academy of sciences, grant No 4.2.
+The~development of the~infrastructure for Mothur command transformation to Rapidminer dataflow diagram is supported by the Irkutsk scientific center of Siberian Branch of Russian Academy of Sciences, grant No 4.2.
 
@@
 \bibitem{abid} T.~Abid, H.~Zarzour.
-Integrating linked open data in geographical information system.
+Integrating linked open data in a geographical information system.
 Procs.
@@
 \bibitem{authoring} E.~Cherkashin, A.~Shigarov, V.~Paramonov, A.~Mikhailov, Digital archives supporting document content inference, Procs.
-of 42-nd International Convention on Information and Communication Technology Electronics and Microelectronics (MIPRO), May, 20–24, 2019.
+of 42-nd International Convention on Information and Communication Technology Electronics and Microelectronics (MIPRO), May 20–24, 2019.
 pp.

### 2022-02-03 paper-aicts-cherkashin.tex
Added subsections for the examples.

@@
 In this section, three examples of application implementation are being presented.
+The first example is an MDA application in the synthesis of a Next-generation sequencing (NGS) investigation environment.
+After the invention of methods of NGS and their introduction in practice of biological systems research, a new direction of molecular genetics is formed, which is referred to as \emph{metagenomics}.
+The method allows biologists to describe a significant number of new groups of organisms at all taxonomic levels.
+A comprehensive review of present sequencing approaches and challenges is presented in~\cite{pere20}.
+
+\subsection{Next-generation sequencing}
+\label{sec:ngs-impl}
+
+The aim of this R\&D is to develop software support of the processes of NGS analysis with techniques and software for a visual representation of the computational process of so-called amplicon analysis so that domain specialists would be able to compose computational pipelines, which are executed on distributed heterogeneous computing resources (clouds).
+The visual models, later, are to be used for organizing the computational process and data store and articular retrieval services.
+
+The tools are based on dataflow representation of various stages of NGS standard operation procedures (SOPs) in Rapidminer studio Java application.
+/Speak about biologists involvement in the process of dataflow diagram construction.
+Analogous soft and integration with them/.
+
+\subsection{Education document authoring}
+\label{sec:doc-impl}
+
+\subsection{Web GIS on the base of a knowledge graph}
+\label{sec:gis-impl}
 
@@
 [we tried to collaborate with software developers using traditional approaches. They are not interested as ]
+
+\section*{Conclusion}
 
@@
 \end{thebibliography}
-\section{Online Resources}
+% \section{Online Resources}
 
-The sources for the viewer are being developed at Github, URL:~\url{https://github.com/De17eon/GRL}.
+% The sources for the viewer are being developed at Github, URL:~\url{https://github.com/De17eon/GRL}.
 

### 2022-02-03 paper-aicts-cherkashin.tex
Final before proof-read.

@@
 \usepackage{hyperref}
+\usepackage{multicol}
 
@@
 \usepackage{pifont}
+\usepackage{xcolor}
 \graphicspath{{pics/}}
+
+\newcommand{\GB}[1]{\colorbox{green}{#1}}
+\newcommand{\BB}[1]{\colorbox{blue}{#1}}
+\newcommand{\RB}[1]{\colorbox{red}{#1}}
 
@@
 Among published global vocabularies, \texttt{DBPedia} has the most value for our projects.
-It is used for concrete object or notion of physical world reference, \emph{e.g.} the term ``passport'' in documents authoring \cite{}, naming (labeling) referenced objects in a user interface with corresponding words in natural language \cite{github-db}.
+It is used for concrete object or notion of physical world reference, \emph{e.g.} the term ``passport'' in documents authoring \cite{}, naming (labeling) referenced objects in a user interface with corresponding words in natural language.
 Another useful resource is \texttt{WordNet} converted in RDF triples.
@@
 A convenient approach of application construction is usage component architectures, where subsystems in runtime are components \emph{providing} interfaces using other components' service available via corresponding interfaces.
-Zope component architecture \cite{zca}, Logtalk, React.js are used to \emph{implement} interfaces.
+Zope component architecture\footnote{On-line book is at \url{https://muthukadan.net/docs/zca.html}}, Logtalk, React.js are used to \emph{implement} interfaces.
 Python allows to wrap whole subsystems, \emph{e.g.}, Elastic search, in imperative paradigm, and Logtalk is chosen for the declarative counterpart.
@@
 The kernel module of applications is \texttt{KG server} (B2) built on the base of ClioPatria \cite{cliopatria} with web-server and Pengines plug-ins.
-ClioPatria is a web-application implemented in SWI-Prolog \cite{swi} being ISO-compliant Prolog implementation.
+ClioPatria is a web-application implemented in SWI-Prolog \cite{cliopatria} being ISO-compliant Prolog implementation.
 KG functions are implemented with SW package, supplied in standard SWI-Prolog distribution.
-Pengines service enables SW web-applications to utilize logical inference on the server-side \cite{pengines} with browser JavaScript environment integration.
+Pengines service enables SW web-applications to utilize logical inference on the server-side \cite{cliopatria} with browser JavaScript environment integration.
 Pengine's knowledge base is extended with the programmer's Prolog modules.
@@
 
-Authoring tool (B1) is used to provide document generation and integration \cite{sibiricon}.
-This tool enables document content and structure inference in the browser utilizing templates and data from other web pages and websites providing LOD capabilities.
+Authoring tool (B1) is used to provide document generation and integration \cite{zont19}.
+This tool enables document content and structure inference in the browser utilizing templates and data from other web pages and websites providing LOD\footnote{Abbreviation of Linked Open Data}с capabilities.
 Programmer implements a view of an object as a web page with an extended version of RDFa.
@@
 The method allows biologists to describe a significant number of new groups of organisms at all taxonomic levels.
-A comprehensive review of present sequencing approaches and challenges is presented in~\cite{pere20}.
 
@@
 
-The aim of this R\&D is to develop software support of the processes of NGS analysis with techniques and software for a visual representation of the computational process of so-called amplicon analysis so that domain specialists would be able to compose computational pipelines, which are executed on distributed heterogeneous computing resources (clouds).
+The aim of this R\&D \cite{zont19, aicts2020} is to develop software support of the processes of NGS analysis with techniques and software for a visual representation of the computational process of so-called amplicon analysis so that domain specialists would be able to compose computational pipelines, which are executed on distributed heterogeneous computing resources (clouds).
 The visual models, later, are to be used for organizing the computational process and data store and articular retrieval services.
 
-The tools are based on dataflow representation of various stages of NGS standard operation procedures (SOPs) in Rapidminer studio Java application.
-/Speak about biologists involvement in the process of dataflow diagram construction.
-Analogous soft and integration with them/.
+The tools are based on dataflow representation of various stages of NGS standard operation procedures (SOPs)  formed out of individual operations of Mothur library in Rapidminer studio Java application.
+To reflect Mothur's operations we implemented source code analysis and transformation software.
+The C++ sources were analyzed by Python syntax analyzer based on regular expressions and a top-level algorithm, generating a \verb|TTL| source of KG representing Mothur modules' interfaces.
+That was a Computation-independent model (CIM) in the implemented MDA.
+CIM transformed into the Platform-independent model (PIM), represented Mothur's modules as dataflow modules, in particular, differencing parameters to input/output dataflow ports and parameters.
+The Platform-specific model (PSM), \emph{i.e.}, models of Java sources of the dataflow blocks, was constructed, in turn, from PSM.
+
+All transformations was carried out with object-oriented (OO) Logtalk represented set of modules that logically infer new data descriptions of the target objects.
+Thanks to OO, the transformation have two level of abstraction in its representation.
+The higher one is a network of objects sending messages to each other via their interfaces.
+The lower is the set of rules analyzing and producing new conclusions on generated object properties.
+
+The PSM objects must have an order in their description as they represent a models of source code to be generated.
+The order is organized with blocks of \verb|assert|ed statement and subobjects sequences referencing KG elements.
+Blocks can be appended and pretended with new elements.
+At the stage of rendering sources, the three of blocks are traversed in-depth.
+This leverages known \verb|llvmlite| open-source approach of code generation.
+Figure~\ref{fig:block-seq} shows the main idea behind PSM object synthesis.
+\begin{figure}
+  \centering
+\begin{multicols}{2}
+\begin{minted}[fontsize=\scriptsize]{logtalk}
+:- object(class, specializes(code_block),
+   imports([named])). % Category of named entities
+:- public([classlist/1, methods/1, attributes/1]).
+% . . . . . . . . . . . . . .
+renderitem(Item, Result):-      % proceed with default
+    ^^renderitem(Item, Result). % rendering
+render(Result):-         % Source generator
+    ^^render(Name),      % implemented in a category
+    ( ::item(classlist(List)) ->
+     % . . . . . . . . . . .
+        [Name]) ),
+    ( ::item(attributes(Attributes))->
+     % . . . . . . . . . . .
+        [DefAttrList]),
+      Attributes::items(InstanceAttrs),
+      findall(S, ( % initialize attributes
+         % . . . . . . . . .
+         ), AttrAssigns),
+        root::unindent,
+        AttrList=[ConstructorDef|AttrAssigns];
+         % . . . . . . . . .
+        AttrList=[ConstructorDef, Pass] ),
+    ( ::item(methods(Methods))-> % If any ...
+      Methods::render(MethodList);
+      MethodList=[] ),
+    lists::append(AttrList,MethodList,StringList),
+    root::unindent, Result=[Signature|StringList].
+:- end_object.
+\end{minted}
+      \includegraphics[width=0.7\linewidth]{code_block_class.pdf}
+\end{multicols}
+  \caption{Procedure of dataflow module source code model synthesis}
+  \label{fig:block-seq}
+\end{figure}
+This example of the application of KG utilizes strongly the property of data representation flexibility for description of the software object model data as well as one central point of data store and access.
+The disadvantageous properties of RDF, namely, the weak expressiveness with triples, was neglected by predicate and object encapsulation.
 
@@
 
+This project \cite{authoring} is aimed at construction of an environment for document authoring out of user editable template elements, database query results, LOD-compliant web-pages, and other document data.
+Target document is a web page which content is inferred while loaded by the browser.
+
+All document elements are either a context structures of templates filled with other elements.
+A context is a RDF hierarchical representation of a document data structure.
+It provides data to be filled in a current rendered template.
+Document is a main template.
+All elements are SW resources.
+Templates can contain JavaScript subroutines for implementing special service, for example, table of contents out of CSS \verb|class| labeled subsections of the generated document.
+
+All elements are stored either in a KG as triples, or as LOD-compliant web-pages, and according to SW consideration all of them form a distributed KG.
+Data access differentiation by security model is implemented programmatically as the web pages disallowing access to the restricted information.
+SW has special predicates for description a possibility and location of a private data sub-KG in the open part.
+
+In Figure~\ref{fig:auth-code}, we presented our small syntactic extension of RDFa for referring template parts.
+A namespace \textbf{taa} was added to define a markup interpreted as XHTML tree manipulation in the image and likeliness of open-source ZPT\footnote{Abbreviation of Zope Page Template}.
+Arguments of \verb|taa:content| are subcontexts (\verb|title| and \verb|neg-UMK|) of a current template under rendering.
+\begin{figure}
+  \centering
+\begin{minted}[escapeinside=||,fontsize=\small]{text}
+<html lang="ru" xmlns=http://www.w3.org/1999/xhtml
+|\GB{xmlns:taa}|=http://irnok.net/engine/rdfa-manipulation
+xml:lang="ru" metal:define-macro="page">
+<head> . . . . </head>
+<body prefix="rdf: http://www.w3.org/1999/...-ns# foaf: http://xmlns.com/foaf/...
+imei: imei.html# course: https://irnok.net/college/plan/01..16-...\
+% D0\%BA_PB-SM.plm.xml.xlsx-....2.3.1.html#"  resource="#post"
+typeof="schema:CreativeWork sioc:Post prov:Entity">
+<!-- The application control panel -->
+
+<main lang="ru" resource="#annotation" typeof="oa:Annotation" id="main-doc-cnt">
+<div property="oa:hasTarget" resource="#course-work-prog"></div>
+<article property="oa:hasBody" typeof="foaf:Document curr:WorkingProgram"
+         resource="#course-work-program" id="main-document">
+  <div |\GB{taa:content}|="imei:title-page"></div>
+  <div |\GB{taa:content}|="imei:neg-UMK"></div>
+  <section id="TOC" class="break-after"> <h2>Table of Contents</h2>
+    <div id="tableOfContents"></div>
+  </section>
+  <section id="course-description" resource="#description"
+           property="schema:hasPart" typeof="schema:CreativeWork">
+    <div property="schema:hasPart" resource="#purpose"
+         typeof="dc:Text cnt:ContentAsText" >
+      <div property="cnt:chars" datatype="xsd:string">
+        <h2 property="dc:title" datatype="xsd:string">
+           Aims and objectives of the discipline (module)</h2>
+        <p>The aim of teaching the discipline ...</p>
+      </div>
+   </div>
+  . . . . . . . .
+\end{minted}
+  \caption{Part of a template representation}
+  \label{fig:auth-code}
+\end{figure}
+
+For document markup (document data representation), the following vocabularies are used:
+\begin{itemize}
+  \item Friend-of-a-friend (\textbf{foaf}), agent information: individuals, legal entities, program agents;
+  \item Provenance (\textbf{prov}), references between documents;
+  \item Dublin Core (\textbf{dc}), web page annotation mark up;
+  \item DBPedia resource (\textbf{dbr}), references to instant objects and classes;
+  \item Schema.org (\textbf{schema}), Google, Yandex, Yahoo, \emph{etc}. searchable objects, structural elements;
+  \item The Bibliographic Ontology (\textbf{bibo}), literature reference mark up.
+\end{itemize}
+
+The general logical structure (sections, subsections, \emph{etc.}) are represented with Open Annotation (\textbf{oa}) vocabulary originally used for annotation of already existing content.
+Using \textbf{oa}, the generated document is converted to PDF and other formats.
+HTML5 default tag structures are used for visual sectioning and style definition.
+
+The \textbf{oa} vocabulary also used to define parts of document, which can be interactively edited after the document have been already rendered.
+The obtained changes are fixed in a commit as new resources, and the containing document obtains new IRI fixing all the set of resource changes.
+
 \subsection{Web GIS on the base of a knowledge graph}
@@
 
-\section{Discussion}
+Project \cite{gisviewer} devoted to investigate capabilities of SW and KG technologies in representation spatially distributed data (GIS\footnote{Abbreviation of Geographical Information System}-data).
+The fault data \cite{lunina} are chosen as subject for representation and publishing in this investigation.
+In the geology, researchers accumulate data obtained after event observations, \emph{e.g.}, earthquakes, landslides, by analyzing remote sensing data and results of field works.
+The obtained data are processed and interpreted, resulting in setting new attributes to a fault or refinement of their values.
+According to the techniques of geological research, additional information are associated with attributes, clarifying their values.
+Such clarification comprises so-called precision characteristics, measurement conditions, reliability assessment, and paper references, where fault data were published.
+
+GIS represents spatial data in semantic-defined layers.
+For each object of a GIS layer, one can associate a set of attribute values of auxiliary data.
+The set of the attributes are the same for each object of the layer, regardless of whether the attribute value is defined for an object or not.
+Empty values are represented as ``\texttt{null}''s.
+In the case of geological exploration, when a lot of attributes are undefined, this approach leads to sparse filled tables.
+This, in turn, requires data modification and analysis algorithms to utilize additional data checking stages when using standard relation operations (SELECT, UPDATE, DELETE).
+
+Web publication applications, as information systems, are to implement filtering functions, differentiating the value attributes and their metadata.
+Screen widgets label names either defined in application configuration or figured out from the attribute names.
+Lack of vocabulary formal domain definition forces developer to spend more efforts for the user interface implementation.
+
+In this project, we constructed a Web-application using contemporary Web and SW technologies for rendering a fault map, which visual content is controlled by data filters implemented as SPARQL queries to KG representing fault data.
+For the purpose, a data conversion from the tabular form to KG was carried out, a CliopPatria server for data storage was installed, and data viewers for the triple sets for a common subject (a fault) has been implemented.
+In Figure~\ref{fig:gis-ex} an example of the filtering and a selected object data viewing form is presented.
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{faults-leaflet-doc.png}
+  % \caption{Viewer window showing description of a fault}
+  \caption{Example of GIS rendering of KG data}
+  \label{fig:gis-ex}
+\end{figure}
+
+The proposed technology and software allows one to construct Web-GIS systems for research communities, as they support constant data accumulation, aggregation and analysis thanks to the properties of knowledge graph (KG) data storage and processing.
+
+\section{Discussion and Conclusion}
 \label{sec:disc}
 
-[we tried to collaborate with software developers using traditional approaches. They are not interested as ]
+Standardized forms of basic concepts of data representation and processing in the filed of Sematic Web (SW) research were considered in this paper.
+A special attention has been paid to reviewing Knowledge Graph (KG) data storage and processing in scientific data processing software.
 
-\section*{Conclusion}
+We did not consider analogous R\&D as it was already reviewed in the corresponding our publications.
+Here we just present a review of obtained SW and KG technologies application experience.
+The main one is the powerful capabilities of application integration, where all the set of implemented software, in fact, form a distributed environment of data processing.
+The main requirement is that the data representation at storage services and in the protocol implementations should obey SW and KG standards.
+
+The main advantages of these technologies for the software developer are semantic markup and generic properties of KG, especially a possibility to postpone formal data structure definition as such declarative data are easily can be processed with a rule-based inference engine.
+Scientific community results in domain description in standardized vocabularies (ontologies), user communities aggregating and formalizing data on real-world objects form a productive background for problem space description and,  in its turn, a common basis for application integration, which result in a cumulative effect of aggregation of developed software in a complex distribute data processing system and environments.
+
+Further development of the technologies include the following directions of a great interest:
+\begin{itemize}
+\item Fibered representation of KG to support consistency and completeness check by reflecting A-Box entities and their relations to template domain models;
+\item Investigate expressive capabilities of Logtalk programming language as an instrument of knowledge (rules) representation for KG processing modules;
+\item Adding extended document analysis services \cite{shigpaper} to the environment (section~\ref{sec:architecture}), which are developed in our research group;
+\item Improve integration techniques with metamodel presentation of the execution environment.
+\end{itemize}
 
@@
 
-\bibitem{tbl} T.B.-L. Scientific American ...
+\bibitem{tbl} Tim Berners-Lee, James Hendler, and Ora Lassila.
+The semantic web: A new form of web contentthat is meaningful to computers will unleash a revolution of new possibilities.
+ScientificAmeri-can.com, 05 2001.
 
@@
 
+\bibitem{aicts2020} Evgeniy Cherkashin, Alexey Shigarov.
+Construction techniques of Baikal microbiome research information-computational environment.
+Proceedings for 1st International Workshop on Advanced Information and Computation Technologies and Systems.
+Irkutsk, Russia, December 7-11, 2020.
+CEUR-WS.org, online \url{http://ceur-ws.org/Vol-2858/paper2.pdf} (access date: 12-Dec-2021)
+
 \bibitem{authoring} E.~Cherkashin, A.~Shigarov, V.~Paramonov, A.~Mikhailov, Digital archives supporting document content inference, Procs.
@@
 Irkutsk, Russia, September 14, 2021.
-CEUR-WS.org, online \url{http://ceur-ws.org/Vol-2984/paper8.pdf}
+CEUR-WS.org, online \url{http://ceur-ws.org/Vol-2984/paper8.pdf} (access date: 12-Dec-2021)
 
-\bibitem{cliopatria}
-\bibitem{sibircon} Logical objects
+\bibitem{cliopatria} J. Wielemaker, W. Beek, M. Hildebrand, J. Ossenbruggen, ``ClioPatria: A
+  SWI-Prolog infrastructure for the Semantic Web,'' Semantic Web.
+  Vol. 7(5). P. 529--541 (2016) \doi{10.3233/SW-150191}
+
+\bibitem{shigpaper} Dorodnykh N.O., Yurin A.Y., Shigarov A.O.
+(2020) Conceptual Model Engineering for Industrial Safety Inspection Based on Spreadsheet Data Analysis.
+In: Simian D., Stoica L.
+(eds) Modelling and Development of Intelligent Systems.
+MDIS 2019.
+Communications in Computer and Information Science, vol 1126.
+Springer, Cham.
+\doi{10.1007/978-3-030-39237-6_4}
 

### 2022-02-03 paper-aicts-cherkashin.tex
Balance to our and other references

@@
 
-Since 2000, the Semantic Web (SW) technologies have become actively used in describing software domains, representing semantic relations between objects.
-For some domains on the base of the SW description, first-order knowledge theories are constructed \cite{}, but in the practical aspect of the SW utilization, the solving data representation and software integration aspects prevail.
-Some tools for automation of conceptual model design and model conversion were proposed \cite{}.
+Since 2001 \cite{tbl}, the Semantic Web (SW) technologies have become actively used in describing software domains, representing semantic relations between objects.
+For some domains on the base of the SW description, first-order knowledge theories are constructed, but in the practical aspect of the SW utilization, the solving data representation and software integration aspects prevail.
+Some tools for automation of conceptual model design and model conversion were proposed \cite{hogan}.
 The R\&D at present are carried on in developing applications, improving and standardizing vocabularies (conceptual domain models), devising tools for solving particular problems.
@@
 Among published global vocabularies, \texttt{DBPedia} has the most value for our projects.
-It is used for concrete object or notion of physical world reference, \emph{e.g.} the term ``passport'' in documents authoring \cite{}, naming (labeling) referenced objects in a user interface with corresponding words in natural language.
+It is used for concrete object or notion of physical world reference, \emph{e.g.} the term ``passport'' in documents authoring \cite{zont19}, naming (labeling) referenced objects in a user interface with corresponding words in natural language.
 Another useful resource is \texttt{WordNet} converted in RDF triples.
@@
 The kernel module of applications is \texttt{KG server} (B2) built on the base of ClioPatria \cite{cliopatria} with web-server and Pengines plug-ins.
-ClioPatria is a web-application implemented in SWI-Prolog \cite{cliopatria} being ISO-compliant Prolog implementation.
+ClioPatria is a web-application implemented in SWI-Prolog \cite{swi} being ISO-compliant Prolog implementation.
 KG functions are implemented with SW package, supplied in standard SWI-Prolog distribution.
-Pengines service enables SW web-applications to utilize logical inference on the server-side \cite{cliopatria} with browser JavaScript environment integration.
+Pengines service enables SW web-applications to utilize logical inference on the server-side \cite{pengines} with browser JavaScript environment integration.
 Pengine's knowledge base is extended with the programmer's Prolog modules.
@@
 Authoring tool (B1) is used to provide document generation and integration \cite{zont19}.
-This tool enables document content and structure inference in the browser utilizing templates and data from other web pages and websites providing LOD\footnote{Abbreviation of Linked Open Data}с capabilities.
+This tool enables document content and structure inference in the browser utilizing templates and data from other web pages and websites providing LOD\footnote{Abbreviation of Linked Open Data} \cite{lod} capabilities.
 Programmer implements a view of an object as a web page with an extended version of RDFa.
@@
 Project \cite{gisviewer} devoted to investigate capabilities of SW and KG technologies in representation spatially distributed data (GIS\footnote{Abbreviation of Geographical Information System}-data).
-The fault data \cite{lunina} are chosen as subject for representation and publishing in this investigation.
+The fault data \cite{lunina,afs} are chosen as subject for representation and publishing in this investigation.
 In the geology, researchers accumulate data obtained after event observations, \emph{e.g.}, earthquakes, landslides, by analyzing remote sensing data and results of field works.
@@
 We did not consider analogous R\&D as it was already reviewed in the corresponding our publications.
-Here we just present a review of obtained SW and KG technologies application experience.
+Another point of our R\&D is a strategy for improving existing technologies, data and software with our tools, such as MDA.
+We present a review of obtained SW and KG technologies application experience.
 The main one is the powerful capabilities of application integration, where all the set of implemented software, in fact, form a distributed environment of data processing.
@@
 \bibitem{tbl} Tim Berners-Lee, James Hendler, and Ora Lassila.
-The semantic web: A new form of web contentthat is meaningful to computers will unleash a revolution of new possibilities.
-ScientificAmeri-can.com, 05 2001.
+The semantic web: A new form of web content that is meaningful to computers will unleash a revolution of new possibilities.
+Scientific American, Vol.~05, 2001.
+
+\bibitem{hogan} A.~Hogan, E.~Blomqvist, M.~Cochez, C.~D’Amato \emph{et al}. Knowledge Graphs, 2020. \url{https://arxiv.org/abs/2003.02320v5}
 
@@
 
-\bibitem{hogan} A.~Hogan, E.~Blomqvist, M.~Cochez, C.~D’Amato \emph{et al}. Knowledge Graphs, 2020. \url{https://arxiv.org/abs/2003.02320v5}
-
 \bibitem{afs} A.~A.~Gladkov, O.~V.~Lunina. Cartographic service ``Activetectonics''. \url{http://activetectonics.ru/} (access date: 20-Sep-2021)
 
-\bibitem{foss} M.~Leidig, R.~Teeuw.
+%\bibitem{foss} M.~Leidig, R.~Teeuw.
 Free software: A review, in the context of disaster management.
@@
 
-\bibitem{lgd} C.~Stadler, J.~Lehmann, K.~Höffner, S.~Auer. LinkedGeoData: A core for a web of spatial open data. Semantic Web 3 (2012) 333–354. \doi{10.3233/SW-2011-0052}
+%\bibitem{lgd} C.~Stadler, J.~Lehmann, K.~Höffner, S.~Auer. LinkedGeoData: A core for a web of spatial open data. Semantic Web 3 (2012) 333–354. \doi{10.3233/SW-2011-0052}
 
-\bibitem{geolink} M.~Cheatham, A.~Krisnadhi, R.~Amini, P.~Hitzler, \emph{et al}. The GeoLink knowledge graph, Big Earth Data, 2:2 (2018), pp.~131-143. \doi{10.1080/20964471.2018.1469291}
+%\bibitem{geolink} M.~Cheatham, A.~Krisnadhi, R.~Amini, P.~Hitzler, \emph{et al}. The GeoLink knowledge graph, Big Earth Data, 2:2 (2018), pp.~131-143. \doi{10.1080/20964471.2018.1469291}
 
-\bibitem{abid} T.~Abid, H.~Zarzour.
+%\bibitem{abid} T.~Abid, H.~Zarzour.
 Integrating linked open data in a geographical information system.
@@
 
-\bibitem{iwaniak1} A.~Iwaniak, I.~Kaczmarek, M.~Strzelecki, J.~Lukowicz, P.~Jankowski.
+%\bibitem{iwaniak1} A.~Iwaniak, I.~Kaczmarek, M.~Strzelecki, J.~Lukowicz, P.~Jankowski.
 Enriching and improving the quality of linked data with GIS.
@@
 
-\bibitem{iwaniak17} A.~Iwaniak, M.~Leszczuk, M.~Strzelecki, F.~Harvey, I.~Kaczmarek.
+%\bibitem{iwaniak17} A.~Iwaniak, M.~Leszczuk, M.~Strzelecki, F.~Harvey, I.~Kaczmarek.
 A Novel Approach for Publishing Linked Open Geodata from National Registries with the Use of Semantically Annotated Context Dependent Web Pages.
@@
 
+\bibitem{pengines} {Torbj{\"{o}}rn Lager, Jan Wielemaker.
+Pengines: Web Logic Programming Made Easy.
+Theory and Practice of Logic Programming.
+Vol.~14, No.~4-5.
+pp.~539--552.
+2014.
+\doi{10.1017/S1471068414000192}
+
+\bibitem{swi} Jan Wielemaker, Tom Schrijvers, Markus  Triska, Torbj\"o{}rn} Lager. SWI-Prolog. Theory and Practice of Logic Programming, Vol.~12, No.~1-2, pp.~67--96, ISSN~1471-0684.
+
 \end{thebibliography}

### 2022-02-04 paper-aicts-cherkashin.tex
Grammarly check done

@@
 %%
-%% One can fix some overfulls
+%% One can fix some overfills
 \sloppy
@@
 Since 2001 \cite{tbl}, the Semantic Web (SW) technologies have become actively used in describing software domains, representing semantic relations between objects.
-For some domains on the base of the SW description, first-order knowledge theories are constructed, but in the practical aspect of the SW utilization, the solving data representation and software integration aspects prevail.
+First-order knowledge theories are constructed for some domains based on the SW description, but in the practical aspect of the SW utilization, the solving data representation problems and software integration aspects prevail.
 Some tools for automation of conceptual model design and model conversion were proposed \cite{hogan}.
@@
 For example \verb|integer| literal \verb|42^^<http://www.w3.org/2001/XMLSchema#integer>| can be reduced to \verb|42^^xs:integer|, where \verb|xs| is namespace prefix of term \verb|integer|.
-Most KG representation formats allow one to define a default namespace, which is used without prefix.
+Most KG representation formats allow one to define a default namespace used without prefix.
 Almost any conceptual model of a domain represented as KG comprises entities taken from multiple vocabularies, and namespaces specify the origin of a term.
@@
 
-KGs and their subgraphs are programmatically processed using libraries developed for all papular programming environments.
+KGs and subgraphs are programmatically processed using libraries developed for all papular programming environments.
 KG is represented with a triple or quadruple set object.
@@
 
-Authoring tool (B1) is used to provide document generation and integration \cite{zont19}.
+Authoring tool (B1) is used to provide document generation, and integration \cite{zont19}.
 This tool enables document content and structure inference in the browser utilizing templates and data from other web pages and websites providing LOD\footnote{Abbreviation of Linked Open Data} \cite{lod} capabilities.
@@
 In this section, three examples of application implementation are being presented.
-The first example is an MDA application in the synthesis of a Next-generation sequencing (NGS) investigation environment.
-After the invention of methods of NGS and their introduction in practice of biological systems research, a new direction of molecular genetics is formed, which is referred to as \emph{metagenomics}.
-The method allows biologists to describe a significant number of new groups of organisms at all taxonomic levels.
+The first example is an MDA application in synthesizing a Next-generation sequencing (NGS) investigation environment.
+After the invention of methods of NGS and their introduction in the practice of biological systems research, a new direction of molecular genetics is formed, which is referred to as \emph{metagenomics}.
+The method allows biologists to describe many new groups of organisms at all taxonomic levels.
 
@@
 
-The aim of this R\&D \cite{zont19, aicts2020} is to develop software support of the processes of NGS analysis with techniques and software for a visual representation of the computational process of so-called amplicon analysis so that domain specialists would be able to compose computational pipelines, which are executed on distributed heterogeneous computing resources (clouds).
-The visual models, later, are to be used for organizing the computational process and data store and articular retrieval services.
+The aim of this R\&D \cite{zont19, aicts2020} is to develop software support of the processes of NGS analysis with techniques and software for a visual representation of the computational process for so-called amplicon analysis so that domain specialists would be able to compose computational pipelines, which are executed on distributed heterogeneous computing resources (clouds).
+Later, the visual models are used to organize the computational process, data store, and particular NGS data retrieval services.
 
 The tools are based on dataflow representation of various stages of NGS standard operation procedures (SOPs)  formed out of individual operations of Mothur library in Rapidminer studio Java application.
-To reflect Mothur's operations we implemented source code analysis and transformation software.
-The C++ sources were analyzed by Python syntax analyzer based on regular expressions and a top-level algorithm, generating a \verb|TTL| source of KG representing Mothur modules' interfaces.
+To reflect Mothur's operations, we implemented source code analysis and transformation software.
+The C++ sources were analyzed by a Python syntax analyzer based on regular expressions and a top-level algorithm, generating a \verb|TTL| source of KG representing Mothur modules' interfaces.
 That was a Computation-independent model (CIM) in the implemented MDA.
-CIM transformed into the Platform-independent model (PIM), represented Mothur's modules as dataflow modules, in particular, differencing parameters to input/output dataflow ports and parameters.
+CIM transformed into the Platform-independent model (PIM), represented Mothur's modules as dataflow modules, mainly differencing parameters to input/output dataflow ports and parameters.
 The Platform-specific model (PSM), \emph{i.e.}, models of Java sources of the dataflow blocks, was constructed, in turn, from PSM.
 
-All transformations was carried out with object-oriented (OO) Logtalk represented set of modules that logically infer new data descriptions of the target objects.
-Thanks to OO, the transformation have two level of abstraction in its representation.
-The higher one is a network of objects sending messages to each other via their interfaces.
-The lower is the set of rules analyzing and producing new conclusions on generated object properties.
+All transformations were carried out with an object-oriented (OO) Logtalk programming environment representing modules that logically infer new data descriptions of the target objects.
+Thanks to OO, the transformation has two levels of abstraction in its representation.
+The higher one is a network of objects sending messages via their interfaces.
+At the lower one, rules analyze and produce new conclusions on generated object properties.
 
-The PSM objects must have an order in their description as they represent a models of source code to be generated.
+The PSM objects must have an order in their description as they represent a model of source code to be generated.
 The order is organized with blocks of \verb|assert|ed statement and subobjects sequences referencing KG elements.
 Blocks can be appended and pretended with new elements.
-At the stage of rendering sources, the three of blocks are traversed in-depth.
-This leverages known \verb|llvmlite| open-source approach of code generation.
+At the rendering sources stage, the three blocks are traversed in-depth.
+It leverages the known \verb|llvmlite| source code generation approach.
 Figure~\ref{fig:block-seq} shows the main idea behind PSM object synthesis.
@@
 
-This project \cite{authoring} is aimed at construction of an environment for document authoring out of user editable template elements, database query results, LOD-compliant web-pages, and other document data.
-Target document is a web page which content is inferred while loaded by the browser.
+This project \cite{authoring} is aimed at the construction of an environment for document authoring out of user-editable template elements, database query results, LOD-compliant web pages, and other document data.
+The target document is a web page where content is inferred while loaded by the browser.
 
-All document elements are either a context structures of templates filled with other elements.
-A context is a RDF hierarchical representation of a document data structure.
-It provides data to be filled in a current rendered template.
-Document is a main template.
+All document elements are either context structures or templates filled with other elements.
+A context is an RDF hierarchical representation of a document data structure, providing data to be filled in a current rendered template.
+The document is rendered from its main template.
 All elements are SW resources.
-Templates can contain JavaScript subroutines for implementing special service, for example, table of contents out of CSS \verb|class| labeled subsections of the generated document.
+Templates can contain JavaScript subroutines for implementing special services, for example, table of contents out of CSS \verb|class| labeled subsections of the generated document.
 
-All elements are stored either in a KG as triples, or as LOD-compliant web-pages, and according to SW consideration all of them form a distributed KG.
-Data access differentiation by security model is implemented programmatically as the web pages disallowing access to the restricted information.
-SW has special predicates for description a possibility and location of a private data sub-KG in the open part.
+All elements are stored in a KG as triples and as LOD-compliant web pages, and according to SW considerations, all of them form a distributed KG.
+Data access differentiation by security model is implemented programmatically as the web pages disallow access to the restricted information.
+SW has special predicates for describing a possibility and location of a private data sub-KG in the open part.
 
@@
 The general logical structure (sections, subsections, \emph{etc.}) are represented with Open Annotation (\textbf{oa}) vocabulary originally used for annotation of already existing content.
-Using \textbf{oa}, the generated document is converted to PDF and other formats.
+The generated document conversion into PDF and other formats is carried out under the control of \textbf{oa}.
 HTML5 default tag structures are used for visual sectioning and style definition.
 
-The \textbf{oa} vocabulary also used to define parts of document, which can be interactively edited after the document have been already rendered.
+The \textbf{oa} vocabulary is also used to define parts of a document, which allow interactive editing after the document has been rendered.
 The obtained changes are fixed in a commit as new resources, and the containing document obtains new IRI fixing all the set of resource changes.
@@
 
-Project \cite{gisviewer} devoted to investigate capabilities of SW and KG technologies in representation spatially distributed data (GIS\footnote{Abbreviation of Geographical Information System}-data).
-The fault data \cite{lunina,afs} are chosen as subject for representation and publishing in this investigation.
-In the geology, researchers accumulate data obtained after event observations, \emph{e.g.}, earthquakes, landslides, by analyzing remote sensing data and results of field works.
-The obtained data are processed and interpreted, resulting in setting new attributes to a fault or refinement of their values.
-According to the techniques of geological research, additional information are associated with attributes, clarifying their values.
-Such clarification comprises so-called precision characteristics, measurement conditions, reliability assessment, and paper references, where fault data were published.
+Project \cite{gisviewer} devoted to investigating capabilities of SW and KG technologies in representation spatially distributed data (GIS\footnote{Abbreviation of Geographical Information System}-data).
+The fault data \cite{lunina,afs} are chosen as a subject for representation and publishing in this investigation.
+In geology, researchers accumulate data obtained after event observations, \emph{e.g.}, earthquakes, landslides, by analyzing remote sensing data and results of field works.
+The obtained data are processed and interpreted, setting new attributes to a fault or refinement of their values.
+According to the geological research techniques, additional information is associated with attributes, clarifying their values.
+Such clarification comprises precision characteristics, measurement conditions, reliability assessment, paper references, and published fault data.
 
@@
 For each object of a GIS layer, one can associate a set of attribute values of auxiliary data.
-The set of the attributes are the same for each object of the layer, regardless of whether the attribute value is defined for an object or not.
+The attribute set is the same for each object of a layer, regardless the attribute's value is defined or not.
 Empty values are represented as ``\texttt{null}''s.
-In the case of geological exploration, when a lot of attributes are undefined, this approach leads to sparse filled tables.
-This, in turn, requires data modification and analysis algorithms to utilize additional data checking stages when using standard relation operations (SELECT, UPDATE, DELETE).
+In the case of geological exploration, when many attributes are undefined, this approach leads to sparsely filled tables.
+In this case, structure modification and data analysis algorithms are required to utilize additional data checking stages when using standard relation operations (SELECT, UPDATE, DELETE).
 
-Web publication applications, as information systems, are to implement filtering functions, differentiating the value attributes and their metadata.
+Web publication applications, like information systems, implement filtering functions, differentiating the value attributes and their metadata.
 Screen widgets label names either defined in application configuration or figured out from the attribute names.
-Lack of vocabulary formal domain definition forces developer to spend more efforts for the user interface implementation.
+Lack of vocabulary formal domain definition forces developers to spend more effort on user interface implementation.
 
-In this project, we constructed a Web-application using contemporary Web and SW technologies for rendering a fault map, which visual content is controlled by data filters implemented as SPARQL queries to KG representing fault data.
-For the purpose, a data conversion from the tabular form to KG was carried out, a CliopPatria server for data storage was installed, and data viewers for the triple sets for a common subject (a fault) has been implemented.
+In this project, we constructed a Web-application using contemporary Web and SW technologies for rendering a fault map, in which visual content is controlled by data filters implemented as SPARQL queries to KG representing fault data.
+For this purpose, data conversion from the tabular form to KG was carried out, a CliopPatria server for data storage was installed, and data viewers for the triple sets for a typical subject (a fault) have been implemented.
 In Figure~\ref{fig:gis-ex} an example of the filtering and a selected object data viewing form is presented.
@@
 
-The proposed technology and software allows one to construct Web-GIS systems for research communities, as they support constant data accumulation, aggregation and analysis thanks to the properties of knowledge graph (KG) data storage and processing.
+The proposed technology and software allow one to construct Web-GIS systems for research communities, as they support constant data accumulation, aggregation, and analysis thanks to the properties of knowledge graph (KG) data storage and processing.
 
@@
 
-Standardized forms of basic concepts of data representation and processing in the filed of Sematic Web (SW) research were considered in this paper.
-A special attention has been paid to reviewing Knowledge Graph (KG) data storage and processing in scientific data processing software.
+Standardized forms of basic concepts of data representation and processing in Semantic Web (SW) research were considered in this paper.
+Special attention has been paid to reviewing Knowledge Graph (KG) data storage and processing in the scientific data processing software.
 
-We did not consider analogous R\&D as it was already reviewed in the corresponding our publications.
-Another point of our R\&D is a strategy for improving existing technologies, data and software with our tools, such as MDA.
+We did not consider analogous R\&D as it was already reviewed in the corresponding publications.
+Another point of our R\&D is improving existing technologies, data, and software with our tools, such as MDA.
 We present a review of obtained SW and KG technologies application experience.
-The main one is the powerful capabilities of application integration, where all the set of implemented software, in fact, form a distributed environment of data processing.
+The main one is the powerful application integration capabilities, where all the set of implemented software forms a distributed environment of data processing.
 The main requirement is that the data representation at storage services and in the protocol implementations should obey SW and KG standards.
 
-The main advantages of these technologies for the software developer are semantic markup and generic properties of KG, especially a possibility to postpone formal data structure definition as such declarative data are easily can be processed with a rule-based inference engine.
-Scientific community results in domain description in standardized vocabularies (ontologies), user communities aggregating and formalizing data on real-world objects form a productive background for problem space description and,  in its turn, a common basis for application integration, which result in a cumulative effect of aggregation of developed software in a complex distribute data processing system and environments.
+The main advantages of these technologies for the software developer are semantic markup and generic properties of KG, especially a possibility to postpone formal data structure definition as such declarative set-based data can be processed with a rule-based inference engine.
+Scientific community results in domain description in standardized vocabularies (ontologies), user communities aggregating and formalizing data on real-world objects form a productive background for problem space description and,  in its turn, a common basis for application integration, which results in a cumulative effect of aggregation of developed software in a complex distributed data processing system and environments.
 
-Further development of the technologies include the following directions of a great interest:
+Further development of the technologies includes the following directions of great interest:
 \begin{itemize}

### 2022-02-04 paper-aicts-cherkashin.tex
Some corrections before references renumbering and formatting

@@
 \end{enumerate}
-The assets  (\ref{kg1}--\ref{kg2}) form knowledge graph technological aspect of the SW that gained a particular interest within R\&D in the last decade.
+The assets  \ref{kg1}--\ref{kg2} form knowledge graph technological aspect of the SW that gained a particular interest within R\&D in the last decade.
 
@@
 Literals are values of basic types (\emph{e.g.}, number, string, date, time) usually accompanied with metadata (type designator or a language tag).
-Types are defined in a vocabulary \texttt{<\url{http://www.w3.org/2001/XMLSchema}\#>}, having well-known namespace abbreviation \ns{xs}\footnote{Hereafter, all well-known namespace names will be given after its IRI in brackets with a bold monospace font.}.
+Types are defined in a vocabulary \texttt{<\url{http://www.w3.org/2001/XMLSchema}\#>}, having well-known namespace abbreviation \ns{xs}\footnote{Hereafter, all well-known namespace names will be written with a bold monospace font.}.
 A language tag defines string values attribute, affiliating them to a cultural context.
@@
 Namespaces are convenient instruments in the triple definition.
-For example \verb|integer| literal \verb|42^^<http://www.w3.org/2001/XMLSchema#integer>| can be reduced to \verb|42^^xs:integer|, where \verb|xs| is namespace prefix of term \verb|integer|.
+For example \verb|integer| literal \verb|42^^<http://www.w3.org/2001/XMLSchema#integer>| can be reduced to \verb|42^^xs:integer|, where \verb|xs| is namespace prefix of a vocabulary having term \verb|integer|.
 Most KG representation formats allow one to define a default namespace used without prefix.
@@
 
-\section{Discussion and Conclusion}
+\section*{Conclusion}
 \label{sec:disc}

### 2022-02-04 paper-aicts-cherkashin.tex
Sorting references
Release

@@
     colorlinks=true,       % false: boxed links; true: colored links
-    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
-    citecolor=green,        % color of links to bibliography
-    filecolor=magenta,      % color of file links
-    urlcolor=blue           % color of external links
+    linkcolor=black,          % color of internal links (change box color with linkbordercolor)
+    citecolor=black,        % color of links to bibliography
+    filecolor=black,      % color of file links
+    urlcolor=black           % color of external links
   }
-
+\urlstyle{tt}
 \usepackage{pifont}
@@
 %% This command is for the conference information
-\conference{2nd International Workshop on Advanced Information and Computation Technologies and Systems, December XX--XX, 2021, Irkutsk, Russia}
+\conference{2nd International Workshop on Advanced Information and Computation Technologies and Systems, December 6--10, 2021, Irkutsk, Russia}
 
@@
   Semantic Web technologies and their standards are becoming productive assets for software development and integration on various design and implementation levels.
-These include OSI level 6 data representation, database and file storage formats, the source data for text indexes, document authoring, hypertext markup engines, a representation for abstract software models, \emph{etc.}  They are also smoothly combined with the development and data analysis tools, giving rise to a new common-ground environment for integrating distributed applications on a semantic level.
+These include OSI level 6 data representation, database and file storage formats, the source data for text indexes, document authoring, hypertext markup engines, a representation for abstract software models, \emph{etc.}  They are also smoothly combined with the development and data analysis tools, giving rise to an environment for integrating distributed applications on a semantic level.
 
@@
 In Russia, domain model designers generally prefer to develop local vocabularies instead of spending time for the routine work of Internet surfing for existing conceptual models and their enrichment.
-It is due to underdevelopment of the metadata resources (metavocabularies) due to lack of a domain classification system as compared, \emph{e.g.} to Universal Decimal Classification (UDC) used in the bibliography.
-Vocabularies also can have common parts, describing the same or similar entities with different terms and relations, \emph{e.g.} semantically more adapted to a specific problem-solving.
+It is due to underdevelopment of the metadata resources (metavocabularies) due to lack of a domain classification system as compared, \emph{e.g.}, to Universal Decimal Classification (UDC) used in the bibliography.
+Vocabularies also can have common parts, describing the same or similar entities with different terms and relations, in particular, semantically more adapted to a specific problem-solving.
 
@@
 Each entity described in SW has its unique global IRI, its virtual or real internet address.
-Document entities' resource IRIs are formed accounting the document structure, \emph{e.g.} a hierarchy of contexts.
+Document entities' resource IRIs are formed accounting the document structure, \emph{e.g.}, a hierarchy of contexts.
 Using IRIs in applications solves the fundamental problem of cross-domain and cross-application object identification and forms a basis for a standardized data processing environment.
@@
 Some libraries allow one to execute \texttt{SPARQL} queries locally.
-KG is evolved by adding new triples and removing irrelevant ones.
+KG is evolved by adding new triples and removing old ones.
 The resulting modified KG is stored as files and published on a server.
@@
   \item Notions \emph{data} and \emph{knowledge} are converged into ``\emph{something} \emph{known}'';
-  \item KG contains data and knowledge in the form of resources and relations, including metadata;
+  \item KG contains data, knowledge and metadata in the form of resources and relations;
   \item The same KG is being filled in and processed by different independent modules, \emph{e.g.}, withing SPARQL queries with UPDATE statements;
-  \item KG fundamentals allow a software developer to \emph{postpone} the formal definition of a data schema;
+  \item KG fundamentals allow a software developer to \emph{postpone} the definition of a data schema;
   \item There are three points of view to graph schemata interpretation:
@@
 
-With general constraints implied by KG based domain modeling, the design of application modules are to be built on standardized data interchange formats, which are to be easily converted from and to KG representation, as well as applications are to be constructed out of service modules supporting SW endpoints (SPARQL) whenever possible.
+With general constraints implied by KG based domain modeling, the design of application modules are to be built on standardized data interchange formats, which are to be easily converted from and to KG representation, as well as applications are to be constructed out of service modules supporting SPARQL endpoints whenever possible.
 
@@
 \end{figure}
-The kernel module of applications is \texttt{KG server} (B2) built on the base of ClioPatria \cite{cliopatria} with web-server and Pengines plug-ins.
+The kernel module of applications is KG server (B2) built on the base of ClioPatria \cite{cliopatria} with web-server and Pengines plug-ins.
 ClioPatria is a web-application implemented in SWI-Prolog \cite{swi} being ISO-compliant Prolog implementation.
@@
 Pengine's knowledge base is extended with the programmer's Prolog modules.
-Kernel module stores a common global part of A-~and T-boxes of applications, \emph{i.e.} it is application database.
+Kernel module stores a common global part of A-~and T-boxes of applications, \emph{i.e.}, it is application database.
 Other modules provide an application (C3) with services.
@@
 The authoring tool engine collects web page chunks and composes the document as an XHTML page.
-This composed page can be stored in a KG as a set of triples and indexed in the Text indexing module (C1), printed with the browser's printing engine, or converted to PDF.
+This composed page can be stored in a KG as a set of triples and indexed in the text indexing module (C1), printed with the browser's printing engine, or converted to PDF.
 The tool allows the user to change document content in WYSIWYG mode.
@@
 Sphinx Search is much faster in text and database records indexing and consumes less RAM as it is implemented in C++.
-The choice of an engine depends on the primary document representation used in an application.
+The choice of an engine depends on the primary document representation used in the application.
 
@@
 The documents could be report data or scientific paper content, DJVU files, or raster images.
-Data recognition is implemented in the PDF/Text document processing module (C2), where data processing workers query the Text indexing module for BLOB documents and add recognized text and high-order structure layers.
+Data recognition is implemented in the PDF/text document processing module (C2), where data processing workers query the text indexing module for BLOB documents and add recognized text and high-order structure layers.
 All obtained layers are stored in the database of C1, and data of general interest are to be converted in triples of KG.
@@
 The aim of this R\&D \cite{zont19, aicts2020} is to develop software support of the processes of NGS analysis with techniques and software for a visual representation of the computational process for so-called amplicon analysis so that domain specialists would be able to compose computational pipelines, which are executed on distributed heterogeneous computing resources (clouds).
-Later, the visual models are used to organize the computational process, data store, and particular NGS data retrieval services.
+Later, the visual models will be used to organize the computational process, data store, and particular NGS data retrieval services.
 
@@
 The higher one is a network of objects sending messages via their interfaces.
-At the lower one, rules analyze and produce new conclusions on generated object properties.
+At the lower one is analyzing and synthesizing rules, generating new conclusions on object properties.
 
@@
 At the rendering sources stage, the three blocks are traversed in-depth.
-It leverages the known \verb|llvmlite| source code generation approach.
+This approach leverages the known \verb|llvmlite| source code generation.
 Figure~\ref{fig:block-seq} shows the main idea behind PSM object synthesis.
@@
 This example of the application of KG utilizes strongly the property of data representation flexibility for description of the software object model data as well as one central point of data store and access.
-The disadvantageous properties of RDF, namely, the weak expressiveness with triples, was neglected by predicate and object encapsulation.
+The disadvantageous properties of RDF, namely, the weak expressiveness with triples, was neglected by Prolog predicate and Logtalk object encapsulation.
 
@@
 
-This project \cite{authoring} is aimed at the construction of an environment for document authoring out of user-editable template elements, database query results, LOD-compliant web pages, and other document data.
+This project \cite{authoring} is aimed at the construction of an environment for document authoring out of user-editable template elements, database query results, LOD-compliant web pages, and other kind of a document content.
 The target document is a web page where content is inferred while loaded by the browser.
 
-All document elements are either context structures or templates filled with other elements.
+All document elements are either \emph{context} structures or templates filled with other elements.
 A context is an RDF hierarchical representation of a document data structure, providing data to be filled in a current rendered template.
@@
 All elements are SW resources.
-Templates can contain JavaScript subroutines for implementing special services, for example, table of contents out of CSS \verb|class| labeled subsections of the generated document.
+Templates can contain JavaScript subroutines for implementing special services, for example, table of contents out of CSS \verb|class| labeled HTML headers of the generated document.
 
@@
 \end{minted}
-  \caption{Part of a template representation}
+  \caption{Part of a template representation of an educational document}
   \label{fig:auth-code}
@@
   \item Friend-of-a-friend (\textbf{foaf}), agent information: individuals, legal entities, program agents;
-  \item Provenance (\textbf{prov}), references between documents;
+  \item Provenance (\textbf{prov}), references between documents, data origins, proofs, \emph{etc.};
   \item Dublin Core (\textbf{dc}), web page annotation mark up;
@@
 For each object of a GIS layer, one can associate a set of attribute values of auxiliary data.
-The attribute set is the same for each object of a layer, regardless the attribute's value is defined or not.
+The attribute set is the same for each object of the layer, regardless the attribute's value is defined or not.
 Empty values are represented as ``\texttt{null}''s.
@@
 
-\bibitem{tbl} Tim Berners-Lee, James Hendler, and Ora Lassila.
+\bibitem{tbl} T.~Berners-Lee, J.~Hendler, O.~Lassila.
 The semantic web: A new form of web content that is meaningful to computers will unleash a revolution of new possibilities.
-Scientific American, Vol.~05, 2001.
+Scientific American, May 2001.
 
-\bibitem{hogan} A.~Hogan, E.~Blomqvist, M.~Cochez, C.~D’Amato \emph{et al}. Knowledge Graphs, 2020. \url{https://arxiv.org/abs/2003.02320v5}
+\bibitem{hogan} A.~Hogan, E.~Blomqvist, M.~Cochez, C.~D’Amato \emph{et al}. Knowledge Graphs, 2020. URL:\url{https://arxiv.org/abs/2003.02320v5} (access date: 12-Dec-2021)
 
-\bibitem{lod} C.~Bizer, T.~Heath, T.~Berners-Lee, Linked Data -- The Story So Far.  Int. J. Semantic Web Inf. Syst., 5 (2009), pp.~1--22. \doi{10.4018/jswis.2009081901}
+\bibitem{zont19} E.~Cherkashin, A.~Shigarov, V.~Paramonov.
+Representation of MDA transformation with logical objects.
+Procs.
+of International Multi-Conference on Engineering, Computer and Information Sciences (SIBIRCON), Novosibirsk, Russia, 2019, pp.~0913--0918 \doi{10.1109/SIBIRCON48586.2019.8958008}
+
+\bibitem{cliopatria} J.~Wielemaker, W.~Beek, M.~Hildebrand, J.~Ossenbruggen, ClioPatria: A SWI-Prolog infrastructure for the Semantic Web, Semantic Web, 2016, volume 7(5).
+pp.~529--541 \doi{10.3233/SW"=150191}
+
+\bibitem{swi} J.~Wielemaker, T.~Schrijvers, M.~Triska, T.~Lager. SWI-Prolog. Theory and Practice of Logic Programming, volume~12, No.~1-2, pp.~67--96, 2011, ISSN~1471-0684.
+
+\bibitem{pengines} T.~Lager, J.~Wielemaker.
+Pengines: Web Logic Programming Made Easy.
+Theory and Practice of Logic Programming, volume~14, No.~4-5, 2014, pp.~539--552.
+\doi{10.1017/S1471068414000192}
+
+\bibitem{lod} C.~Bizer, T.~Heath, T.~Berners-Lee, Linked Data -- The Story So Far.  Int. J. Semantic Web Inf. Syst., volume~5, 2009, pp.~1--22. \doi{10.4018/jswis.2009081901}
+
+\bibitem{aicts2020} E.~Cherkashin, A.~Shigarov.
+Construction techniques of Baikal microbiome research information-computational environment.
+Procs.
+of 1st International Workshop on Advanced Information and Computation Technologies and Systems.
+Irkutsk, Russia, December 7-11, 2020.
+CEUR-WS.org, URL:\url{http://ceur-ws.org/Vol-2858/paper2.pdf} (access date: 12-Dec-2021)
+
+\bibitem{authoring} E.~Cherkashin, A.~Shigarov, V.~Paramonov, A.~Mikhailov, Digital archives supporting document content inference, Procs.
+of 42-nd International Convention on Information and Communication Technology Electronics and Microelectronics (MIPRO), May 20–24, 2019.
+pp.
+1037-1042.
+\doi{10.23919/MIPRO.2019.8757196}
+
+\bibitem{gisviewer} E.~Cherkashin, O.~Lunina, L.~Demyanov, A.~Tsygankov.
+Web-GIS viewer for active faults data represented as a knowledge graph.
+Procs.
+of 4th Scientific-practical Workshop Information Technologies: Algorithms, Models, Systems.
+Irkutsk, Russia, September 14, 2021.
+CEUR-WS.org, URL:\url{http://ceur-ws.org/Vol-2984/paper8.pdf} (access date: 12-Dec-2021)
 
@@
 Geodynamics \& Tectonophysics.
-V.~7(3) (2016).
+volume~7(3), 2016.
 pp.~407-434.
@@
 
-\bibitem{afs} A.~A.~Gladkov, O.~V.~Lunina. Cartographic service ``Activetectonics''. \url{http://activetectonics.ru/} (access date: 20-Sep-2021)
+\bibitem{afs} A.~A.~Gladkov, O.~V.~Lunina. Cartographic service ``Activetectonics''. URL:\url{http://activetectonics.ru/} (access date: 20-Sep-2021)
+
+\bibitem{shigpaper} N.~O.~Dorodnykh, A.~Y.~Yurin, A.~O.~Shigarov.
+Conceptual Model Engineering for Industrial Safety Inspection Based on Spreadsheet Data Analysis.
+In: Simian D., Stoica L.
+(eds) Modelling and Development of Intelligent Systems.
+MDIS 2019.
+Communications in Computer and Information Science, volume~1126, 2020, Springer, Cham.
+\doi{10.1007/978-3-030-39237-6_4}
 
@@
 
-\bibitem{zont19} E.~Cherkashin, A.~Shigarov, V.~Paramonov.
-Representation of MDA transformation with logical objects.
-International Multi-Conference on Engineering, Computer and Information Sciences (SIBIRCON), Novosibirsk, Russia.
-(2019) 0913--0918 \doi{10.1109/SIBIRCON48586.2019.8958008}
-
-\bibitem{aicts2020} Evgeniy Cherkashin, Alexey Shigarov.
-Construction techniques of Baikal microbiome research information-computational environment.
-Proceedings for 1st International Workshop on Advanced Information and Computation Technologies and Systems.
-Irkutsk, Russia, December 7-11, 2020.
-CEUR-WS.org, online \url{http://ceur-ws.org/Vol-2858/paper2.pdf} (access date: 12-Dec-2021)
-
-\bibitem{authoring} E.~Cherkashin, A.~Shigarov, V.~Paramonov, A.~Mikhailov, Digital archives supporting document content inference, Procs.
-of 42-nd International Convention on Information and Communication Technology Electronics and Microelectronics (MIPRO), May 20–24, 2019.
-pp.
-1037-1042.
-\doi{10.23919/MIPRO.2019.8757196}
-
-\bibitem{gisviewer} E.~Cherkashin, O.~Lunina, L.~Demyanov, A.~Tsygankov.
-Web-GIS viewer for active faults data represented as a knowledge graph.
-Proceedings for 4th Scientific-practical Workshop Information Technologies: Algorithms, Models, Systems.
-Irkutsk, Russia, September 14, 2021.
-CEUR-WS.org, online \url{http://ceur-ws.org/Vol-2984/paper8.pdf} (access date: 12-Dec-2021)
-
-\bibitem{cliopatria} J. Wielemaker, W. Beek, M. Hildebrand, J. Ossenbruggen, ``ClioPatria: A
-  SWI-Prolog infrastructure for the Semantic Web,'' Semantic Web.
-  Vol. 7(5).  P. 529--541 (2016) \doi{10.3233/SW-150191}
-
-\bibitem{shigpaper} Dorodnykh N.O., Yurin A.Y., Shigarov A.O.
-(2020) Conceptual Model Engineering for Industrial Safety Inspection Based on Spreadsheet Data Analysis.
-In: Simian D., Stoica L.
-(eds) Modelling and Development of Intelligent Systems.
-MDIS 2019.
-Communications in Computer and Information Science, vol 1126.
-Springer, Cham.
-\doi{10.1007/978-3-030-39237-6_4}
-
-\bibitem{pengines} {Torbj{\"{o}}rn Lager, Jan Wielemaker.
-Pengines: Web Logic Programming Made Easy.
-Theory and Practice of Logic Programming.
-Vol.~14, No.~4-5.
-pp.~539--552.
-2014.
-\doi{10.1017/S1471068414000192}
-
-\bibitem{swi} Jan Wielemaker, Tom Schrijvers, Markus  Triska, Torbj\"o{}rn} Lager. SWI-Prolog. Theory and Practice of Logic Programming, Vol.~12, No.~1-2, pp.~67--96, ISSN~1471-0684.
-
 \end{thebibliography}

### 2022-02-04 other/paper-aicts-chekan.tex
LaTeXified

%% The first command in your LaTeX source must be the \documentclass command.
%%
%% Options:
%% two-column: Two-column layout.
%% hf: enable header and footer.
\documentclass[
% twocolumn,
% hf,
]{ceurart}
\usepackage{amsmath,amssymb,longtable,hhline}
\usepackage{doi}
\def\doitext{DOI:}

%%
%% One can fix some overfills
\sloppy

%%
%% Minted listings support
%% Need pygment <http://pygments.org/> <http://pypi.python.org/pypi/Pygments>
% \usepackage{listings}
\usepackage{minted}
\setminted{fontsize=\footnotesize,mathescape}
\usepackage{hyperref}
\usepackage{multicol}

\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=true,           % non-Latin characters in Acrobat’s bookmarks
    pdftoolbar=false,        % show Acrobat’s toolbar?
    pdfmenubar=false,        % show Acrobat’s menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={},    % title
    pdfauthor={Evgeny Cherkashin},     % author
    pdfsubject={Technologies of Semantic WEB as an environment of application development and integration},   % subject of the document
    pdfnewwindow=true,      % links in new PDF window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=black,          % color of internal links (change box color with linkbordercolor)
    citecolor=black,        % color of links to bibliography
    filecolor=black,      % color of file links
    urlcolor=black           % color of external links
  }
\urlstyle{tt}
\usepackage{pifont}
\usepackage{xcolor}
\graphicspath{{pics/}}

\newcommand{\GB}[1]{\colorbox{green}{#1}}
\newcommand{\BB}[1]{\colorbox{blue}{#1}}
\newcommand{\RB}[1]{\colorbox{red}{#1}}

%% auto break lines
% \lstset{breaklines=true}

\newcommand{\ns}[1]{\textbf{\texttt{#1}}}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% Rights management information.
%% CC-BY is default license.
\copyrightyear{2021}
\copyrightclause{Copyright for this paper by its authors.
  Use permitted under Creative Commons License Attribution 4.0
  International (CC BY 4.0).}

%%
%% This command is for the conference information
\conference{2nd International Workshop on Advanced Information and Computation Technologies and Systems, December 6--10, 2021, Irkutsk, Russia}

%%
%% The "title" command
\title{Distributed Environment for Auctioning Virtual Power System Resources}
%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
\author[1,2]{Mikhail A. Chekan}[%
orcid=0000-0003-2428-2471,
email=eugeneai@icc.ru,
url=https://github.org/eugeneai,
]
\author[1,2]{Evgeny A. Cherkashin}[%
orcid=0000-0003-2428-2471,
email=eugeneai@icc.ru,
url=https://github.org/eugeneai,
]
\address[1]{Matrosov Institute for System Dynamics and Control Theory of Siberian Branch of Russian Academy of Sciences, 134 Lermontov St, Irkutsk, 664033, Russian Federation}

\address[2]{Institute for Mathematics and Information Technologies, Irkutsk State University, 20~Gagarina Bulv, Irkutsk, 664003, Russian Federation}

\begin{abstract}
  The paper discusses the development of an auction environment used as part of a ``smart grid'' simulator. Usage of the Haskell programming language, the Nix package manager, and a set of architectural solutions, such as the relative time windows and the SSH tunnel-based network architecture, made it possible to create a flexible and reliable system. This system was successfully applied both in distributed and in high-load local scenarios.
\end{abstract}

\begin{keywords}
  distributed systems \sep auctions
\end{keywords}

\maketitle

\section*{Introduction}
\label{sec:into}

The current trends in energy sector organizations in the developed countries (including Russia) pay special attention to the concept of ``smart energy.'' The environmental and economic situation forces them to optimize energy processes, switch to alternative sources, and introduce smart grids [1]. Information technologies play a special role in this process as a base for implementing a ``smart energy'' control. In this regard, the issue of training specialists with a comprehensive understanding of energy processes as being based on modern technologies is relevant. On the other hand, smart energy has not yet found good distribution at the household level in some countries. One of the effective methods for dealing with these issues is devising tools for specialized training of young specialist-engineers. The proposed game component plays an essential role in the educational effect, an energetic auction modeling tool as additional didactic material~[2].

The control synthesis is mainly based on the sequence of ``energy market players'' decisions. One of the possible ways of developing the capabilities is to organize the game based on a computer simulation of the operation of the power system. The particular product is the training stand ``Intelligent Energy Systems'' (denoted as \emph{powerstand})~[3], developed by Polyus-NT~Ltd. It allows a tutor to show the basics of planning and managing a real power supply network in a smart grid context. The product has been implemented and deployed in several quantoriums in Russia and used at major educational events, including the final stage of the track of the same name at the annual National Technological Olympiad (NTO). The product can be adapted to various kinds of customers, including industries.

The powerstand game session consists of stages. One of the stages is auctions for the ownership of objects of the conditional microdistrict assembled by the participants of the power grid. The obtained grid is subsequently modeled for energy efficiency and profitability. Since the set of objects directly impacts the simulation result, it is crucial to reliably and promptly carry out this stage of work with the stand. The problem is the imperfection of the manual auction version when the host independently collects the participants' bids using material (paper) or digital media (messages in messenger applications). We proposed automating this work stage and introducing it into the existing software packages to reduce costs and improve reliability. Power system modeling is an already essentially automated process. Reducing external dependencies through additional automation will increase the integrity of the software and hardware complex and increase the dynamism of the activities carried out on it, as well as reduce the requirements for operator qualification, which is especially important when deploying the stand as an educational tool at various sites where the direct presence of administrators and developers is not expected.

The first iteration was successfully developed and implemented as a powerstand module [4]. However, as the popularity of virtual formats of educational and competitive events grows, it becomes necessary to implement a new version of the module for supporting the distributed environment of virtual auctions when participants work at stands located in different geographically distant points. The result of designing and implementing such an environment is presented in the research.

This research and development aim to construct an efficient (in terms of the reaction speed and maintainability) distributed environment to provide game activities enabling students to conduct and participate in auctions as a part of a smart energy training stand.

The theoretical base of distributed software systems has a significant degree of development. Some works consider its aspects in detail [5]. In particular, centralized architectures and their variations (client-server, three-tier, n-tier) are well studied. Therefore, in terms of architecture, modern works [6] pay much attention to distributed decentralized (peer-to-peer) architectures, in particular, based on blockchain [7]. Since the auction environment is designed to be implemented in existing systems with a closed architecture, and the implementation of the game logic requires an arbiter in the form of a central node, a decentralized approach in the current state of work is considered suboptimal. It is also due to the complexity of implementation and support. However, as part of the hypothesis for further research, the use of a decentralized architecture is considered, including implementing the free organization of gaming sessions.

Building a network architecture with the expectation of implementation in consumer products involves following the standards for building computer networks that have been formed over the past half-century, including the OSI model and the protocol group of the transport and session layers of this model [8]. These standards have been tested by time and practice, so the task of developing and designing new solutions belongs to the application and session levels, while solutions at the session level and below are most often the de facto standard, and working with them comes down to choosing options.

In addition to technical standards, there are fundamental requirements for a distributed system, mentioned in the book by W. Jia and W. Zhou. Among these requirements, it is worth highlighting competitiveness (system nodes operate in parallel), openness (the ability to freely use and expand), fault tolerance (recoverability after errors and failures), and security (in particular, the suppression of external interference in the system). The proposed solutions must meet these requirements and be simple to implement. In particular, if we consider the software package as a combination of software modules and the execution environment, the developer has at his disposal system utilities that have already been implemented and debugged to a sufficient extent to delegate part of the functionality to them.

Jia and Zhou also mention replication to provide fault tolerance [5, p. 240]. In addition, replication solves two more problems - increasing the availability of the service (if one node fails, the functionality supports the other) and improving performance (reducing response time by dividing the data flow across different servers). Although these provisions are described mainly for decentralized systems, replication in a centralized architecture (if we consider working with data separately) also allows them to be achieved.

Software solutions for implementing distributed architectures are often reduced to embedded libraries for network interaction and message passing. It is worth highlighting libraries for implementing remote procedure call (RPC), particularly gRPC [9], a full-fledged cross-platform solution for distributed computing. However, the most common solutions related to remote message passing (message passing) require the connection of additional libraries that carry additional dependencies. It makes sense to write code for network functionality optimized for specific conditions and requirements depending on the task.

There is no consensus on the issue of organizing a network infrastructure, namely, connecting services, and most typically, services interact through open channels using standardized TCP and UDP protocols. This option provides sufficient throughput and, in the case of TCP, consistent two-way communication but imposes security and network infrastructure requirements. Problems regarding network infrastructure most frequently lie in the configuration of routing and access ports, and security issues are addressed through various add-ons related to encryption and authorization mechanisms. The most popular access control protocol at present is OAuth [10]. There are ready-made solutions [11] that implement interaction using this protocol. Also, to protect applications from denial of service attacks, balancers are used, in particular, NGINX [12].

At the same time, there are scenarios where it is necessary to block access to the application from the external network to everyone except trusted nodes. It is especially true for corporate systems. To implement such scenarios, VPN tunneling, access via a terminal connection (for example, VNC), and web interface publishing via HTTPS [13] are often used, and the latter method is the most desirable, according to the author of the source. At the same time, VPN tunneling is considered the most popular and common method of creating a closed channel between nodes. However, its use is associated with some risks and requires proper configuration. The same applies to a web application, which will require additional labor to implement. The terminal connection is used mainly for organizing user interaction with the desktop (VNC is used for this). At the same time, the SSH protocol, in addition to its primary functionality - remote interaction with the user environment - provides tunneling of TCP connections over an encrypted channel [14]. To organize a simple SSH tunnel, two nodes are enough, where an SSH server is installed on one, and an SSH client is installed on the second, launched with a specific set of arguments. Both components come standard with many Linux distributions [15]. The further configuration depends on security and infrastructure requirements.

\section{User experience analysis}
\label{sec:user-exp}

The auction, as mentioned above, the module has become a prototype of the distributed environment. Being a part of a powerstand software, the module is used in educational institutions, engineering competitions, and intensive programs. During the powerstand operation and administration and personal observation of auction system operation, a significant amount of user experience was accumulated, based on decisions made regarding developing new versions of software products and creating unique technical solutions. This experience has become the primary source of expertise in forming requirements for a new interpretation of the auction system in distributed scenarios.

Three events made the dominant contribution to the expertise. This section will describe them and briefly analyze the user experience.

The first auction module application was at the finals of the “Intelligent Energy Systems” profile of the 2018 NTI Olympiad [16] at the Sirius Educational Center in Sochi. Even though the module was still in the prototype stage and the game model was more straightforward than manual conduct, the auction automation significantly improved the quality of the game session. Bets were sent directly from the teams’ workplaces, and the round result was determined automatically. That increased the game dynamics, eliminated organizational costs, and brought a positive experience to the process.

The educational intensive program ``Island 10-21'' in the same year from July 10 to July 21 at the Far Eastern Federal University (Russkiy Island) [17]. A modified version of the auction module was used as a part of a stand at a ``smart energy'' educational laboratory. The game model was brought to the current one, the same as in the manual auction. The administrator and observer interfaces were added. The developer was personally present at the event, administering the stand and observing the laboratory conduction. Due to the higher speed of the game and rapid UI feedback, the automatic auctions' usage makes the object distribution fast and dynamic. Playing at the auction evoked positive emotions among the laboratory participants. Displaying auction results through a projector added spectacle to the process and allowed outside observers to track and analyze auction results in real-time. The administrator's interface turned out to be necessary only for the game session starts and restarts, as well as occasional pauses to describe the game situation.

Nevertheless, it turned out to be very useful in terms of information content, as it displayed bids for the current lot (the participants themselves see only the fact of the decision made by the opponent). It allowed the organizers of the educational program to track the participants' behavior and make decisions on further work within the laboratory.

From March 13 to March 17, 2019, the IPS profile finals of the NTI Olympiad took place. It was the first profile final in a distributed format and the first experience of distributed games on the powerstand. The competition was at two venues: the MEPhI Nuclear Research University (Moscow) and the Scientific Library of Irkutsk State University. Most of the stand software was done from scratch for the new game and architectural model. The auction module remained intact, except for a slight modification of the game model and adaptation of the interface for eight players. Thus, the user interface was interacting with the server part of the auction module hosted on the cloud server.

The rules of the event included practice games both in distributed and local formats. Local format implied the core services (incl. the auction module) to be placed in the stand instead of a cloud server. A service manager was implemented to set up and monitor applications based on a configuration file to perform mode switching. In practice, the service manager became a source of hard-to-diagnose bugs but still managed to complete the task.

The result of the final was ambiguous due to technical problems associated with an unstable communication channel on the Irkutsk site, which also held the finals of two more profiles. The protocols of the stand software modules were not adapted to low bandwidth networks, which caused noticeable delays for 5-6 s on average. They were noticeable, disrupted the automatic auction dynamics, and created unequal conditions between sites. Latency was negatively impacted by a query time limit issue in the web framework used. Backend requests took a lot of time in addition to JSON state deserialization (due to a complex structure and more than 100 kilobytes size). Thus, the delays increased to those values ​​at which participation in the auction became difficult. There was also a desynchronization: the round countdown was based on the UNIX timestamp created on the cloud server, while the clock on the terminal could differ from the server clock by more than a few seconds, which, together with the delay, created a situation when the bet was sent but came to the server too late. Despite the current situation, organizers managed to conduct training and final games correctly. Nevertheless, after this event, we decided to review the network architecture organization and the protocol of all stand modules to obtain more stable operation in the conditions of low-quality or slow communication channels. Within this direction, the idea of ​​this work arose.

During observation and experiments related to the auctions module, it was found that a 1-2 second delay in switching the state of the auction does not disrupt the dynamics of the auction process. JSON packets up to 100 kilobytes in size are processed by the browser within a second, which fits within the allowable time with a good communication channel.

\section{Requirements}
\label{sec:reqs}

The powerstand specificity determined the game model, namely its succession to the previous version. Also, the technical component of the stand determined the basic architecture of the environment, consisting of three levels:
\begin{itemize}
\item the cloud level, which implements the logical core of the stand (game session control, game logic);
\item  the stand (relay) level, which coordinates data transfer between the cloud and terminals;
\item  the terminal (interface) level, which implements the user’s interaction and transfers one's actions to the upper levels.
\end{itemize}

The terminal and stand levels exist close to each other as a part of a powerstand. Client stations implement the terminal level and are connected to the controlling unit with the powerstand server inside. Clients and servers are connected via a wired LAN, and the router used for it provides bandwidth up to 1 Gb/s. It provides a high-speed and low response time for communication between terminals and relays. At the same time, data can be transmitted openly, which creates a potential vulnerability in the normal gameplay: a third-party client can intercept the data stream and disrupt the game, ranging from a gameplay disbalance up to system failures. It implies network security and also enables us to implement static module authorization. For example, the configuration of terminals is known in advance, and it can be set strictly statically in one or another module of the system.

In the base scenario (such as holding events using power stands), the system architecture assumes the presence of several such physical stands located at geographically different points and connected to a cloud server. This arrangement implies heterogeneity of network channels and, as a result, the problem of synchronization of transmitted data and control actions. From this follows the requirement for leveling network delays. The helping fact is that the auction gameplay consists of rounds that do not require strict synchronization, but only relative simultaneity (difference within 1-2 seconds is allowed) and equality of a countdown on all stands. It is based on centralized game logic: the cloud level sets the game’s pace, accepts actions from players and physical objects from all stands, processes them, and sends out updated data.

Recalling the third-party interference problem, the data transfer and user actions are required to be minimally subject to outside interference with the relative ease of authorization and connection of modules to each other.

Thus, the resulting network is heterogeneous: there is a high-speed intra-stand network level and an inter-stand layer associated with data transmission through the core. However, there may be a hardware or software problem at both levels, for example, network channel instability or involuntary disconnection of terminals. The solution would be to implement a system that would continue a game session on a module failure and which modules can quickly recover from such a failure.

Finally, the requirement for architecture is its flexibility. So, stand usage scenarios include the possibility of disconnecting the stand from the cloud server for local gaming sessions (standalone mode). In this case, the cloud level is situated at the same physical place as the relay level. It is expected that the stand reconfiguration to a single-mode will require minimal effort and will have minimal impact on the source code of game services.

Based on the experience described above, and the sources studied, the requirements for a distributed environment for conducting virtual auctions have been formed: the implementation of a stable dynamic architecture with the following properties:
\begin{itemize}
\item minimizing the risks of external interference while maintaining the simplicity of connecting and authorizing modules (considering their known configuration);
\item leveling network delays while maintaining the relative synchronization of the game session and the equality of the duration of cycles;
\item system stability preservation in case of a child module failure and fast module recovery in such a case.
\end{itemize}

Game model (перевести картинку)

The construction and justification of the game model are based on the statements of the auction theory, which studies auctions as microeconomic models using game theory. Its accessible and popular presentation was made by A. Filatov and A. Savateev [18]. One frequently mentioned statement is the revenue equivalence theorem, proven by R. Myerson [24]. According to this theorem, the average revenue of the seller and the expected payoffs of the participants do not depend on the model and rules used, provided that three conditions are met:
\begin{enumerate}
\item ``No entry ticket'': the participant with no lot evaluation pays nothing.
\item The lot is always given to the participant with the highest bid.
\item Bid evaluations of auction participants are distributed independently and equally (that is, symmetric equilibrium is presented).
\end{enumerate}

All these statements are true in the context of the chosen applied task, from which we can conclude that it is possible to freely choose the auction format (as well as their combinations) for use in the game model. Thus, the first-price sealed-bid auction is used as a game model. The use of a sealed-bid format is caused by its resistance to short time delays, in contrast to the open dynamic formats such as English and Dutch auctions, where the order of bids is important. It ensures the highest possible equality of conditions among bidders in distributed scenarios.

Lots are put up for auction, which in the applied task are game objects, buildings with an individual address. Participants write bids for the declared lot and send them sealed (without revealing them to competitors) to the host (a person or an automated system).
The bid value has a predetermined limit range, which is individual for each lot (in the applied task, it depends on the object type). Such a restriction is set to form a sensible economic situation. The bottom limit comes from considering that all objects have value, and the upper limit protects participants from excessive losses in an upward auction and excessive profits in a downward auction.
In addition to sending the bid, the participant can refuse to bid on the lot (namely, pass it). Passing does not incur sanctions and penalties to the participant (principle of ``no entry ticket'').

The bid time is limited by a countdown timer, after which the bids are compared. They are sorted in ascending or descending order depending on the type of lot: for consumption objects, rates are tariffs per power unit and are compared downwards; for generation objects, the rates are the rent per game tick and are compared upwards. If the winner is unambiguous, the lot is closed and assigned. If no single bid has been made on the lot, it is declared unplaced and does not participate in the further game at the stand.

The situation is possible when several participants’ evaluation is almost the same, but the probability of an exact match of their rates is insignificant. Therefore, ``catching up'' bids are used for a more competitive distribution of lots. If some bids differ from the leading one by less than 0.5 money units, another round is held for the same lot with the leader and owners of ``catching up'' bids.

If one of the participants has bid a limit price (for example, the minimum value in a downwards auction), then the ``catch-up'' bid mechanic does not work, and the lot is assigned to this participant. If two or more participants set the limit bid, a new round is announced on the all-pay principle. There, bids are fixed amounts that bidders are willing to pay to get the lot at the marginal price. However, regardless of the outcome, the bet sent by the participant is debited from his account as a penalty. In the main game, the total value of the penalty for all-pay is deducted from the initial score.

Given the conditions for revenue equivalence, we can consider an all-pay auction to be equivalent to a sealed-bid first-price auction in terms of determining value. By bidding on the limit value, the participants set the personal value of the object, sacrificing its objective value (economic effect). To resolve the disputable situation, the personal value is determined again by bidding on the amount of money the participant is willing to give to own the object. Since the increase in this value negatively affects the final result of the participants, it was decided not to apply the "catch-up" bets rule in all-pay, holding a second round only if the leading bets completely match.

After the first auction pass, when the objects are pre-assigned, participants can refuse a certain number of their lots, putting them on the second pass of the auction, along with those that were unplaced on the first. In case of refusal of a lot, the participant loses the right to compete for this lot. The value of the refused lots is then determined anew (that is, the bid range is reset to the original one).

At the end of the auction, the results tally is generated to represent information about the ownership of the objects, the bids set on them, and the submitted all-pay bids.

According to the game model described above, a state diagram is drawn, shown in Figure~\ref{fig:1}.

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{pics/statechart.png}
  \caption{Auction state model}
  \label{fig:1}
\end{figure}

This model allows simultaneous independent draw of several lots. This further increases the dynamics and cognitive complexity of the auction, forcing participants to more carefully coordinate their decisions about the value of the lots to keep pace with the pace of the auction.

The described model is abstract enough and contains almost no binds to the stand domain. Therefore, the model enables solving other tasks related to the distribution of virtual resources. It is also possible to change the model implementation, if necessary, i.e., add additional rules.

\section{Implementation architecture}
\label{sec:impl-arch}

A three-level hierarchical architecture is implemented as a result of the analysis of existing developments in networks and distributed systems. For convenience, we also divide it into three layers: system (operating system tools), software (implementation of server and client programs), and transport (protocols and data transfer principles).

We note the need for stable and consistent two-way data transmission in the transport layer. The TCP protocol successfully performs it, so it became the transport core. Typed messages are sent over the established connection. Although it is possible to re-implement the stream-based protocol with length control, it was decided to use the WebSocket protocol [20]. It has high-quality implementations for most programming languages and enables further modifications (direct browser communication, TLS encryption, \emph{etc}.). Messages' size should be reduced to a reasonable minimum, so the CBOR format [21] is used, as it provides the highest serialization speed with the minimum message size [22].

To minimize network traffic between levels, atomic state changes (denoted as effects) are transmitted instead of the new state to minimize network traffic between levels. An effect is generated for each event and passed to all child nodes. The full state is transmitted only when the game session is completely restarted and when the child node is reconnected (in future iterations, it is possible to implement persistent caching and full synchronization of such a state).
When considering the game state, the effects are following:
\begin{itemize}
\item the game state change (for example, the start of the auction is the transition from preparation to the active state);
  \item  a pause setting or removing;
  \item  lots rejection start (time limit, the number of lots to reject);
  \item  lot rejection finish (list of lot indices that need to be reset for the second draw);
  \item  a singular lot of state change;
  \item  bids update for a given lot (index, current bid list);
  \item  mid-rejection update (the current list of lots rejected by the participants).
\end{itemize}

The user response messages consist of the player ID (the number of the site and the number of the player on it) and the required action, accompanied by parameters:
\begin{itemize}
\item  send a bid for a lot (lot ID and bid value);
\item  pass a lot (lot identifier);
\item  reject the lot to the second draw (lot identifier);
\item  finish lot rejection.
\end{itemize}
The auction is designed to support multiple lots simultaneously.  The player's identification and the lot to which the action is applied are needed.

The identification of the client sending the message is done by adding its identifier to the message body. At the same time, due to the fixed connection scheme of nodes in a closed system, a certain level of trust is maintained, which, together with authorization, is provided at the system and program levels. In addition to influencing game state by players, there are actions sent through the admin interface, represented by a separate HTTP endpoint.

The system layer provides the infrastructure for connecting servers by SSH tunnel-based port forwarding. As mentioned, tunneling TCP connections via SSH does not require complex configuration. In particular, the OpenSSH server and client are included in the standard system software of a wide range of Linux distributions, and the ease of configuration makes this tool a suitable solution for building network applications with a distributed architecture. The construction of architecture on SSH tunnels is described in more detail in [7]. Since the SSH protocol supports key-based authorization, this method allows to set up a secure, stable communication channel between network nodes and explicitly define the connection topology. This solution is expected to provide sufficient reliability while meeting the minimum guaranteed transfer speed. It also simplifies the implementation of server programs (because only a connection to a local port is needed) and provides the ability to debug the entire system on a single computer.

For this architecture to function appropriately, SSH tunnels must be configured automatically. If using command-line arguments is sufficient for the client, the server must be configured through a configuration file. At the same time, the configuration may differ depending on the scenario of using the environment, and it may be necessary to quickly reconfigure the system nodes. There are tools for automating these processes (Ansible, Chef), but they are convenient for deploying individual components, but there are scenarios of OS-level configuration for multiple nodes. Therefore, it was decided to choose the NixOS Linux distribution [23].

The main feature of NixOS is that the system is configured using configuration files in the Nix language, which is used in the package manager of the same name. The builder generates a static system environment using the configuration files, consisting of symbolic links to isolated packages in the repository. The Nix package manager allows storing and using different versions of libraries and programs simultaneously. All packages are contained in the repository isolated from each other and the main system; adding or removing a package does not change system files but only changes symbolic links to the necessary components [24].

The main advantage of Nix packages and the NixOS system is reproducibility and portability. Whenever a Nix package generates a workable program, it is highly likely to be deployed on any computer based on the same platform. Likewise, if the NixOS description can form a system, it can be deployed on any computer, and all system instances will be functionally identical. The architecture described in the Nix language is easily parameterized, provides flexibility and ease of deployment on numerous configurations, and allows combining the description and the structure of the network architecture with the deployment of nodes of this architecture, which comprise a complete distributed software package. It leads to the choice of NixOS as a technical solution for organizing the architecture and introducing services into the software system.

Three applications constitute a three-tier client-server architecture in the software layer: the core (cloud level), the relay, and the mediator (terminal level). We consider the user interface a separate entity interacting with the mediator and the administrator interface interacting with a particular HTTP endpoint of the core. The core server controls the game session logic and accepts connections from the relays, updating the current session state for them and processing incoming commands from the players. The relay server is designed to reduce the load on the network and optimize the interaction between participants and the core using replication and time amortization. The mediator server keeps a copy of the state and projects it into a suitable form for processing by the UI. Thus, the tasks of authorization and identification are performed entirely by tunnels and the mediator. In the interface, it is necessary to implement only the display of data and the transmission of commands to the kernel through the mediator using a simple API.

Haskell was chosen as the primary language for implementing services [25]. It is a standardized general-purpose purely functional programming language. The combination of a high-level strict type system, a set of compiler-level guarantees, and the availability of libraries for creating multithreaded web services made Haskell the standard language for working on power bench modules and conditioned the choice of language for this work.

The Dhall [26] format, which also implements strict typing, was guided by the same principles and was selected for configuration files. It prevents configuration errors when the configuration is read and interpreted at the very launch. But at the same time, it is necessary to explicitly specify all types, including sets of valid values ​​for all fields universally. On the other hand, the fundamental descriptions can be placed in an imported prelude module that describes the domain-specific language and greatly simplifies further configuration. Development experience has shown that this format is more reliable and efficient than JSON and YAML.

As mentioned above, atomic changes are transferred between levels, so each server keeps a replica of the game state to apply the changes. It also allows to quickly restore the state of child nodes whether they fail. The general state of the session represents the game state (preparation, first pass, rejection of lots, second pass, completion), the replica, a set of played lots, and a list of current lots with timers.

The lot is represented by a set of fields, among which the most important are
\begin{itemize}
\item  lot identifier;
\item  the initial range of rates and the direction of their comparison;
\item  a list of participants who can compete for the object (or the absence of this restriction);
\item  history of bets (to analyze the game situation by all interested parties);
\item  fines for all-pay for all participants;
\item  current state of the lot (not played, standard round, all-pay round, finished).
\end{itemize}

To effectively update the lots' state, the timers are stored in a separate structure, where the index of a lot currently being played is matched with the UNIX time. Upon reaching this time mark, the state of a lot is updated. This structure is represented by a fixed-length list with option-type elements for efficient use in scenarios with a simultaneous drawing of several lots. This storage method explicitly indicates the limit on the number of lots that can be played simultaneously, making it possible to replicate the state more efficiently.

An important factor is the leveling of delays in data transmission and maintaining the equality of the cycle duration (that is, the time to send a bid). To ensure this factor, it was decided to use the absolute time to control cycles exclusively within the same level and the relative time when transferring countdowns between levels. So, the core sends the relay the time for the next cycle, not the UNIX time of the cycle completion. Still, the number of seconds must be counted from the moment the message was processed. After this period, the transmission of messages for this lot should be stopped. The same is implemented at the level between relays and mediators. The side that sent the relative time also uses it internally but amortizes the time interval by adding a certain (value is selected depending on the characteristics of the channel) number of seconds during which commands continue to be received based on the possible network delay.

Obviously, at the relay-mediator level in the local network, such delays are minimal. Still, an additional delay from 1 to 3 seconds brings additional reliability and a guarantee of message processing in case of possible local network overload. At the same time, the auction dynamics will not undergo negative changes, including because the game logic provides for early completion of the cycle when receiving messages from all players. All participants perform an action before the timer expires in most use cases. One of the possible functional improvements in this direction could be the automatic message sending about skipping a turn after the timer expires. Due to time amortization, timer replication may not be performed in total (when reconnecting during an active draw, it is necessary to transfer the correct relative remaining time per lot, which is difficult to do accurately enough to restore synchronization).

In the current iteration, it is decided that player identification is carried out by mediator settings built into the system configuration. In the future, it is possible to implement authorization and identification by passwords or IP addresses (which will require eliminating tunnels between relays and mediators).

All three programs are launched for greater stability at the system level as Systemd services configured to automatically restart upon a critical error. From the server-side, a sudden disconnection of the client should not incur problems because protocols provide state synchronization, and messages about the players’ actions are not necessary for the continuation of the game session.

\section{Evaluation}
\label{sec:eval}

The developed auction environment was tested at events as part of the third version of the STIES stand. From autumn 2019 to spring 2021, the powerstand was used at two all-Russian events, which are the finals of the "Intelligent Energy Systems" profile of the NTI Kruzhok Movement Olympiad (currently referred to as NTO) in 2020 and 2021, respectively. These activities allowed to perform live testing of the auction environment in the conditions of intense engineering competition to personally observe the user experience, which made it possible to identify the weak and strong points of the developed system and identify areas for further research and development.

\subsection{Distributed scenario}
\label{sec:eval-distr}

The final “Intelligent Energy Systems” profile of the NTO in 2020 was held from March 10 to 14. 60 schoolchildren from different cities of Russia, united in 13 teams, competed in a shared virtual environment while physically being in three different places:
school №709 Moscow (4 teams),
Tyumen State University (5 teams),
Novosibirsk State Technical University (4 teams).
It is worth noting that this was the first approbation of the updated auction system, and the specified configuration is the maximum on which the powerstand has ever been operated. The testing was successful. Thanks to the optimized architecture, the delays between state changes averaged no more than 1 second due to the processing of the JSON packet in the web interface. The peak delay was about 3 seconds, caused by problems with the Internet channel at the Novosibirsk site.

\subsection{High-load local scenario}
\label{sec:eval-high}

From March 29 to April 3, 2021, Tyumen State University hosted the finals of the "Intelligent Energy Systems" profile of the NTO. 10 teams took part in this final. But due to organizational difficulties, all teams were placed on the same site. It meant the need to configure a set of stands for 10 participants. It became the primary technical challenge of the event. So, 4 stands were set up (and, accordingly, 4 servers), the game core and auction relay were placed on one of these servers, the rest of the necessary components were placed on the rest. All computers (4 servers and 10 terminals) were connected to the same physical network with a router.

The event was a success. The auction environment worked well in a local high-load configuration, providing minimal delays in terms of UI perception: state change delays were no more than a second. Moreover, they were caused mainly by the processing of the JSON packet in the browser, and if we do not take them into account, there were practically no delays in data transfer. The maximum size of a LAN packet transmitted was about 60 kibibytes (the CBOR-serialized game state sent on reconnecting or starting a new game session).

The SSH-based architecture flexibility made it possible to separate the auction module from the powerstand core services without additional changes to the program code. The use of statically typed Dhall for configuration ​​justified itself, as with the correct settings, the module is guaranteed to work. In a case of an error, it writes a diagnostic message.

\section{Analogous systems brief review}
\label{sec:analogs}

The initial development of a virtual auction system aimed to automate and optimize the auction as part of the powerstand subject area. When searching for analogs among auction simulators, we posed the criteria of the possibility of complete control over the process of holding an auction, setting up the objects being played, and the possibility of embedding a simulator in an arbitrary subject area with a distributed scenario of operation. It should be noted that no direct analogs of the proposed development that meet the requirements have been found.

Investigated auction simulators are based on an entertaining game component and have no direct customization. One class is mobile games, e.g., Bid Wars [27], whereas auction simulators are frequently built into the existing game model as a side mechanic. Multiplayer online games [28] are examples of simulation. An aggravating factor in the compliance of analogs with the set requirements is a fixed model of conducting, most often corresponding to the English auction, when bids are announced openly, and the auction for the lot is conducted until there are no new bids.

There are also highly specialized commercial developments and resources related to public procurement, such as Sberbank-AST [29], but they are also not suitable for the task in terms of the overwhelming requirements.

\section{Further development}
\label{sec:further}

User experience gained during trial runs of the system during development and events showed the viability of the developed architecture and software solutions in the applied task. Nevertheless, the system has some issues and ambiguous traits with all the effectiveness.

Most of the possible issues are related to SSH tunneling. SSH uses TCP as the base protocol. Using TCP in the internal protocol is redundant and carries the risk of destabilizing the channel [30]. In practice, SSH tunnels did not harm the environment operation.

Being based on SSH tunnels, the architecture configuration is static at the system level, and connections are made according to the “allow list” principle, which is why any change in the architecture requires the action of developers. At the same time, authentication is currently absent, and the player is determined through the mediator configuration based on the trust of the nodes for each other. At the same time, the implementation of the authorization will be required only in scenarios for operating the module outside the current architecture and can be implemented in the future.

Thus, the choice of architecture implementation at this stage is considered justified, taking all positive and negative factors into account. Nevertheless, it makes sense to apply developments to increase the speed and reduce delays in SSH tunneling [31, 32]. Alternatively, even abandon SSH tunnels in favor of other transport protocols, such as a Secure WebSocket with a well-built authorization. It will assume the complexity additions to the software layer but is considered in the future.

Another class of problems is the user interface, its connection with the mediator using an HTTP polling with a JSON response. Although a near-second delay for JSON deserializing does not affect the auction’s perception and dynamics with a fast network channel, this is currently a bottleneck when transferring data to the interface. The solution to this problem could be to use CBOR instead of JSON or implement WebSockets communication, making it possible to eliminate the mediator as an extra intermediate link. Despite this, JavaScript has less performance than Haskell (being an interpreted language), and the use of a mediator for intermediate marshaling is justified at the current stage. However, it creates problems with CORS when accessing localhost from a non-local domain [33], which in our case hosts the user interface. The workaround here is to disable CORS when accessing the UI, but replacing the mediator with direct access to the interface is considered an assumption in conjunction with the above.

As a distant hypothesis, it makes sense to ​​decentralize the control of the state of the game session so that if the core fails, it would be possible not to restart the game but load the last state from one of the relay servers. This feature requires additional elaboration, considering the trust between the core and relay levels. In addition to this, the issue of time control will be relevant because the relay uses timers with relative time settings, and its transfer back to the level up is problematic. It prompts a discussion about the justification of the relative timing mechanism. In the current implementation, this mechanism is justified, since each player is given the same amount of time to decide on a bet on the exposed lot. The lot is closed ahead of schedule as soon as the last participant sends his bet. On the other hand, there is a suggestion that it is possible to synchronize the time qualitatively by using a common NTP (Network Time Protocol) server in all network nodes and minimize the need for amortization and relative timing.

Separately, it is worth highlighting as a direction of research the provision of users with the opportunity to independently organize an auction session with the choice of participants and the organization of data transfer using replication (as well as other developments within the above hypotheses, when they are tested and found effective). The technical stack and set of solutions allow us to work in this direction.

\section*{Conclusions}
\label{sec:conc}

The environment for conducting auctions for virtual resources was designed and developed, operating in a three-level architecture with the possibility of its flexible restructuring. The game model allows objectively and dynamically distributing resources designated by the domain, and the combination of technical solutions ensures the security and reliability of the distribution process.

The resulting environment was successfully implemented in the third version of STIES software. It was tested at two All-Russian engineering competitions, demonstrating its usefulness and reliability and confirming in practice the initial hypothesis about the system architecture.

The developments and concepts identified during the study, in particular, state replication and the use of network architecture based on SSH tunnels, are successfully applied in other modules of the STIES software package, which improves performance reliability and, as a result, the quality of the final product.

Despite the functional and technical viability of the described solutions and software implementation, issues are associated with the irrational use of transmission channels and strong dependence on the core. There is a hypothesis about eliminating centralization, and further research is needed to analyze this issue. It and other tasks are included in the plans for further work, which covers a larger volume than the work on the virtual auction environment.

\section*{Acknowledgments}
\label{sec:acq}

The results were obtained within the state assignment of the Ministry of Education and Science of Russia, the project ``Methods and technologies of a cloud-based service-oriented digital platform for collecting, storing and processing large volumes of multi-format interdisciplinary data and knowledge based on the use of artificial intelligence, a model-driven approach, and machine learning'', No.~FWEW-2021-0005 (State registration No.~121030500071-2).

The results were obtained with the use of the network infrastructure of the Telecommunication center of collective use ``Integrated information-computational network of Irkutsk scientific-educational complex'' (\url{http://net.icc.ru}).

\begin{thebibliography}{99}
   \bibitem{b1} Business for society: how Siemens is changing life in Russia. Smart Energy. RBC, 2016 (in Russian). URL: http://siemens.rbc.ru/article1.html (access date: 05-15-2021)
   \bibitem{b2} Makarov E.B. Game forms of education. Open lesson. (in Russian) URL: https://openlesson.rf/articles/556327/ (access date: 05-15-2021)
   \bibitem{b3} Intelligent energy systems. Stand-simulator for programming and management of urban power networks. (in Russian) URL: http://polyus-nt.ru/powerstand.html (access date: 11/22/2021).
   \bibitem{b4} Chekan M.A. Development of a module for conducting virtual auctions for the training stand ST-IES LLC "Polyus-NT". Procs. of Wiener Readings - 2019: materials of the All-Russian Youth Scientific and Practical. conferences. - Irkutsk: Publishing house INRTU, 2019,  pp. 100-105. (in Russian)
   \bibitem{b5} Jia, W., Zhou, W. Distributed Network Systems: From Concepts to Implementations. Springer, USA, 2005, 513 p. ISBN: 9780387238395
   \bibitem{b6} Burns B. Designing Distributed Systems: Patterns and Paradigms for Scalable, Reliable Services. O'Reilly Media, Inc., Japan, 2018, 149 p. ISBN: 9781491983614
   \bibitem{b7} Raval S. Decentralized applications: harnessing Bitcoin's blockchain technology.  O'Reilly Media, Inc., USA, 2016. 103 p. ISBN: 9781491924549
   \bibitem{b8} Sunshine, C. A. Computer Network Architectures and Protocols., Springer US, 2013. 558 p. ISBN: 9781461308096
   \bibitem{b9} gRPC. URL: https://grpc.io (access date: 05-15-2021)
   \bibitem{b10} OAuth Community Site. URL: https://oauth.net (access date: 05-15-2021)
   \bibitem{b11} ORY Oathkeeper. GitHub. URL: https://github.com/ory/oathkeeper (access date: 05-15-2021)
   \bibitem{b12} NGINX.URL: https://nginx.org (access date: 03-21-2021)
   \bibitem{b13} Skorodumov A. Secure access to the company's information resources.  Information Security No. 2. URL:\url{http://lib.itsec.ru/articles2/sys_ogr_dost/bezopasnyy-dostup-k-informatsionnym-resursam-kompanii} (access date: 05-15-2021)
   \bibitem{b14} SSH Tunnel. SSH.COM URL:https://www.ssh.com/ssh/ tunneling/ (access date: 05-15-2021)
   \bibitem{b15} Users. OpenSSH URL: https://www.openssh.com/users.html (access date: 05-15-2021)
   \bibitem{b16} NTO (in Russian) URL: https://nti-contest.ru/ (access date: 05-20-2021)
   \bibitem{b17} Educational intensive "Island 10-21". Information Bureau 20.35 (in Russian) URL: https://ntinews.ru/event/obrazovatelnyy-intensiv-ostrov-10-21.html (access date: 05-20-2021)
   \bibitem{b18} Savvateev A.V., Filatov A.Yu. Theory and practice of auctions. Bulletin of the Voronezh State University. Series: Economics and Management. Voronezh, 2018. No. 3. P. 119-131. (in Russian)
   \bibitem{b19} Myerson R. Optimal auction design. Mathematics of Operations Research. 1981. Выпуск 6. № 1, С. 58-73. https://doi.org/10.1287/moor.6.1.58
   \bibitem{b20} RFC 6455 — The WebSocket Protocol. IETF URL: https://tools.ietf.org/html/rfc6455 (access date: 05-20-2021)
   \bibitem{b21} Concise Binary Object Representation : [сайт]. 2021. URL: https://cbor.io/ (access date: 20.05.2021)
   \bibitem{b22} Comparison of JSON Like Serializations. URL: http://zderadicka.eu/comparison-of-json-like-serializations-json-vs-ubjson-vs-messagepack-vs-cbor/ (access date: 21.05.2021)
   \bibitem{b23} NixOS Linux. URL: https://nixos.org/ (access date: 05-15-2021)
   \bibitem{b24} Galatenko V.A., Dzabraev M.D., Kostyukhin K.A. Choosing a Package Manager for Multiversion Applications. Software products and systems. V. 31, No. 3, 2018. pp. 469–474. (in Russian) DOI: 10.15827/0236-235X.123.469-474.45
   \bibitem{b25} Haskell Language. : [сайт]. 2021. URL: https://haskell.org (access date: 20.05.2021)
   \bibitem{b26} Dhall configuration language, The : [сайт]. 2021. URL: https://dhall-lang.org/ (access date: 21.05.2021)
   \bibitem{b27} Bid Wars // Google Play [Электронный ресурс]. 2021. URL: https://play.google.com/store/apps/details?id=br.com.tapps.bid wars (access date: 05-15-2021)
   \bibitem{b28} TimeZero. MMORPG Handbook. Auction. (in Russian) URL: http://web.archive.org/web/20160325215014/ https://www.timezero.ru/manual/auction.ru.html (access date: 05-15-2021)
   \bibitem{b29} Sberbank-AST (in Russian) URL: http://www.sberbank-ast.ru/purchaseList.aspx (access date: 05-15-2021)
   \bibitem{b30} Titz O. Why TCP Over TCP Is A Bad Idea. URL: http://sites.inka.de/bigred/devel/tcp-tcp.html (access date: 21.05.2021)
   \bibitem{b31} Rapier C. et al. High performance SSH. SCP-HPN-SSH URL: http://www.psc.edu/networking/projects/hpn-ssh (access date: 21.05.2021)
   \bibitem{b32} Rapier C., Bennett B. High speed bulk data transfer using the SSH protocol. Proceedings of the 15th ACM Mardi Gras conference: From lightweight mash-ups to lambda grids: Understanding the spectrum of distributed computing requirements, applications, tools, infrastructures, interoperability, and the incremental adoption of key capabilities, 2008,  pp. 1-7. DOI: 10.1145/1341811.1341824
   \bibitem{b33} Hobbs, S. CORS Tutorial: A Guide to Cross-Origin Resource Sharing. Auth0. URL: https://auth0.com/ blog/cors-tutorial-a-guide-to-cross-origin-resource-sharing/ (access date: 21.05.2021)
\end{thebibliography}
% \section{Online Resources}

% The sources for the viewer are being developed at Github, URL:~\url{https://github.com/De17eon/GRL}.

\end{document}

%%
%% End of file

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper-aicts-chekan.tex"
%%% End:

### 2022-02-04 other/paper-aicts-chekan.tex
references are ok now.

@@
 \begin{thebibliography}{99}
-   \bibitem{b1} Business for society: how Siemens is changing life in Russia. Smart Energy. RBC, 2016 (in Russian). URL: http://siemens.rbc.ru/article1.html (access date: 05-15-2021)
-   \bibitem{b2} Makarov E.B. Game forms of education. Open lesson. (in Russian) URL: https://openlesson.rf/articles/556327/ (access date: 05-15-2021)
-   \bibitem{b3} Intelligent energy systems. Stand-simulator for programming and management of urban power networks. (in Russian) URL: http://polyus-nt.ru/powerstand.html (access date: 11/22/2021).
-   \bibitem{b4} Chekan M.A.
+   \bibitem{b1} Business for society: how Siemens is changing life in Russia. Smart Energy. RBC, 2016 (in Russian). URL:\url{http://siemens.rbc.ru/article1.html} (access date: 05-15-2021)
+   \bibitem{b2} E.~B.~Makarov. Game forms of education. Open lesson. (in Russian) URL:\url{https://openlesson.rf/articles/556327/} (access date: 05-15-2021)
+   \bibitem{b3} Intelligent energy systems.
+Stand-simulator for programming and management of urban power networks.
+(in Russian) URL:\url{http://polyus-nt.ru/powerstand.html} (access date: 11/22/2021).
+   \bibitem{b4} M.~A.~Chekan.
 Development of a module for conducting virtual auctions for the training stand ST-IES LLC "Polyus-NT".
@@
 conferences.
-- Irkutsk: Publishing house INRTU, 2019,  pp.
-100-105.
+Irkutsk: Publishing house INRTU, 2019,  pp.~100-105.
 (in Russian)
-   \bibitem{b5} Jia, W., Zhou, W. Distributed Network Systems: From Concepts to Implementations. Springer, USA, 2005, 513 p. ISBN: 9780387238395
-   \bibitem{b6} Burns B. Designing Distributed Systems: Patterns and Paradigms for Scalable, Reliable Services. O'Reilly Media, Inc., Japan, 2018, 149 p. ISBN: 9781491983614
-   \bibitem{b7} Raval S. Decentralized applications: harnessing Bitcoin's blockchain technology.  O'Reilly Media, Inc., USA, 2016. 103 p. ISBN: 9781491924549
-   \bibitem{b8} Sunshine, C. A. Computer Network Architectures and Protocols., Springer US, 2013. 558 p. ISBN: 9781461308096
-   \bibitem{b9} gRPC. URL: https://grpc.io (access date: 05-15-2021)
-   \bibitem{b10} OAuth Community Site. URL: https://oauth.net (access date: 05-15-2021)
-   \bibitem{b11} ORY Oathkeeper. GitHub. URL: https://github.com/ory/oathkeeper (access date: 05-15-2021)
-   \bibitem{b12} NGINX.URL: https://nginx.org (access date: 03-21-2021)
-   \bibitem{b13} Skorodumov A.
+   \bibitem{b5} W.~Jia, W.~Zhou. Distributed Network Systems: From Concepts to Implementations. Springer, USA, 2005, 513~p. ISBN: 9780387238395
+   \bibitem{b6} Burns B. Designing Distributed Systems: Patterns and Paradigms for Scalable, Reliable Services. O'Reilly Media, Inc., Japan, 2018, 149~p. ISBN:~9781491983614
+   \bibitem{b7} S.~Raval. Decentralized applications: harnessing Bitcoin's blockchain technology.  O'Reilly Media, Inc., USA, 2016. 103~p. ISBN:~9781491924549
+   \bibitem{b8} C.~A.~Sunshine. Computer Network Architectures and Protocols. Springer US, 2013. 558~p. ISBN:~9781461308096
+   \bibitem{b9} gRPC. URL:\url{https://grpc.io} (access date: 05-15-2021)
+   \bibitem{b10} OAuth Community Site. URL:\url{https://oauth.net} (access date: 05-15-2021)
+   \bibitem{b11} ORY Oathkeeper. GitHub. URL:\url{https://github.com/ory/oathkeeper} (access date: 05-15-2021)
+   \bibitem{b12} NGINX. URL:\url{https://nginx.org} (access date: 03-21-2021)
+   \bibitem{b13} A.~Skorodumov.
 Secure access to the company's information resources.
-Information Security No.
-2.
+Information Security No.~2.
 URL:\url{http://lib.itsec.ru/articles2/sys_ogr_dost/bezopasnyy-dostup-k-informatsionnym-resursam-kompanii} (access date: 05-15-2021)
-   \bibitem{b14} SSH Tunnel. SSH.COM URL:https://www.ssh.com/ssh/ tunneling/ (access date: 05-15-2021)
-   \bibitem{b15} Users. OpenSSH URL: https://www.openssh.com/users.html (access date: 05-15-2021)
-   \bibitem{b16} NTO (in Russian) URL: https://nti-contest.ru/ (access date: 05-20-2021)
-   \bibitem{b17} Educational intensive "Island 10-21". Information Bureau 20.35 (in Russian) URL: https://ntinews.ru/event/obrazovatelnyy-intensiv-ostrov-10-21.html (access date: 05-20-2021)
-   \bibitem{b18} Savvateev A.V., Filatov A.Yu.
+   \bibitem{b14} SSH Tunnel. URL:\url{https://www.ssh.com/ssh/ tunneling/} (access date: 05-15-2021)
+   \bibitem{b15} Users. OpenSSH URL:\url{https://www.openssh.com/users.html} (access date: 05-15-2021)
+   \bibitem{b16} NTO (in Russian) URL:\url{https://nti-contest.ru/} (access date: 05-20-2021)
+   \bibitem{b17} Educational intensive ``Island 10-21''. Information Bureau 20.35. (in Russian) URL:\url{https://ntinews.ru/event/obrazovatelnyy-intensiv-ostrov-10-21.html} (access date: 05-20-2021)
+   \bibitem{b18} A.~V.~Savvateev, A.~Yu.~Filatov A.Yu.
 Theory and practice of auctions.
@@
 Voronezh, 2018.
-No.
-3.
-P.
-119-131.
+No.~3.
+pp.~119-131.
 (in Russian)
-   \bibitem{b19} Myerson R. Optimal auction design. Mathematics of Operations Research. 1981. Выпуск 6. № 1, С. 58-73. https://doi.org/10.1287/moor.6.1.58
-   \bibitem{b20} RFC 6455 — The WebSocket Protocol. IETF URL: https://tools.ietf.org/html/rfc6455 (access date: 05-20-2021)
-   \bibitem{b21} Concise Binary Object Representation : [сайт]. 2021. URL: https://cbor.io/ (access date: 20.05.2021)
-   \bibitem{b22} Comparison of JSON Like Serializations. URL: http://zderadicka.eu/comparison-of-json-like-serializations-json-vs-ubjson-vs-messagepack-vs-cbor/ (access date: 21.05.2021)
-   \bibitem{b23} NixOS Linux. URL: https://nixos.org/ (access date: 05-15-2021)
-   \bibitem{b24} Galatenko V.A., Dzabraev M.D., Kostyukhin K.A.
+   \bibitem{b19} R.~Myerson. Optimal auction design. Mathematics of Operations Research. 1981. volume~6. No.~1, pp.~58-73. \doi{10.1287/moor.6.1.58}
+   \bibitem{b20} RFC 6455 — The WebSocket Protocol. IETF. URL:\url{https://tools.ietf.org/html/rfc6455} (access date: 05-20-2021)
+   \bibitem{b21} Concise Binary Object Representation. URL:\url{https://cbor.io/} (access date: 20.05.2021)
+   \bibitem{b22} Comparison of JSON Like Serializations. URL:\url{http://zderadicka.eu/comparison-of-json-like-serializations-json-vs-ubjson-vs-messagepack-vs-cbor/} (access date: 21.05.2021)
+   \bibitem{b23} NixOS Linux. URL:\url{https://nixos.org/} (access date: 05-15-2021)
+   \bibitem{b24} V.~A.~Galatenko, M.~D.~Dzabraev, K.~A.~Kostyukhin.
 Choosing a Package Manager for Multiversion Applications.
 Software products and systems.
-V.
-31, No.
-3, 2018.
-pp.
-469–474.
-(in Russian) DOI: 10.15827/0236-235X.123.469-474.45
-   \bibitem{b25} Haskell Language. : [сайт]. 2021. URL: https://haskell.org (access date: 20.05.2021)
-   \bibitem{b26} Dhall configuration language, The : [сайт]. 2021. URL: https://dhall-lang.org/ (access date: 21.05.2021)
-   \bibitem{b27} Bid Wars // Google Play [Электронный ресурс]. 2021. URL: https://play.google.com/store/apps/details?id=br.com.tapps.bid wars (access date: 05-15-2021)
-   \bibitem{b28} TimeZero. MMORPG Handbook. Auction. (in Russian) URL: http://web.archive.org/web/20160325215014/ https://www.timezero.ru/manual/auction.ru.html (access date: 05-15-2021)
-   \bibitem{b29} Sberbank-AST (in Russian) URL: http://www.sberbank-ast.ru/purchaseList.aspx (access date: 05-15-2021)
-   \bibitem{b30} Titz O. Why TCP Over TCP Is A Bad Idea. URL: http://sites.inka.de/bigred/devel/tcp-tcp.html (access date: 21.05.2021)
-   \bibitem{b31} Rapier C. et al. High performance SSH. SCP-HPN-SSH URL: http://www.psc.edu/networking/projects/hpn-ssh (access date: 21.05.2021)
-   \bibitem{b32} Rapier C., Bennett B.
+volume~31, No.~3, 2018.
+pp.~469–474.
+(in Russian) \doi{10.15827/0236-235X.123.469-474.45}
+   \bibitem{b25} Haskell Language. URL:\url{https://haskell.org} (access date: 20.05.2021)
+   \bibitem{b26} Theh Dhall configuration language. URL:\url{https://dhall-lang.org/} (access date: 21.05.2021)
+   \bibitem{b27} Bid Wars // Google Play [Электронный ресурс]. 2021. URL:\url{https://play.google.com/store/apps/details?id=br.com.tapps.bidwars} (access date: 05-15-2021)
+   \bibitem{b28} TimeZero. MMORPG Handbook. Auction. (in Russian) URLs:\url{http://web.archive.org/web/20160325215014/}, \url{https://www.timezero.ru/manual/auction.ru.html} (access date: 05-15-2021)
+   \bibitem{b29} Sberbank-AST (in Russian) URL:\url{http://www.sberbank-ast.ru/purchaseList.aspx} (access date: 05-15-2021)
+   \bibitem{b30} O.~Titz. Why TCP Over TCP Is A Bad Idea. URL:\url{http://sites.inka.de/bigred/devel/tcp-tcp.html} (access date: 21.05.2021)
+   \bibitem{b31} C.~Rapier et al. High performance SSH. SCP-HPN-SSH URL:\url{http://www.psc.edu/networking/projects/hpn-ssh} (access date: 21.05.2021)
+   \bibitem{b32} C.~Rapier, B.~Bennett.
 High speed bulk data transfer using the SSH protocol.
@@
 1-7.
-DOI: 10.1145/1341811.1341824
-   \bibitem{b33} Hobbs, S.
+\doi{10.1145/1341811.1341824}
+   \bibitem{b33} S.~Hobbs.
 CORS Tutorial: A Guide to Cross-Origin Resource Sharing.
 Auth0.
-URL: https://auth0.com/ blog/cors-tutorial-a-guide-to-cross-origin-resource-sharing/ (access date: 21.05.2021)
+URL:\url{https://auth0.com/blog/cors-tutorial-a-guide-to-cross-origin-resource-sharing/} (access date: 21.05.2021)
 \end{thebibliography}

### 2022-02-04 other/paper-aicts-chekan.tex
Reduced 1 page with tail suppression.

@@
 The current trends in energy sector organizations in the developed countries (including Russia) pay special attention to the concept of ``smart energy.'' The environmental and economic situation forces them to optimize energy processes, switch to alternative sources, and introduce smart grids [1].
-Information technologies play a special role in this process as a base for implementing a ``smart energy'' control.
+Information technologies play a special role here as implementers of a ``smart energy'' control.
 In this regard, the issue of training specialists with a comprehensive understanding of energy processes as being based on modern technologies is relevant.
@@
 One of the stages is auctions for the ownership of objects of the conditional microdistrict assembled by the participants of the power grid.
-The obtained grid is subsequently modeled for energy efficiency and profitability.
-Since the set of objects directly impacts the simulation result, it is crucial to reliably and promptly carry out this stage of work with the stand.
-The problem is the imperfection of the manual auction version when the host independently collects the participants' bids using material (paper) or digital media (messages in messenger applications).
-We proposed automating this work stage and introducing it into the existing software packages to reduce costs and improve reliability.
+The grid is subsequently simulated, accounting parameters of energy efficiency and profitability.
+Since the set of objects directly impacts the simulation result, it is crucial to carry out reliably and promptly this stage of work with the stand.
+The problem is the imperfection of the manual auction version when the hosts independently collect the participants' bids using material (paper) or digital media (messages in a messenger application).
+We proposed automating this stage and introducing it into the existing software packages to improve reliability.
 Power system modeling is an already essentially automated process.
-Reducing external dependencies through additional automation will increase the integrity of the software and hardware complex and increase the dynamism of the activities carried out on it, as well as reduce the requirements for operator qualification, which is especially important when deploying the stand as an educational tool at various sites where the direct presence of administrators and developers is not expected.
+Reducing external dependencies through additional automation will increase the integrity of the software and hardware complex and increase the dynamics of the activities, as well as reduce the requirements for operator qualification requirements, which is especially important when deploying the stand as an educational tool at various sites where the direct presence of administrators is not expected.
 
@@
 
-There is no consensus on the issue of organizing a network infrastructure, namely, connecting services, and most typically, services interact through open channels using standardized TCP and UDP protocols.
+There is no consensus on the issue of organizing a network infrastructure, namely, connecting services, and most typically, services interact through open channels using TCP and UDP protocols.
 This option provides sufficient throughput and, in the case of TCP, consistent two-way communication but imposes security and network infrastructure requirements.
@@
 There are ready-made solutions [11] that implement interaction using this protocol.
-Also, to protect applications from denial of service attacks, balancers are used, in particular, NGINX [12].
+To protect applications from denial of service attacks, balancers are used, \emph{e.g.}, NGINX [12].
 
-At the same time, there are scenarios where it is necessary to block access to the application from the external network to everyone except trusted nodes.
-It is especially true for corporate systems.
-To implement such scenarios, VPN tunneling, access via a terminal connection (for example, VNC), and web interface publishing via HTTPS [13] are often used, and the latter method is the most desirable, according to the author of the source.
-At the same time, VPN tunneling is considered the most popular and common method of creating a closed channel between nodes.
+There are scenarios where it is necessary to block access to the application from the external network to everyone except trusted nodes, namely, in corporate systems.
+To implement such scenarios, VPN tunneling, access via a terminal connection (for example, VNC), and web interface publishing via HTTPS [13] are often used, and the latter method prevails by author's opinion.
+VPN tunneling is considered the most popular and common method of creating a closed channel between nodes.
 However, its use is associated with some risks and requires proper configuration.
-The same applies to a web application, which will require additional labor to implement.
+The same applies to a web application.
 The terminal connection is used mainly for organizing user interaction with the desktop (VNC is used for this).
@@
 This experience has become the primary source of expertise in forming requirements for a new interpretation of the auction system in distributed scenarios.
-
-Three events made the dominant contribution to the expertise. This section will describe them and briefly analyze the user experience.
+Three events made the dominant contribution to the expertise.
+This section will describe them and briefly analyze the user experience.
 
@@
 
-Given the conditions for revenue equivalence, we can consider an all-pay auction to be equivalent to a sealed-bid first-price auction in terms of determining value.
+Given the conditions for revenue equivalence, we can consider an all-pay auction to be equivalent to a sealed-bid first-price auction in determining value.
 By bidding on the limit value, the participants set the personal value of the object, sacrificing its objective value (economic effect).
-To resolve the disputable situation, the personal value is determined again by bidding on the amount of money the participant is willing to give to own the object.
-Since the increase in this value negatively affects the final result of the participants, it was decided not to apply the "catch-up" bets rule in all-pay, holding a second round only if the leading bets completely match.
+To resolve the disputable situation, the personal value is determined again by bidding on the amount the participant is willing to give to own the object.
+Since the increase in this value negatively affects the final result of the participants, it was decided not to apply the ``catch-up'' bets rule in all-pay, holding a second round only if the leading bets completely match.
 
@@
 The system layer provides the infrastructure for connecting servers by SSH tunnel-based port forwarding.
-As mentioned, tunneling TCP connections via SSH does not require complex configuration.
-In particular, the OpenSSH server and client are included in the standard system software of a wide range of Linux distributions, and the ease of configuration makes this tool a suitable solution for building network applications with a distributed architecture.
-The construction of architecture on SSH tunnels is described in more detail in [7].
+Tunneling TCP connections via SSH does not require complex configuration.
+In particular, the OpenSSH server and client are included in the standard system software of a wide range of Linux distributions, and the ease of configuration makes this tool a suitable solution for building network applications with a distributed architecture [7].
 Since the SSH protocol supports key-based authorization, this method allows to set up a secure, stable communication channel between network nodes and explicitly define the connection topology.
@@
 Three applications constitute a three-tier client-server architecture in the software layer: the core (cloud level), the relay, and the mediator (terminal level).
-We consider the user interface a separate entity interacting with the mediator and the administrator interface interacting with a particular HTTP endpoint of the core.
+The user interface is a separate entity interacting with the mediator and the administrator interface interacting with a particular HTTP endpoint of the core.
 The core server controls the game session logic and accepts connections from the relays, updating the current session state for them and processing incoming commands from the players.
@@
 Thus, the tasks of authorization and identification are performed entirely by tunnels and the mediator.
-In the interface, it is necessary to implement only the display of data and the transmission of commands to the kernel through the mediator using a simple API.
+In the interface, we implemented only the display of data and the transmission of commands to the kernel through the mediator using an API.
 
@@
 
-In the current iteration, it is decided that player identification is carried out by mediator settings built into the system configuration.
-In the future, it is possible to implement authorization and identification by passwords or IP addresses (which will require eliminating tunnels between relays and mediators).
+As of today, player identification is carried out by mediator settings built into the system configuration.
+In the future, it is possible to implement authorization and identification by passwords or IP addresses (it requires eliminating tunnels between relays and mediators).
 
-All three programs are launched for greater stability at the system level as Systemd services configured to automatically restart upon a critical error.
-From the server-side, a sudden disconnection of the client should not incur problems because protocols provide state synchronization, and messages about the players’ actions are not necessary for the continuation of the game session.
+All three programs are launched at the system level as Systemd services configured to automatically restart upon a critical error.
+From the server-side, a sudden disconnection of a client should not incur problems because protocols provide state synchronization, and messages about the players’ actions are not necessary for the continuation of the game session.
 
@@
 The developed auction environment was tested at events as part of the third version of the STIES stand.
-From autumn 2019 to spring 2021, the powerstand was used at two all-Russian events, which are the finals of the "Intelligent Energy Systems" profile of the NTI Kruzhok Movement Olympiad (currently referred to as NTO) in 2020 and 2021, respectively.
+From autumn 2019 to spring 2021, the powerstand was used at two all-Russian events, which are the finals of the ``Intelligent Energy Systems'' profile of the NTI Kruzhok Movement Olympiad (referred to as NTO) in 2020 and 2021, respectively.
 These activities allowed to perform live testing of the auction environment in the conditions of intense engineering competition to personally observe the user experience, which made it possible to identify the weak and strong points of the developed system and identify areas for further research and development.
 
-\subsection{Distributed scenario}
+\paragraph{Distributed scenario}
 \label{sec:eval-distr}
-
 The final “Intelligent Energy Systems” profile of the NTO in 2020 was held from March 10 to 14.
@@
 
-\subsection{High-load local scenario}
+\paragraph{High-load local scenario}
 \label{sec:eval-high}
-
 From March 29 to April 3, 2021, Tyumen State University hosted the finals of the "Intelligent Energy Systems" profile of the NTO.
@@
 Investigated auction simulators are based on an entertaining game component and have no direct customization.
-One class is mobile games, e.g., Bid Wars [27], whereas auction simulators are frequently built into the existing game model as a side mechanic.
-Multiplayer online games [28] are examples of simulation.
+One class is mobile games, \emph{e.g}., Bid Wars~[27], whereas auction simulators are frequently built in the game model as a side mechanic.
+Multiplayer online games [28] are such examples.
 An aggravating factor in the compliance of analogs with the set requirements is a fixed model of conducting, most often corresponding to the English auction, when bids are announced openly, and the auction for the lot is conducted until there are no new bids.
@@
 
-Thus, the choice of architecture implementation at this stage is considered justified, taking all positive and negative factors into account.
-Nevertheless, it makes sense to apply developments to increase the speed and reduce delays in SSH tunneling [31, 32].
+Thus, the choice of architecture implementation at this stage is considered justified, taking all factors into account.
+Nevertheless, it makes sense to apply developments to increase the speed and reduce delays in SSH tunneling [31,32].
 Alternatively, even abandon SSH tunnels in favor of other transport protocols, such as a Secure WebSocket with a well-built authorization.
@@
 Although a near-second delay for JSON deserializing does not affect the auction’s perception and dynamics with a fast network channel, this is currently a bottleneck when transferring data to the interface.
-The solution to this problem could be to use CBOR instead of JSON or implement WebSockets communication, making it possible to eliminate the mediator as an extra intermediate link.
+The solution to this problem is to use CBOR instead of JSON or implement WebSockets communication, making it possible to eliminate the mediator as an extra intermediate link.
 Despite this, JavaScript has less performance than Haskell (being an interpreted language), and the use of a mediator for intermediate marshaling is justified at the current stage.

### 2022-02-04 other/paper-aicts-chekan.tex
Elimination of overfill box
Proof read of Introduction

@@
 The powerstand game session consists of stages.
-One of the stages is auctions for the ownership of objects of the conditional microdistrict assembled by the participants of the power grid.
+One of the stages is auctions for the ownership of objects of a playing ground district assembled by the participants of the power grid.
 The grid is subsequently simulated, accounting parameters of energy efficiency and profitability.
@@
 
-The first iteration was successfully developed and implemented as a powerstand module [4].
-However, as the popularity of virtual formats of educational and competitive events grows, it becomes necessary to implement a new version of the module for supporting the distributed environment of virtual auctions when participants work at stands located in different geographically distant points.
+The first version of software was successfully developed and implemented as a powerstand module~[4].
+However, as the popularity of virtual formats of educational and competitive events grows, it becomes necessary to implement an extension of the module for supporting the distributed environment of virtual auctions when participants work at stands located in different geographically distant points.
 The result of designing and implementing such an environment is presented in the research.
@@
 Building a network architecture with the expectation of implementation in consumer products involves following the standards for building computer networks that have been formed over the past half-century, including the OSI model and the protocol group of the transport and session layers of this model [8].
-These standards have been tested by time and practice, so the task of developing and designing new solutions belongs to the application and session levels, while solutions at the session level and below are most often the de facto standard, and working with them comes down to choosing options.
+These standards have been tested by time and practice, so the task of developing and designing new solutions belongs to the application and session levels, while solutions at the session level and below are most often the \emph{de facto} standard, and working with them comes down to choosing options.
 
@@
 240].
-In addition, replication solves two more problems - increasing the availability of the service (if one node fails, the functionality supports the other) and improving performance (reducing response time by dividing the data flow across different servers).
+In addition, replication solves two more problems - increasing the availability of the service (if one node fails, the functionality supports the other) and improving performance (reducing response time by distributing the data flow across different servers).
 Although these provisions are described mainly for decentralized systems, replication in a centralized architecture (if we consider working with data separately) also allows them to be achieved.
@@
 It is worth highlighting libraries for implementing remote procedure call (RPC), particularly gRPC [9], a full-fledged cross-platform solution for distributed computing.
-However, the most common solutions related to remote message passing (message passing) require the connection of additional libraries that carry additional dependencies.
+However, the most common solutions related to remote message passing require utilization of other libraries that pull additional dependencies.
 It makes sense to write code for network functionality optimized for specific conditions and requirements depending on the task.
@@
    \bibitem{b21} Concise Binary Object Representation. URL:\url{https://cbor.io/} (access date: 20.05.2021)
-   \bibitem{b22} Comparison of JSON Like Serializations. URL:\url{http://zderadicka.eu/comparison-of-json-like-serializations-json-vs-ubjson-vs-messagepack-vs-cbor/} (access date: 21.05.2021)
+   \bibitem{b22} Comparison of JSON Like Serializations. URL:\texttt{http://zderadicka.eu/comparison-of-json-like-serializations-json-vs-ubjson-vs-messagepack-vs-cbor/} (access date: 21.05.2021)
    \bibitem{b23} NixOS Linux. URL:\url{https://nixos.org/} (access date: 05-15-2021)

### 2022-02-04 other/paper-aicts-chekan.tex
chekan: added M's contacts, fixed a header and access dates

--- 2022-02-04 other/paper-aicts-chekan.tex
@@
 \author[1,2]{Mikhail A. Chekan}[%
-orcid=0000-0003-2428-2471,
-email=eugeneai@icc.ru,
-url=https://github.org/eugeneai,
+orcid=0000-0001-7622-421X,
+email=chekoopa@mail.ru,
+url=https://github.org/chekoopa,
 ]
@@
 
-Game model (перевести картинку)
+\section{Game model}
+\label{sec:gamemodel}
 
@@
 \begin{thebibliography}{99}
-   \bibitem{b1} Business for society: how Siemens is changing life in Russia. Smart Energy. RBC, 2016 (in Russian). URL:\url{http://siemens.rbc.ru/article1.html} (access date: 05-15-2021)
-   \bibitem{b2} E.~B.~Makarov. Game forms of education. Open lesson. (in Russian) URL:\url{https://openlesson.rf/articles/556327/} (access date: 05-15-2021)
+   \bibitem{b1} Business for society: how Siemens is changing life in Russia. Smart Energy. RBC, 2016 (in Russian). URL:\url{http://siemens.rbc.ru/article1.html} (access date: 11-22-2021)
+   \bibitem{b2} E.~B.~Makarov. Game forms of education. Open lesson. (in Russian) URL:\url{https://urok.1sept.ru/articles/556327/} (access date: 11-22-2021)
    \bibitem{b3} Intelligent energy systems.
 Stand-simulator for programming and management of urban power networks.
-(in Russian) URL:\url{http://polyus-nt.ru/powerstand.html} (access date: 11/22/2021).
+(in Russian) URL:\url{http://polyus-nt.ru/powerstand.html} (access date: 11-22-2021).
    \bibitem{b4} M.~A.~Chekan.
@@
    \bibitem{b8} C.~A.~Sunshine. Computer Network Architectures and Protocols. Springer US, 2013. 558~p. ISBN:~9781461308096
-   \bibitem{b9} gRPC. URL:\url{https://grpc.io} (access date: 05-15-2021)
-   \bibitem{b10} OAuth Community Site. URL:\url{https://oauth.net} (access date: 05-15-2021)
-   \bibitem{b11} ORY Oathkeeper. GitHub. URL:\url{https://github.com/ory/oathkeeper} (access date: 05-15-2021)
+   \bibitem{b9} gRPC. URL:\url{https://grpc.io} (access date: 11-22-2021)
+   \bibitem{b10} OAuth Community Site. URL:\url{https://oauth.net} (access date: 11-22-2021)
+   \bibitem{b11} ORY Oathkeeper. GitHub. URL:\url{https://github.com/ory/oathkeeper} (access date: 11-22-2021)
    \bibitem{b12} NGINX. URL:\url{https://nginx.org} (access date: 03-21-2021)
@@
 Information Security No.~2.
-URL:\url{http://lib.itsec.ru/articles2/sys_ogr_dost/bezopasnyy-dostup-k-informatsionnym-resursam-kompanii} (access date: 05-15-2021)
-   \bibitem{b14} SSH Tunnel. URL:\url{https://www.ssh.com/ssh/ tunneling/} (access date: 05-15-2021)
-   \bibitem{b15} Users. OpenSSH URL:\url{https://www.openssh.com/users.html} (access date: 05-15-2021)
-   \bibitem{b16} NTO (in Russian) URL:\url{https://nti-contest.ru/} (access date: 05-20-2021)
-   \bibitem{b17} Educational intensive ``Island 10-21''. Information Bureau 20.35. (in Russian) URL:\url{https://ntinews.ru/event/obrazovatelnyy-intensiv-ostrov-10-21.html} (access date: 05-20-2021)
+URL:\url{http://lib.itsec.ru/articles2/sys_ogr_dost/bezopasnyy-dostup-k-informatsionnym-resursam-kompanii} (access date: 11-22-2021)
+   \bibitem{b14} SSH Tunnel. URL:\url{https://www.ssh.com/ssh/ tunneling/} (access date: 11-22-2021)
+   \bibitem{b15} Users. OpenSSH URL:\url{https://www.openssh.com/users.html} (access date: 11-22-2021)
+   \bibitem{b16} NTO (in Russian) URL:\url{https://ntcontest.ru/} (access date: 11-22-2021)
+   \bibitem{b17} Educational intensive ``Island 10-21''. Information Bureau 20.35. (in Russian) URL:\url{https://ntinews.ru/event/obrazovatelnyy-intensiv-ostrov-10-21.html} (access date: 11-22-2021)
    \bibitem{b18} A.~V.~Savvateev, A.~Yu.~Filatov A.Yu.
@@
    \bibitem{b19} R.~Myerson. Optimal auction design. Mathematics of Operations Research. 1981. volume~6. No.~1, pp.~58-73. \doi{10.1287/moor.6.1.58}
-   \bibitem{b20} RFC 6455 — The WebSocket Protocol. IETF. URL:\url{https://tools.ietf.org/html/rfc6455} (access date: 05-20-2021)
-   \bibitem{b21} Concise Binary Object Representation. URL:\url{https://cbor.io/} (access date: 20.05.2021)
-   \bibitem{b22} Comparison of JSON Like Serializations. URL:\url{http://zderadicka.eu/comparison-of-json-like-serializations-json-vs-ubjson-vs-messagepack-vs-cbor/} (access date: 21.05.2021)
-   \bibitem{b23} NixOS Linux. URL:\url{https://nixos.org/} (access date: 05-15-2021)
+   \bibitem{b20} RFC 6455 — The WebSocket Protocol. IETF. URL:\url{https://tools.ietf.org/html/rfc6455} (access date: 12-22-2021)
+   \bibitem{b21} Concise Binary Object Representation. URL:\url{https://cbor.io/} (access date: 11-22-2021)
+   \bibitem{b22} Comparison of JSON Like Serializations. URL:\url{http://zderadicka.eu/comparison-of-json-like-serializations-json-vs-ubjson-vs-messagepack-vs-cbor/} (access date: 11-22-2021)
+   \bibitem{b23} NixOS Linux. URL:\url{https://nixos.org/} (access date: 11-22-2021)
    \bibitem{b24} V.~A.~Galatenko, M.~D.~Dzabraev, K.~A.~Kostyukhin.
@@
 (in Russian) \doi{10.15827/0236-235X.123.469-474.45}
-   \bibitem{b25} Haskell Language. URL:\url{https://haskell.org} (access date: 20.05.2021)
-   \bibitem{b26} Theh Dhall configuration language. URL:\url{https://dhall-lang.org/} (access date: 21.05.2021)
-   \bibitem{b27} Bid Wars // Google Play [Электронный ресурс]. 2021. URL:\url{https://play.google.com/store/apps/details?id=br.com.tapps.bidwars} (access date: 05-15-2021)
-   \bibitem{b28} TimeZero. MMORPG Handbook. Auction. (in Russian) URLs:\url{http://web.archive.org/web/20160325215014/}, \url{https://www.timezero.ru/manual/auction.ru.html} (access date: 05-15-2021)
-   \bibitem{b29} Sberbank-AST (in Russian) URL:\url{http://www.sberbank-ast.ru/purchaseList.aspx} (access date: 05-15-2021)
-   \bibitem{b30} O.~Titz. Why TCP Over TCP Is A Bad Idea. URL:\url{http://sites.inka.de/bigred/devel/tcp-tcp.html} (access date: 21.05.2021)
-   \bibitem{b31} C.~Rapier et al. High performance SSH. SCP-HPN-SSH URL:\url{http://www.psc.edu/networking/projects/hpn-ssh} (access date: 21.05.2021)
+   \bibitem{b25} Haskell Language. URL:\url{https://haskell.org} (access date: 11-22-2021)
+   \bibitem{b26} Theh Dhall configuration language. URL:\url{https://dhall-lang.org/} (access date: 11-22-2021)
+	\bibitem{b27} Bid Wars // Google Play [Электронный ресурс]. 2021. URL:\url{https://play.google.com/store/apps/details?id=br.com.tapps.bidwars} (access date: 11-22-2021)
+   \bibitem{b28} TimeZero. MMORPG Handbook. Auction. (in Russian) URLs:\url{http://web.archive.org/web/20160325215014/}, \url{https://www.timezero.ru/manual/auction.ru.html} (access date: 11-22-2021)
+   \bibitem{b29} Sberbank-AST (in Russian) URL:\url{http://www.sberbank-ast.ru/purchaseList.aspx} (access date: 11-22-2021)
+   \bibitem{b30} O.~Titz. Why TCP Over TCP Is A Bad Idea. URL:\url{http://sites.inka.de/bigred/devel/tcp-tcp.html} (access date: 11-22-2021)
+   \bibitem{b31} C.~Rapier et al. High performance SSH. SCP-HPN-SSH URL:\url{http://www.psc.edu/networking/projects/hpn-ssh} (access date: 11-22-2021)
    \bibitem{b32} C.~Rapier, B.~Bennett.
@@
 Auth0.
-URL:\url{https://auth0.com/blog/cors-tutorial-a-guide-to-cross-origin-resource-sharing/} (access date: 21.05.2021)
+URL:\url{https://auth0.com/blog/cors-tutorial-a-guide-to-cross-origin-resource-sharing/} (access date: 11-22-2021)
 \end{thebibliography}

### 2022-02-04 other/paper-aicts-chekan.tex

--- 2022-02-04 other/paper-aicts-chekan.tex
@@
    \bibitem{b22} Comparison of JSON Like Serializations. URL:\texttt{http://zderadicka.eu/comparison-of-json-like-serializations-json-vs-ubjson-vs-messagepack-vs-cbor/} (access date: 21.05.2021)
-   \bibitem{b23} NixOS Linux. URL:\url{https://nixos.org/} (access date: 05-15-2021)
+   \bibitem{b23} NixOS Linux. URL:\url{https://nixos.org/} (access date: 11-22-2021)
    \bibitem{b24} V.~A.~Galatenko, M.~D.~Dzabraev, K.~A.~Kostyukhin.

### 2022-02-04 other/paper-aicts-chekan.tex
Update

@@
 
-The auction, as mentioned above, the module has become a prototype of the distributed environment.
+The aforementioned auction module has become a prototype of the distributed environment.
 Being a part of a powerstand software, the module is used in educational institutions, engineering competitions, and intensive programs.
@@
 The rules of the event included practice games both in distributed and local formats.
-Local format implied the core services (incl.
-the auction module) to be placed in the stand instead of a cloud server.
+Local format implied the core services (including the auction module) to be placed in the stand instead of a cloud server.
 A service manager was implemented to set up and monitor applications based on a configuration file to perform mode switching.

### 2022-02-04 other/paper-aicts-chekan.tex
Update

@@
 
-After the first auction pass, when the objects are pre-assigned, participants can refuse a certain number of their lots, putting them on the second pass of the auction, along with those that were unplaced on the first.
+After the first auction pass, when the objects are pre-assigned, participants can refuse a certain number of their lots, putting them on the second pass of the auction, along those were unplaced on the first.
 In case of refusal of a lot, the participant loses the right to compete for this lot.
@@
 At the end of the auction, the results tally is generated to represent information about the ownership of the objects, the bids set on them, and the submitted all-pay bids.
-
 According to the game model described above, a state diagram is drawn, shown in Figure~\ref{fig:1}.
-
 \begin{figure}
@@
 \end{figure}
-
 This model allows simultaneous independent draw of several lots.
@@
 Therefore, the model enables solving other tasks related to the distribution of virtual resources.
-It is also possible to change the model implementation, if necessary, i.e., add additional rules.
+It is also possible to change the model implementation, if necessary, \emph{i.e}., add additional rules.
 
@@
 
-To minimize network traffic between levels, atomic state changes (denoted as effects) are transmitted instead of the new state to minimize network traffic between levels.
+To minimize network traffic between levels, atomic state changes (hereafter denoted as \emph{effects}) are transmitted instead of the new state to minimize network traffic between levels.
 An effect is generated for each event and passed to all child nodes.

### 2022-02-04 other/paper-aicts-chekan.tex
The final update, hopefully

@@
 
-\paragraph{Distributed scenario}
+\paragraph{Distributed scenario.}
 \label{sec:eval-distr}
@@
 
-\paragraph{High-load local scenario}
+\paragraph{High-load local scenario.}
 \label{sec:eval-high}
-From March 29 to April 3, 2021, Tyumen State University hosted the finals of the "Intelligent Energy Systems" profile of the NTO.
-10 teams took part in this final.
+From March 29 to April 3, 2021, Tyumen State University hosted the finals of the ``Intelligent Energy Systems'' profile of the NTO.
+Ten teams took part in this final.
 But due to organizational difficulties, all teams were placed on the same site.
@@
 
-Being based on SSH tunnels, the architecture configuration is static at the system level, and connections are made according to the “allow list” principle, which is why any change in the architecture requires the action of developers.
+Being based on SSH tunnels, the architecture configuration is static at the system level, and connections are made according to the “allow list” principle, that is why any change in the architecture requires the interference of developers.
 At the same time, authentication is currently absent, and the player is determined through the mediator configuration based on the trust of the nodes for each other.

### 2022-04-15 paper-aicts-jcart.tex
JCTart compliance

--- 2022-02-04 paper-aicts-cherkashin.tex
@@
 %% hf: enable header and footer.
-\documentclass[
-% twocolumn,
-% hf,
-]{ceurart}
-\usepackage{amsmath,amssymb,longtable,hhline}
+\documentclass[a4paper]{jctart19a}
+
+\oddsidemargin=-1.5mm \evensidemargin=-2.5mm
+
+\usepackage[pdftex]{graphicx}
+\usepackage{epstopdf}
+\usepackage{amsthm,amsfonts,amsmath,amssymb}
+
+\usepackage{color}
+\usepackage{multirow}
+
+\newcommand{\jctyear}{????}
+\newcommand{\jctvolume}{??}
+\newcommand{\jctnumber}{??}
 \usepackage{doi}
 \def\doitext{DOI:}
+\newcommand{\jctdoi}{\doitext{}10.25743/ICT.\jctyear.\jctvolume.\jctnumber.???}
+
+% \usepackage{amsmath,amssymb,longtable,hhline}
 
@@
 %% One can fix some overfills
-\sloppy
+% \sloppy
 
@@
 
-\hypersetup{
-    bookmarks=true,         % show bookmarks bar?
-    unicode=true,           % non-Latin characters in Acrobat’s bookmarks
-    pdftoolbar=false,        % show Acrobat’s toolbar?
-    pdfmenubar=false,        % show Acrobat’s menu?
-    pdffitwindow=false,     % window fit to page when opened
-    pdfstartview={FitH},    % fits the width of the page to the window
-    pdftitle={},    % title
-    pdfauthor={Evgeny Cherkashin},     % author
-    pdfsubject={Technologies of Semantic WEB as an environment of application development and integration},   % subject of the document
-    pdfnewwindow=true,      % links in new PDF window
-    colorlinks=true,       % false: boxed links; true: colored links
-    linkcolor=black,          % color of internal links (change box color with linkbordercolor)
-    citecolor=black,        % color of links to bibliography
-    filecolor=black,      % color of file links
-    urlcolor=black           % color of external links
-  }
+% \hypersetup{
+%     bookmarks=true,         % show bookmarks bar?
+%     unicode=true,           % non-Latin characters in Acrobat’s bookmarks
+%     pdftoolbar=false,        % show Acrobat’s toolbar?
+%     pdfmenubar=false,        % show Acrobat’s menu?
+%     pdffitwindow=false,     % window fit to page when opened
+%     pdfstartview={FitH},    % fits the width of the page to the window
+%     pdftitle={},    % title
+%     pdfauthor={Evgeny Cherkashin},     % author
+%     pdfsubject={Technologies of Semantic WEB as an environment of application development and integration},   % subject of the document
+%     pdfnewwindow=true,      % links in new PDF window
+%     colorlinks=true,       % false: boxed links; true: colored links
+%     linkcolor=black,          % color of internal links (change box color with linkbordercolor)
+%     citecolor=black,        % color of links to bibliography
+%     filecolor=black,      % color of file links
+%     urlcolor=black           % color of external links
+%   }
 \urlstyle{tt}
@@
 
-%%
-%% Rights management information.
-%% CC-BY is default license.
-\copyrightyear{2021}
-\copyrightclause{Copyright for this paper by its authors.
-  Use permitted under Creative Commons License Attribution 4.0
-  International (CC BY 4.0).}
+\setcounter{page}{1}
 
-%%
-%% This command is for the conference information
-\conference{2nd International Workshop on Advanced Information and Computation Technologies and Systems, December 6--10, 2021, Irkutsk, Russia}
+\widowpenalty=10000
+\clubpenalty=10000
+
+\label{pagebegin}
+
+\vspace*{-20mm}
+
+\noindent{{\scriptsize Вычислительные технологии, \jctyear, том \jctvolume, № \jctnumber, с.~\pageref*{pagebegin}--\pageref*{pageend}. \copyright\ ФИЦ ИВТ, \jctyear\hfill ISSN 1560-7534\\[-1.5mm]
+Computational Technologies, \jctyear,  vol. \jctvolume, no. \jctnumber, pp.~\pageref*{pagebegin}--\pageref*{pageend}. \copyright\ FRC ICT, \jctyear\hfill eISSN 2313-691X}}\\[-3mm]
+
+{\centering
+\noindent \rule{16.49cm}{0.5mm}\\
+{\small ??? }\\[-2.5mm]
+\noindent \rule{16.49cm}{0.5mm}}\\[3mm]
+
+\markboth{<И.\,О.~Фамилия первого автора>, <И.\,О.~Фамилия остальных авторов>}{<Сокращенное название статьи (до 50 символов)>}
+\vspace*{-7.5mm}
+
+\begin{flushright}
+	{\small \jctdoi}
+\end{flushright}
 
@@
 %% the authors and their affiliations.
-\author[1,2]{Evgeny A. Cherkashin}[%
-orcid=0000-0003-2428-2471,
-email=eugeneai@icc.ru,
-url=https://github.org/eugeneai,
-]
-\address[1]{Matrosov Institute for System Dynamics and Control Theory of Siberian Branch of Russian Academy of Sciences, 134 Lermontov St, Irkutsk, 664033, Russian Federation}
+\author{%
+\vspace*{-3mm}
+	{\sc Evgeny A. Cherkashin$^{1,2,\ast}$}\\
+	$^1$\vtop{\small Matrosov Institute for System Dynamics and Control Theory of Siberian Branch of Russian Academy of Sciences, 134 Lermontov St, 664033, Irkutsk, Russian Federation}\\
+	$^2$\vtop{\small Institute for Mathematics and Information Technologies, Irkutsk State University, 20~Gagarina Bulv, 664003, Irkutsk, Russian Federation>}\\
+	$^\ast$\vtop{\small Контактный автор: Evgeny A. Cherkashin, e-mail: \tt eugeneai\@icc.ru}\\
+	{\scriptsize\it Поступила ?? ??? ????~г., доработана ?? ??? ????~г., принята в печать ?? ??? ????~г.}
+}
 
-\address[2]{Institute for Mathematics and Information Technologies, Irkutsk State University, 20~Gagarina Bulv, Irkutsk, 664003, Russian Federation}
+\date{}\maketitle
 
@@
 Three examples show the technologies usage, exposing our experience.
-\end{abstract}
+% \end{abstract}
 
@@
 %% the work being presented. Separate the keywords with commas.
-\begin{keywords}
+\emph{Keywords:}
   knowledge graph \sep
@@
   geographical information system
-\end{keywords}
+
+\emph{Citing:} Evgeny A.
+Cherkashin.
+Technologies of Semantic WEB as an environment of application development and integration.
+Вычислительные технологии.
+\jctyear; \jctvolume(\jctnumber):\pageref*{pagebegin}--\pageref*{pageend}.\linebreak \jctdoi.
+\end{abstract}
 
@@
 %% information and builds the first part of the formatted document.
-\maketitle
 
@@
 
-\begin{figure}[bth]
+\begin{figure}[!b]
   \centering
@@
 Our R\&D resulted in a set of components and a general application architecture showed in Figure~\ref{fig:arch-services}.
-\begin{figure}[bth]
+\begin{figure}[!b]
   \begin{center}
@@
 Figure~\ref{fig:block-seq} shows the main idea behind PSM object synthesis.
-\begin{figure}
+\begin{figure}[!b]
   \centering
@@
 Arguments of \verb|taa:content| are subcontexts (\verb|title| and \verb|neg-UMK|) of a current template under rendering.
-\begin{figure}
+\begin{figure}[!b]
   \centering
@@
 In Figure~\ref{fig:gis-ex} an example of the filtering and a selected object data viewing form is presented.
-\begin{figure}
+\begin{figure}[!b]
   \centering
@@
 
-\begin{acknowledgments}
+%%%%%%%%% Конец текста статьи %%%%%%%%%%%%%
+\pagebreak
 
-The results were obtained within the state assignment of the Ministry of Education and Science of Russia, the project ``Methods and technologies of a cloud-based service-oriented digital platform for collecting, storing and processing large volumes of multi-format interdisciplinary data and knowledge based on the use of artificial intelligence, a model-driven approach, and machine learning'', No.~FWEW-2021-0005 (State registration No.~121030500071-2).
+\paragraph{\small Acknowledgements.}\hspace*{-3.5mm} The results were obtained within the state assignment of the Ministry of Education and Science of Russia, the project ``Methods and technologies of a cloud-based service-oriented digital platform for collecting, storing and processing large volumes of multi-format interdisciplinary data and knowledge based on the use of artificial intelligence, a model-driven approach, and machine learning'', No.~FWEW-2021-0005 (State registration No.~121030500071-2).
 
@@
 
-\end{acknowledgments}
 %%

### 2022-04-16 paper-aicts-jcart.tex
Cut crying the images and text.

@@
 % \usepackage{listings}
-\usepackage{minted}
-\setminted{fontsize=\footnotesize,mathescape}
+% \usepackage{minted}
+% \setminted{fontsize=\footnotesize,mathescape}
 \usepackage{hyperref}
@@
 
-\markboth{<И.\,О.~Фамилия первого автора>, <И.\,О.~Фамилия остальных авторов>}{<Сокращенное название статьи (до 50 символов)>}
+\en
+\markboth{Evgeny A. Cherkashin}{Semantic WEB in application development and integration}
 \vspace*{-7.5mm}
@@
 	$^1$\vtop{\small Matrosov Institute for System Dynamics and Control Theory of Siberian Branch of Russian Academy of Sciences, 134 Lermontov St, 664033, Irkutsk, Russian Federation}\\
-	$^2$\vtop{\small Institute for Mathematics and Information Technologies, Irkutsk State University, 20~Gagarina Bulv, 664003, Irkutsk, Russian Federation>}\\
-	$^\ast$\vtop{\small Контактный автор: Evgeny A. Cherkashin, e-mail: \tt eugeneai\@icc.ru}\\
+	$^2$\vtop{\small Institute for Mathematics and Information Technologies, Irkutsk State University, 20~Gagarina Bulv, 664003, Irkutsk, Russian Federation}\\
+    \ru
+	$^\ast$\vtop{\small Контактный автор: Evgeny A. Cherkashin, e-mail: \tt eugeneai@icc.ru}\\
 	{\scriptsize\it Поступила ?? ??? ????~г., доработана ?? ??? ????~г., принята в печать ?? ??? ????~г.}
@@
 
+\en
 %%
@@
   Semantic Web technologies and their standards are becoming productive assets for software development and integration on various design and implementation levels.
-These include OSI level 6 data representation, database and file storage formats, the source data for text indexes, document authoring, hypertext markup engines, a representation for abstract software models, \emph{etc.}  They are also smoothly combined with the development and data analysis tools, giving rise to an environment for integrating distributed applications on a semantic level.
+These include OSI level 6 data representation, database and file storage formats, %the source data for text indexes, document authoring, hypertext markup engines, a representation for abstract software models,
+\emph{etc.}  They are also smoothly combined with the development and data analysis tools, giving rise to an environment for integrating distributed applications on a semantic level.
 
-  This paper presents a view on the Semantic Web technologies as data representation for software development tools with model analysis, transformation, and retrospection capabilities.
-Three examples show the technologies usage, exposing our experience.
+This paper presents a view on the technologies as data representation for software development tools based on model analysis, and transformation, %and retrospection capabilities,
+examples are provided.  %Three examples show the technologies usage, exposing our experience.
 % \end{abstract}
@@
 %% the work being presented. Separate the keywords with commas.
+\providecommand{\sep}{\unskip, }
 \emph{Keywords:}
@@
   model transformation \sep
-  model-driven architecture \sep
+  MDA \sep
   object-oriented logical programming \sep
   web application \sep
-  geographical information system
+  GIS
 
@@
 Technologies of Semantic WEB as an environment of application development and integration.
-Вычислительные технологии.
-\jctyear; \jctvolume(\jctnumber):\pageref*{pagebegin}--\pageref*{pageend}.\linebreak \jctdoi.
+\ru Вычислительные технологии.
+\jctyear; \jctvolume(\jctnumber):\pageref*{pagebegin}--\pageref*{pageend}.
+\en \linebreak \jctdoi.
 \end{abstract}
@@
 Some tools for automation of conceptual model design and model conversion were proposed \cite{hogan}.
-The R\&D at present are carried on in developing applications, improving and standardizing vocabularies (conceptual domain models), devising tools for solving particular problems.
+The R\&D at present are carried on in developing applications, improving and standardizing vocabularies (conceptual domain models).
+%, devising tools for solving particular problems.
 
@@
 These include data representation formats, vocabulary standards, query languages, tools for accessing and managing data in a KG.
-Software development techniques are mediated with properties of the new-created resources and their APIs\footnote{Abbreviation of Application Programming Interface}.
-Many internet resources provide standard ways of formalized domain data access using query endpoints, and these endpoints allow one to organize federated KG data access.
+Software development techniques are mediated with properties of the new-created resources and their APIs (Application Programming Interfaces).
+Many internet resources provide standard ways of formalized domain data access using query endpoints.%, and these endpoints allow one to organize federated KG data access.
 
-In Russia, domain model designers generally prefer to develop local vocabularies instead of spending time for the routine work of Internet surfing for existing conceptual models and their enrichment.
+In Russia, domain model designers generally prefer to develop local vocabularies instead of spending time for the routine work of internet surfing for existing conceptual models and their enrichment.
 It is due to underdevelopment of the metadata resources (metavocabularies) due to lack of a domain classification system as compared, \emph{e.g.}, to Universal Decimal Classification (UDC) used in the bibliography.
-Vocabularies also can have common parts, describing the same or similar entities with different terms and relations, in particular, semantically more adapted to a specific problem-solving.
+Vocabularies also can have common parts, describing the same or similar entities with different terms and relations.%, in particular, semantically more adapted to a specific problem-solving.
 
-Unitizing standardized vocabularies as conceptual domain models (ontologies) in our projects allows us to compare other research groups' points of view to a problem with a formalized context and figure out underestimated and not fully formularized aspects of a domain under investigation or automation.
+Unitizing standardized vocabularies (ontologies) in our projects allows us to compare other research groups' points of view to a problem with a formalized context and figure out underestimated and not fully formularized aspects of a domain under investigation or automation.
 The vocabularies are similar to information sources as scientific papers but have a concrete formal presentation of the information objects, allowing one to get rid of particular uncertainties.
 Another advantage of standardization is constructing a common ground of research, development, data formalization, and interchange formats and semantics.
-It includes an aggregation of pattern recognition results applications of machine learning with their following use as facts in knowledge-based systems.
+%It includes an aggregation of pattern recognition results applications of machine learning with their following use as facts in knowledge-based systems.
 
-This paper aims to present a contemporary experience of SW and KG utilization in the development of applications in a context of performance improvement of R\&D.
+This paper presents an author's experience of SW and KG utilization in the development of applications in a context of performance improvement of R\&D.
 Three examples were reviewed, showing good points of the technologies' application.
@@
 Briefly, SW technologies from a developer point of view include the following assets:
+% Es sheist dass diest Enumrate ist Itemize????
 \begin{enumerate}
-\item Global identification of the entities via URI\footnote{Abbreviation of Universal Resource Identifiers}, IRI\footnote{Abbreviation of International Resource Identifier}, URL\footnote{Abbreviation of Universal Resource Locator}, which are essentially the same notions but reflecting different representation properties of an identified resource;
-\item Conceptual set-theoretic and category-theoretic data representation notation as relations between resources and literals represented with triples of a form \texttt{<subject, predicate, object>};
-\item Triples used in a description of a domain form a corresponding semantic graph, conditionally subdivided into terminological subgraph (T-Box) and instance subgraph (A-Box);
+\item Global identification of the entities via URI, IRI, URL (\{Universal, International\} Resource \{Identifier, Locator\}) are essentially the same notions but reflecting different representation properties of an identified resource;
+\item Conceptual set-theoretic and category-theoretic data representation notation as relations between resources and literals represented % with triples of a form
+as \texttt{<subject, predicate, object>};
+\item Triples used in a description of a domain form a corresponding semantic graph, conditionally subdivided into terminological subgraph (T-Box) and instances %subgraph
+(A-Box);
 \item Graph presentation formats and techniques are implemented in files, databases, markups of documents, a data flow between applicants; \label{kg1}
-\item There are public semantic databases comprising the generic domain objects descriptions, which are mostly A-Boxes (databases), for example, \texttt{DBPedia.org}, a formalized version of \texttt{Wikipedia.org}.
-% wikidata
+\item There are %public semantic
+databases comprising the generic domain objects descriptions, which are mostly A-Boxes (databases), for example, \texttt{DBPedia.org}, (formalized \texttt{Wikipedia.org}). % wikidata
 \item For graphs being databases, query services are implemented to allow users to obtain its data by portions, as well as implement modifications;
-\item API and libraries for processing files and documents with semantic data acquisition (\emph{e.g.} by extraction of the markup).
-\item Software for graph data processing to improve its structure or induct general relation patterns;
+\item API and libraries for processing files and documents with semantic data acquisition.% (\emph{e.g.} by extraction of the markup).
+\item Software for graph data processing to improve its structure or induct %general relation
+patterns;
 \item Validation and verification services allow one to explore consistency and completeness of the conceptual domain models and their implementations in software. \label{kg2}
 \end{enumerate}
-The assets  \ref{kg1}--\ref{kg2} form knowledge graph technological aspect of the SW that gained a particular interest within R\&D in the last decade.
+Some of the assets % \ref{kg1}--\ref{kg2}
+form knowledge graph (KG) technological aspect of the SW that gained a particular interest within R\&D in the last decade.
 
@@
 Namespaces are convenient instruments in the triple definition.
-For example \verb|integer| literal \verb|42^^<http://www.w3.org/2001/XMLSchema#integer>| can be reduced to \verb|42^^xs:integer|, where \verb|xs| is namespace prefix of a vocabulary having term \verb|integer|.
+For example \verb|integer| literal \verb|42^^<http://www.w3.org/2001/XMLSchema#integer>| is reduced to \verb|42^^xs:integer|, where \verb|xs| is namespace prefix of a vocabulary having term \verb|integer|.
 Most KG representation formats allow one to define a default namespace used without prefix.
@@
 
-Access by program agents to the global resources are standardized as well.  Each serious vocabulary have API access point (endpoints) of form
+Program agents' access to the global resources is standardized as well.  Serious vocabulary providers have an API endpoint of form
 \verb|http[s]://<address>/sparql|\footnote{For example, see DBpedia access point at the URL \url{https://dbpedia.org/sparql}} supporting direct \texttt{GET}-queries returning a user form, and \texttt{SPARQL} \texttt{POST} query with answering \texttt{JSON} encoded tabular result.
@@
 
-KGs and subgraphs are programmatically processed using libraries developed for all papular programming environments.
+% Me'sa crying.... cutting this.
+%KGs and subgraphs are programmatically processed using libraries developed for all papular programming environments.
 KG is represented with a triple or quadruple set object.
@@
 Filters are usually defined as patterns, where each component is either a free variable/placeholder or a concrete value (IRI, literal), which must match exactly.
-Some libraries allow one to execute \texttt{SPARQL} queries locally.
-KG is evolved by adding new triples and removing old ones.
-The resulting modified KG is stored as files and published on a server.
-An example of a federated SPARQL query is presented in Figure~\ref{fig:sparql-ex1}, which results in counting parts classes as the first level of a three widget.
+Some libraries allow one to execute \texttt{SPARQL} queries locally and then propagate to other endpoints, resulting in a federated query.
+% KG is evolved by adding new triples and removing old ones.
+% The resulting modified KG is stored as files and published on a server.
+% An example of a federated SPARQL query is presented in Figure~\ref{fig:sparql-ex1}, which results in counting parts classes as the first level of a three widget.
 Each class is named in Russian if \texttt{DBPedia} contains a corresponding label with \texttt{ru} language tag.
@@
 
-\begin{figure}[!b]
-  \centering
-\begin{minted}{sparql}
-PREFIX dbr: <http://dbpedia.org/resource/>
-PREFIX owl: <http://www.w3.org/2002/07/owl#>
-PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
-select * where {
-  ?class rdfs:subClassOf <http://dbpedia.org/resource/Electronic_component> .
-  bind(replace(strafter(str(?class),str(dbr:)),'_',' ') as ?class_name) .
-  OPTIONAL {
-    SERVICE <http://dbpedia.org/sparql> {
-      ?class rdfs:label ?title .
-      FILTER (
-        langMatches(lang(?title), 'ru')
-      )
-    }
-  }
-} limit 100
-\end{minted}
-  \caption{Example of a federated SPARQL query usage in an electronic component database}
-  \label{fig:sparql-ex1}
-\end{figure}
+% \begin{figure}%[!b]
+%   \centering
+% \begin{minted}{sparql}
+% PREFIX dbr: <http://dbpedia.org/resource/>
+% PREFIX owl: <http://www.w3.org/2002/07/owl#>
+% PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
+% select * where {
+%   ?class rdfs:subClassOf <http://dbpedia.org/resource/Electronic_component> .
+%   bind(replace(strafter(str(?class),str(dbr:)),'_',' ') as ?class_name) .
+%   OPTIONAL {
+%     SERVICE <http://dbpedia.org/sparql> {
+%       ?class rdfs:label ?title .
+%       FILTER (
+%         langMatches(lang(?title), 'ru')
+%       )
+%     }
+%   }
+% } limit 100
+% \end{minted}
+%   \caption{Example of a federated SPARQL query usage in an electronic component database}
+%   \label{fig:sparql-ex1}
+% \end{figure}
 
@@
 The inference is implemented in various ways.
-All of them belong either to semantic ones, usually implemented as solving a corresponding CSP\footnote{Abbreviation of Constraint Satisfaction Problem}, or syntax-driven procedures that mediate the previous structures and infer one of the specific properties (Automatic Theorem Proving).
-Theoretic research is being carried on classification the properties of vocabularies by the complexity of verification and development software implementing it for these classes.
-This software is a  part of RDF-processing libraries and stand-alone system extensions.
+All of them belong either to semantic ones, usually implemented as solving a corresponding CSP (Constraint Satisfaction Problem), or syntax-driven procedures that mediate the previous structures and infer one of the specific properties (Automatic Theorem Proving), modifying a KG.
+Theoretic research is being carried on classification % the properties
+of vocabularies by the complexity of verification and development software implementing it for these classes.  % This software is a  part of RDF-processing libraries and stand-alone system extensions.
 
@@
 
-In order to present the general architecture of application development, we have to speak about KG definitions and useful properties briefly of KGs.
+In order to present the general architecture of application development, %we have to speak about
+consider briefly
+KG definitions and its useful properties.
 % \subsection{Knowledge graph properties as application basis}
 % \label{sec:kg-as-basis}
-KGs changed the viewpoint to the knowledge base as being a part of the whole totality of knowledge, implying the obeying the global standards and techniques of its acquisition and processing \cite{hogan}.
-The following outlines their definitions \cite{hogan}.
+KGs changed the viewpoint to the knowledge base as being a part of the whole totality of knowledge, implying the obeying the %global %standards and
+techniques of its acquisition and processing \cite{hogan}.  The following outlines their definitions \cite{hogan}.
  \begin{itemize}
@@
   \item The same KG is being filled in and processed by different independent modules, \emph{e.g.}, withing SPARQL queries with UPDATE statements;
-  \item KG fundamentals allow a software developer to \emph{postpone} the definition of a data schema;
+  \item KG fundamentals allow %a software
+developer to \emph{postpone} the definition of a data schema;
   \item There are three points of view to graph schemata interpretation:
     \begin{itemize}
-    \item \emph{semantic}, which is aimed at domain modeling paying attention to generalization relations (classes);
-    \item \emph{validating}, reflecting semantic verification, checking completeness with respect to sets of predefined general relations, and
-    \item \emph{emergent}, observed or deduced from exploitation experience analysis, as a set of new generalized structures and \emph{refactoring} of the KG.
+    \item \emph{semantic} aimed at domain modeling %paying attention to
+describing
+generalization relations (classes);
+    \item \emph{validating} reflecting semantic verification, checking completeness with respect to sets of predefined general relations, and
+    \item \emph{emergent} observed or deduced from exploitation experience analysis as a set of new generalized structures and \emph{refactoring} of the KG.
     \end{itemize}
@@
 Such a way partially solves the main technical problems of KG storage and query execution.
-Being a triple set, whole KGs are usually stored in computer RAM, and their structures expose no heuristic information to query engines, which could be used for optimizations compared to relation databases technologies.
+Being a triple set, whole KGs are usually stored in computer RAM, and their structures expose no heuristic information for query engines, which could be used for optimizations.% compared to relation databases technologies.
 
@@
 Python allows to wrap whole subsystems, \emph{e.g.}, Elastic search, in imperative paradigm, and Logtalk is chosen for the declarative counterpart.
-Inter-process component communications is realized usually with \texttt{HTTP} protocol, representing data in XML and JSON formats.
+Inter-process component communications is realized usually with \texttt{HTTP} protocol.%, representing data in XML and JSON formats.
 
 Our R\&D resulted in a set of components and a general application architecture showed in Figure~\ref{fig:arch-services}.
-\begin{figure}[!b]
+\begin{figure}%[!b]
   \begin{center}
@@
 Kernel module stores a common global part of A-~and T-boxes of applications, \emph{i.e.}, it is application database.
-Other modules provide an application (C3) with services.
+Other modules provide an application (C3).
+% with services.
 
@@
 This composed page can be stored in a KG as a set of triples and indexed in the text indexing module (C1), printed with the browser's printing engine, or converted to PDF.
-The tool allows the user to change document content in WYSIWYG mode.
-Its general structure is defined by changing the template on the server-side.
+The tool allows the user to change document content in WYSIWYG mode,
+its general structure is controlled by templates.
+% . Its general structure is defined by changing the template on the server-side.
 
@@
 Sphinx Search is much faster in text and database records indexing and consumes less RAM as it is implemented in C++.
-The choice of an engine depends on the primary document representation used in the application.
+The choice of an engine depends on the primary document representation.
+% used in the application.
 
 BLOB stored documents having no markup (PDFs, scanned documents) usually contain valuable data to be revealed for application use.
-The documents could be report data or scientific paper content, DJVU files, or raster images.
-Data recognition is implemented in the PDF/text document processing module (C2), where data processing workers query the text indexing module for BLOB documents and add recognized text and high-order structure layers.
-All obtained layers are stored in the database of C1, and data of general interest are to be converted in triples of KG.
+The documents could be report data or scientific paper content, %DJVU files,
+or raster images.
+Recognition is implemented in the PDF/text document processing module (C2), where workers query the text indexing module for BLOB documents and add recognized text and high-order structure layers.
+All obtained layers are stored in the database of C1,
+and generally useful data are
+% and data of general interest are to be
+converted in triples of KG.
 
@@
 It allows us to store resultant data in KG for further usage in module (B3) for problem-solving and decision-making.
-Module A3 denote external services and KG with valuable resources, \emph{e.g.}, DBPedia global objects.
+Module A3 denote external services and KG with valuable resources.
+%, \emph{e.g.}, DBPedia global objects.
 
@@
 This approach leverages the known \verb|llvmlite| source code generation.
-Figure~\ref{fig:block-seq} shows the main idea behind PSM object synthesis.
-\begin{figure}[!b]
-  \centering
-\begin{multicols}{2}
-\begin{minted}[fontsize=\scriptsize]{logtalk}
-:- object(class, specializes(code_block),
-   imports([named])). % Category of named entities
-:- public([classlist/1, methods/1, attributes/1]).
-% . . . . . . . . . . . . . .
-renderitem(Item, Result):-      % proceed with default
-    ^^renderitem(Item, Result). % rendering
-render(Result):-         % Source generator
-    ^^render(Name),      % implemented in a category
-    ( ::item(classlist(List)) ->
-     % . . . . . . . . . . .
-        [Name]) ),
-    ( ::item(attributes(Attributes))->
-     % . . . . . . . . . . .
-        [DefAttrList]),
-      Attributes::items(InstanceAttrs),
-      findall(S, ( % initialize attributes
-         % . . . . . . . . .
-         ), AttrAssigns),
-        root::unindent,
-        AttrList=[ConstructorDef|AttrAssigns];
-         % . . . . . . . . .
-        AttrList=[ConstructorDef, Pass] ),
-    ( ::item(methods(Methods))-> % If any ...
-      Methods::render(MethodList);
-      MethodList=[] ),
-    lists::append(AttrList,MethodList,StringList),
-    root::unindent, Result=[Signature|StringList].
-:- end_object.
-\end{minted}
-      \includegraphics[width=0.7\linewidth]{code_block_class.pdf}
-\end{multicols}
-  \caption{Procedure of dataflow module source code model synthesis}
-  \label{fig:block-seq}
-\end{figure}
+%Figure~\ref{fig:block-seq} shows the main idea behind PSM object synthesis.
+% \begin{figure}%[!b]
+%   \centering
+% \begin{multicols}{2}
+% \begin{minted}[fontsize=\scriptsize]{logtalk}
+% :- object(class, specializes(code_block),
+%    imports([named])). % Category of named entities
+% :- public([classlist/1, methods/1, attributes/1]).
+% % . . . . . . . . . . . . . .
+% renderitem(Item, Result):-      % proceed with default
+%     ^^renderitem(Item, Result). % rendering
+% render(Result):-         % Source generator
+%     ^^render(Name),      % implemented in a category
+%     ( ::item(classlist(List)) ->
+%      % . . . . . . . . . . .
+%         [Name]) ),
+%     ( ::item(attributes(Attributes))->
+%      % . . . . . . . . . . .
+%         [DefAttrList]),
+%       Attributes::items(InstanceAttrs),
+%       findall(S, ( % initialize attributes
+%          % . . . . . . . . .
+%          ), AttrAssigns),
+%         root::unindent,
+%         AttrList=[ConstructorDef|AttrAssigns];
+%          % . . . . . . . . .
+%         AttrList=[ConstructorDef, Pass] ),
+%     ( ::item(methods(Methods))-> % If any ...
+%       Methods::render(MethodList);
+%       MethodList=[] ),
+%     lists::append(AttrList,MethodList,StringList),
+%     root::unindent, Result=[Signature|StringList].
+% :- end_object.
+% \end{minted}
+%       \includegraphics[width=0.7\linewidth]{code_block_class.pdf}
+% \end{multicols}
+%   \caption{Procedure of dataflow module source code model synthesis}
+%   \label{fig:block-seq}
+% \end{figure}
 This example of the application of KG utilizes strongly the property of data representation flexibility for description of the software object model data as well as one central point of data store and access.
@@
 Data access differentiation by security model is implemented programmatically as the web pages disallow access to the restricted information.
-SW has special predicates for describing a possibility and location of a private data sub-KG in the open part.
+SW has special predicates for describing locations of a private data sub-KG in the open part.
+%The In Figure~\ref{fig:auth-code}, we presented our syntactic extension of RDFa for referring template parts.
+A namespace \textbf{taa} was added to define a markup interpreted as XHTML tree manipulation in the image and likeliness of open-source ZPT\footnote{Abbreviation of Zope Page Template}.
+% Arguments of \verb|taa:content| are subcontexts (\verb|title| and \verb|neg-UMK|) of a current template under rendering.
 
-In Figure~\ref{fig:auth-code}, we presented our small syntactic extension of RDFa for referring template parts.
+%In Figure~\ref{fig:auth-code}, we presented our small syntactic extension of RDFa for referring template parts.
 A namespace \textbf{taa} was added to define a markup interpreted as XHTML tree manipulation in the image and likeliness of open-source ZPT\footnote{Abbreviation of Zope Page Template}.
 Arguments of \verb|taa:content| are subcontexts (\verb|title| and \verb|neg-UMK|) of a current template under rendering.
-\begin{figure}[!b]
-  \centering
-\begin{minted}[escapeinside=||,fontsize=\small]{text}
-<html lang="ru" xmlns=http://www.w3.org/1999/xhtml
-|\GB{xmlns:taa}|=http://irnok.net/engine/rdfa-manipulation
-xml:lang="ru" metal:define-macro="page">
-<head> . . . . </head>
-<body prefix="rdf: http://www.w3.org/1999/...-ns# foaf: http://xmlns.com/foaf/...
-imei: imei.html# course: https://irnok.net/college/plan/01..16-...\
-% D0\%BA_PB-SM.plm.xml.xlsx-....2.3.1.html#"  resource="#post"
-typeof="schema:CreativeWork sioc:Post prov:Entity">
-<!-- The application control panel -->
+% \begin{figure}[!b]
+%   \centering
+% \begin{minted}[escapeinside=||,fontsize=\small]{text}
+% <html lang="ru" xmlns=http://www.w3.org/1999/xhtml
+% |\GB{xmlns:taa}|=http://irnok.net/engine/rdfa-manipulation
+% xml:lang="ru" metal:define-macro="page">
+% <head> . . . . </head>
+% <body prefix="rdf: http://www.w3.org/1999/...-ns# foaf: http://xmlns.com/foaf/...
+% imei: imei.html# course: https://irnok.net/college/plan/01..16-...\
+% % D0\%BA_PB-SM.plm.xml.xlsx-....2.3.1.html#"  resource="#post"
+% typeof="schema:CreativeWork sioc:Post prov:Entity">
+% <!-- The application control panel -->
 
-<main lang="ru" resource="#annotation" typeof="oa:Annotation" id="main-doc-cnt">
-<div property="oa:hasTarget" resource="#course-work-prog"></div>
-<article property="oa:hasBody" typeof="foaf:Document curr:WorkingProgram"
-         resource="#course-work-program" id="main-document">
-  <div |\GB{taa:content}|="imei:title-page"></div>
-  <div |\GB{taa:content}|="imei:neg-UMK"></div>
-  <section id="TOC" class="break-after"> <h2>Table of Contents</h2>
-    <div id="tableOfContents"></div>
-  </section>
-  <section id="course-description" resource="#description"
-           property="schema:hasPart" typeof="schema:CreativeWork">
-    <div property="schema:hasPart" resource="#purpose"
-         typeof="dc:Text cnt:ContentAsText" >
-      <div property="cnt:chars" datatype="xsd:string">
-        <h2 property="dc:title" datatype="xsd:string">
-           Aims and objectives of the discipline (module)</h2>
-        <p>The aim of teaching the discipline ...</p>
-      </div>
-   </div>
-  . . . . . . . .
-\end{minted}
-  \caption{Part of a template representation of an educational document}
-  \label{fig:auth-code}
-\end{figure}
+% <main lang="ru" resource="#annotation" typeof="oa:Annotation" id="main-doc-cnt">
+% <div property="oa:hasTarget" resource="#course-work-prog"></div>
+% <article property="oa:hasBody" typeof="foaf:Document curr:WorkingProgram"
+%          resource="#course-work-program" id="main-document">
+%   <div |\GB{taa:content}|="imei:title-page"></div>
+%   <div |\GB{taa:content}|="imei:neg-UMK"></div>
+%   <section id="TOC" class="break-after"> <h2>Table of Contents</h2>
+%     <div id="tableOfContents"></div>
+%   </section>
+%   <section id="course-description" resource="#description"
+%            property="schema:hasPart" typeof="schema:CreativeWork">
+%     <div property="schema:hasPart" resource="#purpose"
+%          typeof="dc:Text cnt:ContentAsText" >
+%       <div property="cnt:chars" datatype="xsd:string">
+%         <h2 property="dc:title" datatype="xsd:string">
+%            Aims and objectives of the discipline (module)</h2>
+%         <p>The aim of teaching the discipline ...</p>
+%       </div>
+%    </div>
+%   . . . . . . . .
+% \end{minted}
+%   \caption{Part of a template representation of an educational document}
+%   \label{fig:auth-code}
+% \end{figure}
 
@@
 HTML5 default tag structures are used for visual sectioning and style definition.
-
 The \textbf{oa} vocabulary is also used to define parts of a document, which allow interactive editing after the document has been rendered.
@@
 For this purpose, data conversion from the tabular form to KG was carried out, a CliopPatria server for data storage was installed, and data viewers for the triple sets for a typical subject (a fault) have been implemented.
-In Figure~\ref{fig:gis-ex} an example of the filtering and a selected object data viewing form is presented.
-\begin{figure}[!b]
-  \centering
-  \includegraphics[width=\linewidth]{faults-leaflet-doc.png}
-  % \caption{Viewer window showing description of a fault}
-  \caption{Example of GIS rendering of KG data}
-  \label{fig:gis-ex}
-\end{figure}
+%In Figure~\ref{fig:gis-ex} an example of the filtering and a selected object data viewing form is presented.
+% \begin{figure}%[!t]
+%   \centering
+%   \includegraphics[width=\linewidth]{faults-leaflet-doc.png}
+%   % \caption{Viewer window showing description of a fault}
+%   \caption{Example of GIS rendering of KG data}
+%   \label{fig:gis-ex}
+% \end{figure}
 
@@
 %%%%%%%%% Конец текста статьи %%%%%%%%%%%%%
-\pagebreak
+% \pagebreak  % Растягивает предыдцщую страницу, что не красиво.
+\newpage
 
@@
 
-The~R\&D on GIS-viewer of faults involved the Centre of Geodynamics and Geochronology equipment at the Institute of the Earth's Crust, Siberian Branch of the Russian Academy of Sciences (grant No.~075-15-2021-682).
+%The~R\&D on GIS-viewer of faults involved the Centre of Geodynamics and Geochronology equipment at the Institute of the Earth's Crust, Siberian Branch of the Russian Academy of Sciences (grant No.~075-15-2021-682).
 This study direction was partially carried out within the basic budgetary research project ``Modern geodynamics, mechanisms of destruction of the lithosphere and hazardous geological processes in Central Asia'', No.~FWEF-2021-0009.
 
-The~development of the~infrastructure for Mothur command transformation to Rapidminer dataflow diagram is supported by the Irkutsk scientific center of Siberian Branch of Russian Academy of Sciences, grant No 4.2.
+%The~development of the~infrastructure for Mothur command transformation to Rapidminer dataflow diagram is supported by the Irkutsk scientific center of Siberian Branch of Russian Academy of Sciences, grant No 4.2.
 
@@
 
-\bibitem{swi} J.~Wielemaker, T.~Schrijvers, M.~Triska, T.~Lager. SWI-Prolog. Theory and Practice of Logic Programming, volume~12, No.~1-2, pp.~67--96, 2011, ISSN~1471-0684.
+% \bibitem{swi} J.~Wielemaker, T.~Schrijvers, M.~Triska, T.~Lager. SWI-Prolog. Theory and Practice of Logic Programming, volume~12, No.~1-2, pp.~67--96, 2011, ISSN~1471-0684.
 
@@
 
-\bibitem{lunina} O.~V.~Lunina.
+% \bibitem{lunina} O.~V.~Lunina.
 The digital map of the Pliocene–Quaternary crustal faults in the Southern East Siberia and the adjacent Northern Mongolia.
@@
 (in Russian) \doi{10.5800/GT-2016-7-3-0215}
-
-\bibitem{afs} A.~A.~Gladkov, O.~V.~Lunina. Cartographic service ``Activetectonics''. URL:\url{http://activetectonics.ru/} (access date: 20-Sep-2021)
+% \bibitem{afs} A.~A.~Gladkov, O.~V.~Lunina. Cartographic service ``Activetectonics''. URL:\url{http://activetectonics.ru/} (access date: 20-Sep-2021)
 
@@
 
-%\bibitem{lgd} C.~Stadler, J.~Lehmann, K.~Höffner, S.~Auer. LinkedGeoData: A core for a web of spatial open data. Semantic Web 3 (2012) 333–354. \doi{10.3233/SW-2011-0052}
+%\bibitem{lgd} C.~Stadler, J.~Lehmann, K.~Hffner, S.~Auer. LinkedGeoData: A core for a web of spatial open data. Semantic Web 3 (2012) 333–354. \doi{10.3233/SW-2011-0052}
 

### 2022-11-27 paper-aicts-jcart.tex
Release corrected

@@
 This paper presents a view on the technologies as data representation for software development tools based on model analysis, and transformation, %and retrospection capabilities,
-examples are provided.  %Three examples show the technologies usage, exposing our experience.
+examples are provided.  %Three examples show the technologies' usage, exposing our experience.
 % \end{abstract}
@@
 \end{enumerate}
-Some of the assets % \ref{kg1}--\ref{kg2}
+Some assets % \ref{kg1}--\ref{kg2}
 form knowledge graph (KG) technological aspect of the SW that gained a particular interest within R\&D in the last decade.
@@
 Subject and predicate are always IRIs, whereas objects are either IRI (a reference to a resource) or a literal.
-Literals are values of basic types (\emph{e.g.}, number, string, date, time) usually accompanied with metadata (type designator or a language tag).
+Literals are values of basic types (\emph{e.g.}, number, string, date, time) usually accompanied by metadata (type designator or a language tag).
 Types are defined in a vocabulary \texttt{<\url{http://www.w3.org/2001/XMLSchema}\#>}, having well-known namespace abbreviation \ns{xs}\footnote{Hereafter, all well-known namespace names will be written with a bold monospace font.}.
@@
 Much R\&D efforts since 2001 \cite{tbl} were spent to develop SW data formats and their coupling with document representation.
-The simplest way to provide a printed document with semantic representation is by pointing the user's software agent to an URL serving the semantic data.
+The simplest way to provide a printed document with semantic representation is by pointing the user's software agent to a URL serving the semantic data.
 However, this approach requires implementing two different procedures for document generation and its metadata serving.
@@
 Zope component architecture\footnote{On-line book is at \url{https://muthukadan.net/docs/zca.html}}, Logtalk, React.js are used to \emph{implement} interfaces.
-Python allows to wrap whole subsystems, \emph{e.g.}, Elastic search, in imperative paradigm, and Logtalk is chosen for the declarative counterpart.
+Python allows wrapping whole subsystems, \emph{e.g.}, Elastic search, in imperative paradigm, and Logtalk is chosen for the declarative counterpart.
 Inter-process component communications is realized usually with \texttt{HTTP} protocol.%, representing data in XML and JSON formats.
@@
 The kernel module of applications is KG server (B2) built on the base of ClioPatria \cite{cliopatria} with web-server and Pengines plug-ins.
-ClioPatria is a web-application implemented in SWI-Prolog \cite{swi} being ISO-compliant Prolog implementation.
+ClioPatria is a web-application implemented in SWI-Prolog being ISO-compliant Prolog implementation.
 KG functions are implemented with SW package, supplied in standard SWI-Prolog distribution.
@@
 
-This project \cite{authoring} is aimed at the construction of an environment for document authoring out of user-editable template elements, database query results, LOD-compliant web pages, and other kind of a document content.
+This project \cite{authoring} is aimed at the construction of an environment for document authoring out of user-editable template elements, database query results, LOD-compliant web pages, and other document content.
 The target document is a web page where content is inferred while loaded by the browser.
@@
   \item DBPedia resource (\textbf{dbr}), references to instant objects and classes;
-  \item Schema.org (\textbf{schema}), Google, Yandex, Yahoo, \emph{etc}. searchable objects, structural elements;
+  \item Schema.org (\textbf{schema}), Google, Yandex, Yahoo, \emph{i.e.}, searchable objects, structural elements;
   \item The Bibliographic Ontology (\textbf{bibo}), literature reference mark up.
@@
 
-The general logical structure (sections, subsections, \emph{etc.}) are represented with Open Annotation (\textbf{oa}) vocabulary originally used for annotation of already existing content.
+The general logical structure (sections, subsections) is represented with Open Annotation (\textbf{oa}) vocabulary originally used for annotation of already existing content.
 The generated document conversion into PDF and other formats is carried out under the control of \textbf{oa}.
@@
 Project \cite{gisviewer} devoted to investigating capabilities of SW and KG technologies in representation spatially distributed data (GIS\footnote{Abbreviation of Geographical Information System}-data).
-The fault data \cite{lunina,afs} are chosen as a subject for representation and publishing in this investigation.
+The fault data \cite{gisviewer} are chosen as a subject for representation and publishing in this investigation.
 In geology, researchers accumulate data obtained after event observations, \emph{e.g.}, earthquakes, landslides, by analyzing remote sensing data and results of field works.
@@
 
-\bibitem{tbl} T.~Berners-Lee, J.~Hendler, O.~Lassila.
+\bibitem{tbl} Berners-Lee~T., Hendler~J., Lassila~O.
 The semantic web: A new form of web content that is meaningful to computers will unleash a revolution of new possibilities.
-Scientific American, May 2001.
+Scientific American.
+2001; 284:~1--5
 
-\bibitem{hogan} A.~Hogan, E.~Blomqvist, M.~Cochez, C.~D’Amato \emph{et al}. Knowledge Graphs, 2020. URL:\url{https://arxiv.org/abs/2003.02320v5} (access date: 12-Dec-2021)
+\bibitem{hogan} Hogan~A., Blomqvist~E., Cochez~M., D’Amato~C., \emph{et al}. Knowledge Graphs. 2020. Available~at:~\url{https://arxiv.org/abs/2003.02320v5} (Accessed December 12, 2021)
 
-\bibitem{zont19} E.~Cherkashin, A.~Shigarov, V.~Paramonov.
+\bibitem{zont19} Cherkashin~E., Shigarov~A., Paramonov~V.
 Representation of MDA transformation with logical objects.
-Procs.
-of International Multi-Conference on Engineering, Computer and Information Sciences (SIBIRCON), Novosibirsk, Russia, 2019, pp.~0913--0918 \doi{10.1109/SIBIRCON48586.2019.8958008}
+Proc.
+of ``International Multi-Conference on Engineering, Computer and Information Sciences (SIBIRCON)''.
+Novosibirsk, Russia; 2019:~0913--0918 \doi{10.1109/SIBIRCON48586.2019.8958008}
 
-\bibitem{cliopatria} J.~Wielemaker, W.~Beek, M.~Hildebrand, J.~Ossenbruggen, ClioPatria: A SWI-Prolog infrastructure for the Semantic Web, Semantic Web, 2016, volume 7(5).
-pp.~529--541 \doi{10.3233/SW"=150191}
+\bibitem{cliopatria} Wielemaker~J., Beek~W., Hildebrand~M., Ossenbruggen~J. ClioPatria: A SWI-Prolog infrastructure for the Semantic Web. Semantic Web. 2016; 7(5):~529--541 \doi{10.3233/SW-150191}
 
-% \bibitem{swi} J.~Wielemaker, T.~Schrijvers, M.~Triska, T.~Lager. SWI-Prolog. Theory and Practice of Logic Programming, volume~12, No.~1-2, pp.~67--96, 2011, ISSN~1471-0684.
+\bibitem{pengines} Lager~T., Wielemaker~J. Pengines: Web Logic Programming Made Easy.  Theory and Practice of Logic Programming. 2014; 14(4-5):~539--552. \doi{10.1017/S1471068414000192}
 
-\bibitem{pengines} T.~Lager, J.~Wielemaker.
-Pengines: Web Logic Programming Made Easy.
-Theory and Practice of Logic Programming, volume~14, No.~4-5, 2014, pp.~539--552.
-\doi{10.1017/S1471068414000192}
+\bibitem{lod} Bizer~C., Heath~T., Berners-Lee~T., Linked Data -- The Story So Far.  Int. J. Semantic Web Inf. Syst. 2009; 5:~1--22. \doi{10.4018/jswis.2009081901}
 
-\bibitem{lod} C.~Bizer, T.~Heath, T.~Berners-Lee, Linked Data -- The Story So Far.  Int. J. Semantic Web Inf. Syst., volume~5, 2009, pp.~1--22. \doi{10.4018/jswis.2009081901}
+\bibitem{aicts2020} Cherkashin~E., Shigarov~A.
+Construction techniques of Baikal microbiome research information-computational environment.
+Proc.
+of ``1st International Workshop on Advanced Information and Computation Technologies and Systems''.
+December 7--11.
+Irkutsk, Russia.
+CEUR-WS.org; 2020.
+Available~at:~\url{http://ceur-ws.org/Vol-2858/paper2.pdf} (Accessed December 12, 2021)
 
-\bibitem{aicts2020} E.~Cherkashin, A.~Shigarov.
-Construction techniques of Baikal microbiome research information-computational environment.
-Procs.
-of 1st International Workshop on Advanced Information and Computation Technologies and Systems.
-Irkutsk, Russia, December 7-11, 2020.
-CEUR-WS.org, URL:\url{http://ceur-ws.org/Vol-2858/paper2.pdf} (access date: 12-Dec-2021)
-
-\bibitem{authoring} E.~Cherkashin, A.~Shigarov, V.~Paramonov, A.~Mikhailov, Digital archives supporting document content inference, Procs.
-of 42-nd International Convention on Information and Communication Technology Electronics and Microelectronics (MIPRO), May 20–24, 2019.
-pp.
-1037-1042.
+\bibitem{authoring} Cherkashin~E., Shigarov~A., Paramonov~V., Mikhailov~A., Digital archives supporting document content inference.
+Proc.
+of ``42-nd International Convention on Information and Communication Technology Electronics and Microelectronics (MIPRO)''.
+May 20--24.
+Opatia, Chroatia; 2019; 1037--1042.
 \doi{10.23919/MIPRO.2019.8757196}
 
-\bibitem{gisviewer} E.~Cherkashin, O.~Lunina, L.~Demyanov, A.~Tsygankov.
-Web-GIS viewer for active faults data represented as a knowledge graph.
-Procs.
-of 4th Scientific-practical Workshop Information Technologies: Algorithms, Models, Systems.
-Irkutsk, Russia, September 14, 2021.
-CEUR-WS.org, URL:\url{http://ceur-ws.org/Vol-2984/paper8.pdf} (access date: 12-Dec-2021)
+\bibitem{gisviewer} Cherkashin~E., Lunina~J., Demyanov~L., Tsygankov~A.
+Web--GIS viewer for active faults data represented as a knowledge graph.
+Proc.
+of ``4th Scientific-practical Workshop Information Technologies: Algorithms, Models, Systems''.
+September 14.
+Irkutsk, Russia, CEUR-WS.org; 2021.
+Available~at:~\url{http://ceur-ws.org/Vol-2984/paper8.pdf} (Accessed December 12, 2021)
 
-% \bibitem{lunina} O.~V.~Lunina.
-The digital map of the Pliocene–Quaternary crustal faults in the Southern East Siberia and the adjacent Northern Mongolia.
-Geodynamics \& Tectonophysics.
-volume~7(3), 2016.
-pp.~407-434.
-(in Russian) \doi{10.5800/GT-2016-7-3-0215}
-% \bibitem{afs} A.~A.~Gladkov, O.~V.~Lunina. Cartographic service ``Activetectonics''. URL:\url{http://activetectonics.ru/} (access date: 20-Sep-2021)
-
-\bibitem{shigpaper} N.~O.~Dorodnykh, A.~Y.~Yurin, A.~O.~Shigarov.
+\bibitem{shigpaper} Dorodnykh~N., Yurin~A., Shigarov~A.
 Conceptual Model Engineering for Industrial Safety Inspection Based on Spreadsheet Data Analysis.
@@
 MDIS 2019.
-Communications in Computer and Information Science, volume~1126, 2020, Springer, Cham.
+Communications in Computer and Information Science.
+Springer, Cham.
+2020; 1126.
 \doi{10.1007/978-3-030-39237-6_4}
 
-%\bibitem{foss} M.~Leidig, R.~Teeuw.
-Free software: A review, in the context of disaster management.
-International Journal of Applied Earth Observation and Geoinformation, 42 (2015), pp.~49-56.
-\doi{10.1016/j.jag.2015.05.012}.
+\end{thebibliography}
+% \section{On-line Resources}
 
-%\bibitem{lgd} C.~Stadler, J.~Lehmann, K.~Hffner, S.~Auer. LinkedGeoData: A core for a web of spatial open data. Semantic Web 3 (2012) 333–354. \doi{10.3233/SW-2011-0052}
-
-%\bibitem{geolink} M.~Cheatham, A.~Krisnadhi, R.~Amini, P.~Hitzler, \emph{et al}. The GeoLink knowledge graph, Big Earth Data, 2:2 (2018), pp.~131-143. \doi{10.1080/20964471.2018.1469291}
-
-%\bibitem{abid} T.~Abid, H.~Zarzour.
-Integrating linked open data in a geographical information system.
-Procs.
-of.
-International Conference on Information Technology for Organization Development.
-Oct 19-20, 2014, University of Tebessa, Tebessa, Algeria (2014).
-
-%\bibitem{iwaniak1} A.~Iwaniak, I.~Kaczmarek, M.~Strzelecki, J.~Lukowicz, P.~Jankowski.
-Enriching and improving the quality of linked data with GIS.
-Open Geosciences, Vol.~8, 1~(2016) pp.~323-336.
-\doi{10.1515/geo-2016-0020}
-
-%\bibitem{iwaniak17} A.~Iwaniak, M.~Leszczuk, M.~Strzelecki, F.~Harvey, I.~Kaczmarek.
-A Novel Approach for Publishing Linked Open Geodata from National Registries with the Use of Semantically Annotated Context Dependent Web Pages.
-International Journal of Geo-Information.
-6, 252 (2017).
-\doi{10.3390/ijgi6080252}
-
-\end{thebibliography}
-% \section{Online Resources}
-
-% The sources for the viewer are being developed at Github, URL:~\url{https://github.com/De17eon/GRL}.
-
+% The sources for the viewer are being developed at GitHub, URL:~\url{https://github.com/De17eon/GRL}.
+\label{pageend}
 \end{document}
